### 2024-10-17 Article/main.tex
Initial commit

\documentclass[a4paper, 11pt]{article}
\usepackage[utf8]{inputenc}

\usepackage[breaklinks, colorlinks=true, citecolor=black, linkcolor=blue, urlcolor=blue, filecolor=blue]{hyperref}

\usepackage{comment}

\usepackage[center, labelfont=bf]{caption}
\captionsetup{justification=justified, singlelinecheck=false} % Justify left

\usepackage{amsmath} % Math equation

% For table
\usepackage{multirow}
\usepackage[table, xcdraw]{xcolor}

% Page definitions
\setlength{\topmargin}{1.5cm}
\setlength{\headheight}{1.1\baselineskip}
\setlength{\headsep}{20pt}
\setlength{\topskip}{12pt}
\setlength{\evensidemargin}{0pt}
\setlength{\oddsidemargin}{0pt}
\setlength{\textheight}{240mm}
\setlength{\textwidth}{160mm}
\setlength{\voffset}{-2cm}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\usepackage{natbib} % Bibliography
\bibliographystyle{unsrt}

\usepackage{graphicx}
\graphicspath{ {../Figures/} }

%opening
\title{The Impact of Model Assumptions on RES Drought Analysis}
\author{Boris Morin, Damian Flynn, Aina Maimó Far, Conor Sweeney}

\begin{document}

\maketitle

\begin{abstract}

Keywords: 
\end{abstract}

\section{Introduction}

As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy. Under the revised European Green Deal \cite{greendeal2023report}, the EU has set ambitious targets to achieve at least 69\% of electricity generation from renewable energy sources (RES) by 2030. This is a significant increase from 2022, when RES accounted for over 41\% of the EU’s electricity generation \cite{europe2023stat}. Meeting this goal is a key step in reducing greenhouse gas emissions and ensuring a sustainable energy future.

However, as reliance on RES grows, so does the challenge of managing the variability of weather-dependent energy sources like wind and PhotoVoltaic (PV) power. This challenge is further intensified by the increasing electrification of energy sectors which places additional demand on the power system~\cite{bloomfield2021}. As more sectors shift from fossil fuels to electricity, the power grid becomes increasingly sensitive to meteorological conditions~\cite{bloomfield2016, vanderWiel2019drought}. In this context, periods of low renewable generation often referred to as \textit{Dunkelflaute} or RES droughts, pose a greater risk to grid stability and energy security. Ensuring a reliable electricity supply during these RES droughts is crucial for maintaining a resilient energy system capable of meeting both the growing demand for electricity and the EU’s decarbonization targets.

For this study, a RES drought event is defined when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology established in previous research~\cite{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}. Other studies have employed varying methodologies for defining RES droughts. One approach uses relative thresholds that change over the course of the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}. Another common method relies on quantile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\cite{bracken2024drought, allen2023drought}. Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\cite{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}. In this paper, the focus is exclusively on energy generation, using a fixed threshold approach to define RES droughts.

RES drought events can be identified from wind and PV CF time series. In this study, we use three different datasets, all of them driven by ERA5 data~\cite{hersbach2020era5}. The first two datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy} (C3S). One of the C3S-E datasets provides CF time series aggregated at the national scale. Another C3S-E dataset provides the CF time series at each grid point, at the ERA5 resolution of 0.25°. The third dataset is produced by the authors using the Atlite model~\cite{hofman2021atlite}, which converts the  ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.

\begin{comment}
The C3S-E national dataset assumes RES farms exist at every grid point, whereas for the C3S-E Gridded and Atlite datasets, the location of each RES farm is applied. \textcolor{magenta}{Do you think you could add somehow that the C3S-E datasets assume specific wind and PV conversion models? Perhaps this should not be in the introduction, but since farms are mentioned I would include the model as well.}
\end{comment}

This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios. The first represents the current energy mix, predominantly driven by wind, while the second is based on projected energy scenarios, featuring a more balanced share between wind and PV. A consistent methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.

\begin{comment}
	 \textcolor{magenta}{Did we say we are avoiding temporal references in scenarios? I get that ignoring what they represent is not the way to go but the line is very fine I would say}
\end{comment}

The datasets used in this study are detailed in section~\ref{sec:data}, which describes their sources, characteristics, and relevance for evaluating RES droughts. Section~\ref{sec:methods} outlines the energy models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied. In section~\ref{sec:analysis}, an in-depth analysis of RES drought occurrences is presented for the two energy scenarios, highlighting key differences across the datasets and the effects of varying energy capacities. Finally, section~\ref{sec:DiscConc} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.

\section{Data}
\label{sec:data}

To construct and validate the models, this study uses publicly available datasets from EirGrid, the Republic of Ireland's transmission system operator, and the ERA5 reanalysis dataset. These datasets provide necessary inputs for estimating the CF for wind and PV energy generation.

\subsection{EirGrid}
\label{sec:eirgrid}

EirGrid provides detailed datasets for all wind and PV farms in the Republic of Ireland and Northern Ireland (collectively referred to as Ireland) from 1990 to the present~\cite{eirgrid}. These datasets include information on each farm’s installed capacity, name, and connection date. To improve these datasets, the GPS coordinates for each farm were manually determined using online search.

EirGrid also provides a wide range of energy system variables for wind (2014-present) and PV (2018-present) energy generation. For our study, we use generation which refers to the actual amount of energy produced by the wind or PV farms. This is then combined with the installed capacity to compute the CF. \textcolor{cyan}{Change}

\begin{comment}
boris: we use availability which indicates the potential energy output if the farms were operating at full capacity 
\end{comment}

\subsection{Atmospheric Variables}
\label{sec:era5}

Atlite and C3S-E datasets are driven by the ERA5 reanalysis dataset~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF). This global gridded dataset provides hourly atmospheric variables from 1940 to the present. It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}. Table~\ref{tab:var_name} lists the ERA5 variables used in the models to generate the Atlite and C3S-Energy datasets.

\begin{table}[h!]
	\centering
	\begin{tabular}{|l|c|}
		\hline
		{\textbf{ERA5 name}}      & \textbf{variable} \\ \hline
		100 metre zonal and meridional wind speed   & $u_{100}$, $v_{100}$ \\
		2 metre temperature                         & $t2m$ \\
		Surface net solar radiation                 & $ssr$ \\
		Surface solar radiation downwards           & $ssrd$  \\
		Top of atmosphere incident radiation        & $tisr$  \\
		Total sky direct solar radiation at surface & $fdir$  \\ \hline
	\end{tabular}
	\caption{Variables retrieved from ERA5 dataset}
	\label{tab:var_name}
\end{table}

\subsection{C3S Energy}
\label{sec:c3se}

The Copernicus Climate Change Service (\textcolor{cyan}{I have defined C3S-E before, should I use it here ?} \textcolor{magenta}{I am not against introducing some acronyms twice, same as for some references being in the introduction and data, but I would say there is no right or wrong way necessarily} )has developed a renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models. This dataset provides hourly CF for wind and PV energy from 1979 to the present. The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°. Data are also available for download at two aggregated scales: regional (NUTS 2) and national.

The C3S-E dataset estimates wind energy using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 m. This choice is based on expert advice and actual trend in wind turbine installation. The PV generation model used by C3S-E uses two ERA5 variables: global horizontal irradiance (referred to here as $ssrd$) and air temperature ($t2m$). PV generation is calculated multiple times using the same model with different azimuth and tilt angles. The results are aggregated based on a statistical distribution of the module angles~\cite{saintdrenan2018solar}. \textcolor{cyan}{I added a sentence to explain the reason of the Vestas choice} \textcolor{magenta}{So, basically what they are saying is that they are looking at a realistic turbine for the future, makes sense. Not necessarily linked to the article, but I wonder how realistic this turbine choice is for other countries. We know it is not for Ireland, but I am curious about the rest of Europe}

\section{Methods}
\label{sec:methods}

In this study, we use three datasets to analyse RES droughts. Data downloaded from C3S-E are used to obtain two datasets: one based on national-level (C3S-E N), and another on grid-level (C3S-E G). The third dataset is computed using the Atlite model (Atlite).

\subsection{C3S-E N}
\label{sec:C3SEN}

For national-level analyses, we use the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0). We compute a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland. This CF time series for Ireland is based on the underlying C3S-E assumption of RES generation at every 0.25\textdegree grid point in Ireland.

\subsection{C3S-E G}
\label{sec:C3SEG}

For the second dataset, we use the gridded dataset from C3S-E to create a CF time series which is representative of the actual RES farm distribution. Using the farms' coordinates, we identify the nearest grid point on the C3S-E grid for each farm. We retrieve the CF values from the C3S-E dataset for its corresponding grid point. We then average the CF for each farm using the installed capacity as weights. This process results in a CF time series for each renewable energy source (wind and PV), which is representative of the location of the RES farms in Ireland.

\subsection{Atlite} 
\label{sec:atlite}

Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite}, and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}. This model transforms weather data into energy data, using the locations of existing RES farms as described in C3S-E G. A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.

For wind power estimation, Atlite uses the wind speed at 100 metres ($u_{100}$, $v_{100}$). We provide Atlite with a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake} at a hub height of 100 metres, which it uses when converting wind speed to CF.

For PV power estimation, Atlite uses four irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) as well as the air temperature ($t2m$). We use a single PV model~\cite{beyer2004pv}, with the Kaneka Hybrid panel option. The azimuth angle is fixed at 180\textdegree (due south) and we select the optimal tilt angle option.

\subsection{Energy Scenarios}
\label{sec:scenarios}

This study investigates wind and PV CF using three models, resulting in six CF time series with the installed capacity at the end of 2023. Additionally, we analyse the combined CF of wind and PV for each model by averaging the two sources, weighted by their capacities (6.7 GW for wind, 0.6 GW for PV). This approach gives us a total of nine CF time series for the 2023 scenario.

Given the relatively low PV capacity in Ireland in 2023 and the projected increase by 2030 (expected to increase from 0.6 GW to 8 GW), we also explore a future scenario with higher PV capacity. We generate additional CF time series reflecting the projected PV capacity in 2030 for both the Atlite and C3S-E G models, based on a roadmap published by EirGrid~\cite{eirgrid2023future}.

The wind capacity is projected to increase to 9 GW in 2030, however as the spatial distribution is not expected to significantly change we use the same CF time series. The combination of wind and PV CF is calculated once again using the new PV time series, and by averaging with the future projected capacities of RES farms. In total, we evaluate fourteen CF time series across different scenarios and models, as summarised in Table~\ref{tab:datasets}.

\begin{table}[h!]
	\centering
	\begin{tabular}{l|ccc|ccc|}
		\cline{2-7}
		& \multicolumn{3}{c|}{\textbf{Current}}                                   & \multicolumn{3}{c|}{\textbf{Projected}}                                   \\ \cline{2-7} 
	    & \multicolumn{1}{c|}{Atlite} & \multicolumn{1}{c|}{C3S-E G} & C3S-E N & \multicolumn{1}{c|}{Atlite} & \multicolumn{1}{c|}{C3S-E G} & C3S-E N \\ \hline
		\multicolumn{1}{|l|}{\textbf{PV}}      & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       &         \\
		\multicolumn{1}{|l|}{\textbf{Wind}}    & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{}       & \multicolumn{1}{c|}{}        &         \\
		\multicolumn{1}{|l|}{\textbf{Combine}} & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       \\ \hline
	\end{tabular}
	\caption{Summary of the time series compared in this study}
	\label{tab:datasets}
\end{table}

\subsection{RES drought definition}
\label{sec:res_drought}

In this study, we define a RES drought event as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours. Using the moving average captures fewer, but longer-lasting events than using the regular time series.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{data_methods/find_res_droughts}
	\caption{Time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021. CF threshold (dashed black). The grey bar shows the period identified as a wind drought.}
	\label{fig:find_res_droughts}
\end{figure}

Fig.~\ref{fig:find_res_droughts} illustrates the moving average  identification method. When using the moving average method, this CF time series results in one event being identified, which lasts 161 hours. In contrast, using the regular CF series would have resulted in three separate events being identified, with durations of 36, 35, and 40 hours.

\section{Results}
\label{sec:results}

\subsection{Verification}
\label{sec:verification}

Before going into the analysis of RES droughts, it is important to first verify the accuracy of the datasets used in this study. For the verification process, we use time-varying values of installed capacity to account for changes in RES development over the verification period. This step allows us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.

\subsubsection{Wind}
\label{sec:wind_verification}

The C3S-E datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve} a). The Atlite model allows the user to specify the power curve. We considered the 121 power curves available for download from Renewables.ninja~\cite{staffell2016wake}. For each power curve, Renewables.ninja also provide four associated smoothed power curves. The smoothing is done by using a Gaussian filter with different standard deviations that depend on the wind speed. We generated a wind CF time series for Ireland for each of these wind turbine power curve options. The performance of each wind turbine power curve was calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve. The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023. Based on these metrics, the most representative power curve for Ireland was the Enercon E112.4500 power curve with the $0.3w$ filter.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification/power_curve}
	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
	\label{fig:power_curve}
\end{figure}

The smoothing of the wind turbine power curve represents the wake effect, which is important when modelling wind energy on larger scales. The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind data from EirGrid.

\begin{table}[!ht]
	\centering
	\begin{tabular}{l|lll|}
		\cline{2-4}
		& \textbf{Atlite} & \textbf{C3S-E G} & \textbf{C3S-E N} \\ \hline
		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.981           & 0.972            & 0.970            \\ \hline
		\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.045           & 0.177            & 0.162            \\ \hline
		\multicolumn{1}{|l|}{\textbf{MBE}}   & -0.003          & 0.137            & 0.121            \\ \hline
	\end{tabular}
	\caption{Skill scores for wind for the three datasets compared to EirGrid data}
	\label{tab:wind_skill_scores}
\end{table}

The effect of the differences in the power curves are also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data. In contrast, the two C3S-E datasets tend to overestimate the EirGrid CF. This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification/wind_verification_contour}
	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
	\label{fig:wind_verification_contour}
\end{figure}

Fig.~\ref{fig:bar_number_events_verification_wind} presents results for the average annual number of wind drought events during the 2014 to 2023 validation period. The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification/bar_number_events_verification_wind}
	\caption{Average annual number of wind drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple) / Change plot}
	\label{fig:bar_number_events_verification_wind}
\end{figure}

\subsubsection{PV}
\label{sec:pv_verification}

The C3S-E datasets calculate PV generation multiple times using the same model with varying azimuth and tilt angles. The results are then aggregated based on a statistical distribution of module angles. Similarly to the wind CF verification, the Atlite model allows the user to select from one of three different PV panel characteristics. Following the same methodology as in the previous section, we compare the three available models using four skill scores (CC, RMSE, MB, and area under the curve). Based on the best-performing metrics, we selected the Breyer PV panel model \cite{beyer2004pv}, specifically using the Kaneka Hybrid panel option.

Figure \ref{fig:solar_verification_contour} shows that the three datasets have similar correlation when compared to the EirGrid data, which is further supported by the results shown in Table~\ref{tab:pv_skill_scores}.

\begin{table}[!ht]
	\centering
	\begin{tabular}{l|lll|}
	\cline{2-4}
	& \textbf{Atlite} & \textbf{C3S-E G} & \textbf{C3S-E N} \\ \hline
	\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.913            \\ \hline
	\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.070           & 0.065            & 0.100            \\ \hline
	\multicolumn{1}{|l|}{\textbf{MBE}}   & 0.002           & -0.012           & -0.013           \\ \hline
	\end{tabular}
	\caption{Skill scores for wind CF for the three datasets compared to EirGrid data}
	\label{tab:pv_skill_scores}
\end{table}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{verification/solar_verification_contour}
	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}	
	\label{fig:solar_verification_contour}
\end{figure}

Fig.~\ref{fig:bar_number_events_verification_pv} shows results for PV drought events. The figure reveals a lack of significant agreement between the three datasets and the EirGrid data. The main challenge in validating PV stems from the recent installation of a large proportion of the PV installed capacity in Ireland. PV Generation data suffers from large uncertainties for the first few months after each PV farm is connected, making it difficult to know what the actual CF for PV was. Given that PV farms installed in 2023 represent more than 60\% of the total capacity, this significantly impacts our ability to perform vigorous verification for PV.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification/bar_number_events_verification_pv}
	\caption{Average number of PV drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
	\label{fig:bar_number_events_verification_pv}
\end{figure}

\newpage
\subsection{Analysis}
\label{sec:analysis}

In this section, we evaluate RES drought events under two different scenarios with fixed installed capacities: the current scenario, with 89\% of wind capacity(6.7 GW) and 11\% of PV capacity (0.6 GW), and a projected scenario for 2030, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW). Both scenarios are driven by 45 years of ERA5 data. Using the RES drought identification process described earlier (Section~\ref{sec:res_drought}), we first analyse wind and PV droughts separately before presenting the results for combined (wind + PV) RES droughts.

\subsubsection{Annual Number of RES Droughts}

The first part of our analysis examines the annual number of RES drought events for each of the three datasets. For wind results (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting longer than seven days. For PV results (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations. Comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal than those for wind, regardless of duration range or dataset.

The bottom panels show the combination of wind and PV using two different capacity scenarios. With the current RES installation (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts are similar to those for wind alone, as expected, due to the dominance of installed wind generation. However, in the projected scenario (Fig.~\ref{fig:boxplot_number_events}d), there is a distinct reduction in the number of events across all datasets and durations (26\% for Atlite, 20\% for C3S-E G, 18\% for C3S-E N).

The median, first, and third quartiles for the Atlite dataset are consistently greater than or equal to those of the other two datasets, regardless of the duration range or type of renewable energy considered.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{number_events/boxplot_number_events}
	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events. The boxes display the first and third quartiles and the median is marked by a black line. The whiskers indicate the 5th and 95th percentiles}
	\label{fig:boxplot_number_events}	
\end{figure}

\newpage
\subsubsection{Return Periods of RES Drought Duration}

The RES drought events identified over the 45 year period were used to calculate the return periods for different RES drought durations.

For wind (Fig.~\ref{fig:return_periods}a), the duration of RES droughts increases in a log-linear fashion across the three datasets.

For PV (Fig.~\ref{fig:return_periods}b), the duration of RES droughts follows a log-linear increase for events lasting less than 16 days. Beyond this threshold, there is a sharp rise in RES drought duration for events up to a one-year return period. This sudden jump reflects the effect of winter on PV generation in Ireland, as PV generation does not exceed the CF threshold for a large part of winter. The jump for the Atlite dataset is not as large as that for the two C3S-E datasets, due to differences between the datasets in PV generation around this critical 0.1 CF threshold during winter. After this jump, the increase in drought duration continues linearly relative to the log of time.

Similar to Fig.~\ref{fig:boxplot_number_events}c, the current installed capacity combination (Fig.~\ref{fig:return_periods}c), shows comparable results as the wind panel (Fig.~\ref{fig:return_periods}a). In the projected scenario (Fig.~\ref{fig:return_periods}d) the return periods for RES droughts increase across all durations. For example, the return period for a five-day event increases from around half a year to 21 months in the Atlite dataset, and from around a year and a half to almost three years for the two C3S-E datasets.

Across Fig.~\ref{fig:return_periods}a,~c,~d, the return periods for the Atlite dataset are consistently higher than those for the two C3S-E datasets. For example, in Fig.~\ref{fig:return_periods}c, an one-year event lasts for six months according to Atlite, but only five according to the two C3S-E datasets. Additionally, in all four panels, the results for the two C3S-E datasets are closer to each other than they are to Atlite.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.9\textwidth]{return_periods/return_periods}
	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it}
	\label{fig:return_periods}
\end{figure}

\newpage
\subsubsection{Seasonal Distribution of RES Droughts}

The seasonality of RES droughts is analysed by comparing the percentage of hours in a month identified as a RES drought. For wind (Fig.~\ref{fig:res_droughts_seasonality}a), the drought percentages are higher during the summer than in winter. For Atlite, during summer (JJA), there are on average 24\% of hours which are considered as RES drought, and only 4\% during winter (DJF). In contrast the results for PV (Fig.~\ref{fig:res_droughts_seasonality}b)  show a higher percentage in winter, with RES droughts occurring  over 60\% of the time regardless of the dataset. The Atlite results show a higher percentage of RES drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.

Similar to previous results, the current installed capacity wind-PV combination (Fig.~\ref{fig:res_droughts_seasonality}c), shows comparable results to the wind panel (Fig.~\ref{fig:res_droughts_seasonality}a). There is a reduction in RES drought frequency for all three datasets when looking at the projected scenario (Fig.~\ref{fig:res_droughts_seasonality}d). The reduction is greatest during the summer months, which reduces the overall seasonality. Annual reductions in the median values are from 14 to 9\% for Atlite, from 8 to 6\% for C3S-E~G, and 10 to 7\% for C3S-E~N.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{seasonality/percentage_time_events}
	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents the month of the year, and the y-axis indicates the percentage of hours. Lines correspond to the median values and the area between the first and third quartiles is shaded.}
	\label{fig:res_droughts_seasonality}
\end{figure}

\newpage
\subsubsection{Change of Threshold}

In this study, we have considered a threshold of 0.1 in CF to identify a RES drought. In order to check how the results may dependent on this threshold, we have calculated the annual number of events across a range of threshold CF values (Fig~\ref{fig:number_days_threshold}).

The results for wind (Fig.~\ref{fig:number_days_threshold}a) show an increase of number events when the threshold increases. For PV (Fig.~\ref{fig:number_days_threshold}b), the number of events first increases and then start decreasing, as smaller events are continuously merging with each other to produce a smaller number of loner events. For the current (\ref{fig:number_days_threshold}c) and projected (\ref{fig:number_days_threshold}d) installed capacity scenario, the annual number of events increase with the threshold used.

Overall, the four panels show that regardless of the threshold used, the number of events identified by Atlite is consistently higher than those identified by the two C3S-E datasets.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{multiple_th/number_days_threshold}
	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment. The y-axis indicates the annual number of events. Each line correspond to the median value and the first and third quartiles are shown in shaded area}
	\label{fig:number_days_threshold}
\end{figure}

\newpage
\section{Discussion and Conclusions}
\label{sec:DiscConc}

\begin{itemize}
	\item Remind differences between three datasets
	\begin{itemize}
		\item Atlite and C3S-E G: incorporate locations of RES farms
		\item Atlite models better follows EirGrid data (for Wind)
	\end{itemize}
	\item Point Arguments with examples from the figures
	\begin{itemize}
		\item The two C3S-E datasets show similar results
		\item Atlite stands out of these two datasets
	\end{itemize}
	\item Conclusion: Location of RES farms has less impact than the accuracy of modelling the CF!
	\\
	\item Impact of the 2030 scenario (Anti correlation between wind and PV) | Seasonality
	\\
	\item Risk for policymakers. The more accurate Atlite model produces more RES drought events and longer RES drought durations, an important consideration for risk-adverse decision making.
	\\
	\item Where to go next ?
	\begin{itemize}
		\item Looking at future climate data (Combine Climate and Energy scenarios)
		\item Studying broader region (Europe), which is more relevant as everything will be interconnected but more difficult as we will need to repeat verification for all EU countries
	\end{itemize}
\end{itemize}

\begin{comment}
This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts. One of the most significant differences is how each dataset incorporates the specific locations of RES farms. Both Atlite and C3S-E G take into account the exact locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation. However, our analysis shows that the Atlite dataset generally performs better, particularly in its wind generation estimates, where it closely follows the EirGrid data. This suggests that while the inclusion of RES farm locations is important, the accuracy of the CF modeling used by Atlite may be the determining factor in its superior performance.

When analysing the results, we found that the two C3S-E datasets (C3S-E G and C3S-E N) consistently show similar outcomes across various figures (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods},~\ref{fig:res_droughts_seasonality}), suggesting that their differences in methodology have minimal impact on the results. In contrast, Atlite stands out, in its ability to model wind generation, as evidenced by the more accurate correlation with EirGrid data seen in the verification process (Section \ref{sec:wind_verification}). This distinction is further underscored in the figures, where Atlite often reports higher or more accurate return periods and a greater number of RES droughts (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods}), particularly in scenarios with increased PV capacity.

One of the most critical conclusions of this study is that the specific locations of RES farms, while important, have less impact on the accuracy of RES drought modeling than the precision of the CF models themselves. Atlite's superior performance suggests that refining these models could lead to better forecasting and planning for RES droughts, which is crucial for energy security.

Looking ahead to the projected scenario, our analysis suggests a significant shift in the dynamics of RES droughts due to the anti-correlation between wind and PV generation. This scenario reveals the importance of seasonality in RES droughts (Fig.~\ref{fig:res_droughts_seasonality}), where a diversified energy portfolio that includes both wind and PV could decrease the frequency and duration of RES droughts.

For policymakers, the findings of this study highlight the risks of inadequate modeling and planning. Accurate RES drought forecasting is essential for ensuring energy security, especially as Ireland moves toward an ambitious renewable energy target.

In terms of future research, there are two main topics to explore. First, integrating future climate data with energy scenarios to better understand how climate change might affect RES droughts. Second, expanding the scope of the study to include Europe, would provide a more comprehensive understanding of RES drought in an interconnected energy grid. However, this expansion would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.

\end{comment}

\newpage
\section{Acknowledgment}
Using the Climate Data Store. Thank EirGrid for the data and add a reference to the funding project or projects

\section{Funding information}

\section{Author contribution}

\bibliography{litrev}

\newpage
\section*{Appendix}

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.7\textwidth]{data_methods/farm_location_wind_pv}
	\caption{Wind and PV Farms Location in Ireland. The capacity of a farm is illustrated by the size of the dot, the year of installation is represented by its color}	
	\label{fig:farm_location_wind_pv}
\end{figure}

Figure

\begin{table}[!ht]
	\centering
	\begin{tabular}{l|cc|l|}
		\cline{2-4}
		& \textbf{Wind} & \textbf{PV} & \textbf{Total} \\ \hline
		\multicolumn{1}{|l|}{\textbf{Republic of Ireland}} & 4,730         & 445         & 5,175          \\
		\multicolumn{1}{|l|}{\textbf{Northern Ireland}}    & 1,364         & 183         & 1,547          \\ \hline
		\multicolumn{1}{|l|}{\textbf{Total}} & \multicolumn{1}{l}{6,094} & \multicolumn{1}{l|}{628} & \textbf{6,712} \\ \hline
	\end{tabular}
	\caption{Installed Capacities of wind and PV farms (MW) as of January 2024! To be modified !}
	\label{tab:installed_capacities}
\end{table}

Figure

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.9\textwidth]{data_methods/pv_wind_capacity_time}
	\caption{Installed capacity in Ireland}
	\label{fig:pv_wind_capacity_time}
\end{figure}

\end{document}

### 2024-11-13 Article/main.tex
Update Article

@@
 \section{Introduction}
+\label{sec:Intro}
 
@@
 
-\begin{comment}
-The C3S-E national dataset assumes RES farms exist at every grid point, whereas for the C3S-E Gridded and Atlite datasets, the location of each RES farm is applied.
-\textcolor{magenta}{Do you think you could add somehow that the C3S-E datasets assume specific wind and PV conversion models?
-Perhaps this should not be in the introduction, but since farms are mentioned I would include the model as well.}
-\end{comment}
-
 This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
 The first represents the current energy mix, predominantly driven by wind, while the second is based on projected energy scenarios, featuring a more balanced share between wind and PV.
-A consistent methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.
+A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.
 
-\begin{comment}
-	 \textcolor{magenta}{Did we say we are avoiding temporal references in scenarios? I get that ignoring what they represent is not the way to go but the line is very fine I would say}
-\end{comment}
-
-The datasets used in this study are detailed in section~\ref{sec:data}, which describes their sources, characteristics, and relevance for evaluating RES droughts.
-Section~\ref{sec:methods} outlines the energy models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:analysis}, an in-depth analysis of RES drought occurrences is presented for the two energy scenarios, highlighting key differences across the datasets and the effects of varying energy capacities.
-Finally, section~\ref{sec:DiscConc} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
+The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their sources, characteristics, and relevance for evaluating RES droughts.
+Section~\ref{sec:Methods} outlines the energy models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
+In section~\ref{sec:Results}, the models are first verified against actual energy data to assess their accuracy, followed by an analysis of RES drought occurrences across the two energy scenarios, highlighting key differences across the datasets and the effects of varying energy capacities.
+Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
 
 \section{Data}
-\label{sec:data}
+\label{sec:Data}
 
-To construct and validate the models, this study uses publicly available datasets from EirGrid, the Republic of Ireland's transmission system operator, and the ERA5 reanalysis dataset.
-These datasets provide necessary inputs for estimating the CF for wind and PV energy generation.
+This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
+The primary data sources include EirGrid, the transmission system operator for the Republic of Ireland, and the ERA5 reanalysis dataset.
 
-\subsection{EirGrid}
+\subsection{Wind and PV Generation \& Capacity}
 \label{sec:eirgrid}
 
-EirGrid provides detailed datasets for all wind and PV farms in the Republic of Ireland and Northern Ireland (collectively referred to as Ireland) from 1990 to the present~\cite{eirgrid}.
-These datasets include information on each farm’s installed capacity, name, and connection date.
-To improve these datasets, the GPS coordinates for each farm were manually determined using online search.
+EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
+These datasets include critical information such as each farm’s installed capacity, name, and connection date.
+To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
 
-EirGrid also provides a wide range of energy system variables for wind (2014-present) and PV (2018-present) energy generation.
-For our study, we use generation which refers to the actual amount of energy produced by the wind or PV farms.
-This is then combined with the installed capacity to compute the CF.
-\textcolor{cyan}{Change}
+The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
+For this study, availability data, which indicates the amount of energy being delivered to the grid, is used.
+This value is combined with installed capacity to compute the CF, the main metric for assessing renewable energy output.
 
-\begin{comment}
-boris: we use availability which indicates the potential energy output if the farms were operating at full capacity 
-\end{comment}
+Detailed data analysis, supported by access to additional data from EirGrid, revealed that the installed capacity alone does not accurately reflect the true capacity used by the grid.
+Instead, the Maximum Export Capacity (MEC) values from the spreadsheet are used as a basis.
+To adjust for actual grid conditions, these MEC values are further multiplied by a factor of 1.4, which provides a more realistic representation of the PV generation potential.
 
@@
 
-The Copernicus Climate Change Service (\textcolor{cyan}{I have defined C3S-E before, should I use it here ?} \textcolor{magenta}{I am not against introducing some acronyms twice, same as for some references being in the introduction and data, but I would say there is no right or wrong way necessarily} )has developed a renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
+C3S has developed a renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
 This dataset provides hourly CF for wind and PV energy from 1979 to the present.
 The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°.
-Data are also available for download at two aggregated scales: regional (NUTS 2) and national.
+The time series are also available for download at two aggregated scales: regional (NUTS 2) and national.
 
@@
 PV generation is calculated multiple times using the same model with different azimuth and tilt angles.
-The results are aggregated based on a statistical distribution of the module angles~\cite{saintdrenan2018solar}.
-\textcolor{cyan}{I added a sentence to explain the reason of the Vestas choice} \textcolor{magenta}{So, basically what they are saying is that they are looking at a realistic turbine for the future, makes sense.
-Not necessarily linked to the article, but I wonder how realistic this turbine choice is for other countries.
-We know it is not for Ireland, but I am curious about the rest of Europe}
+The results are aggregated based on a statistical distribution of the module angles found in Germany and replicated in European countries~\cite{saintdrenan2018solar}.
 
 \section{Methods}
-\label{sec:methods}
+\label{sec:Methods}
 
@@
 
-\subsection{C3S-E N}
-\label{sec:C3SEN}
+\subsection{C3S-Energy National}
+\label{sec:c3se_n}
 
@@
 
-\subsection{C3S-E G}
-\label{sec:C3SEG}
+\subsection{C3S-E Gridded}
+\label{sec:c3se_g}
 
@@
 We then average the CF for each farm using the installed capacity as weights.
-This process results in a CF time series for each renewable energy source (wind and PV), which is representative of the location of the RES farms in Ireland.
+This process results in a CF time series for each renewable energy source (wind energy and PV energy), which is representative of the location of the RES farms in Ireland.
 
@@
 We provide Atlite with a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake} at a hub height of 100 metres, which it uses when converting wind speed to CF.
-
 For PV power estimation, Atlite uses four irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) as well as the air temperature ($t2m$).
@@
 
+The selection of a specific wind turbine and PV panel characteristics, are further discussed and explained in section \ref{sec:verification}.
+
 \subsection{Energy Scenarios}
@@
 
-This study investigates wind and PV CF using three models, resulting in six CF time series with the installed capacity at the end of 2023.
-Additionally, we analyse the combined CF of wind and PV for each model by averaging the two sources, weighted by their capacities (6.7 GW for wind, 0.6 GW for PV).
-This approach gives us a total of nine CF time series for the 2023 scenario.
+This study investigates wind and PV capacity factors (CF) using three different models, resulting in six individual CF time series based on the installed capacities of 2023.
+In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
+This results in a total of nine CF time series for the existing scenario, which represents the current distribution of RES installations in Ireland.
 
-Given the relatively low PV capacity in Ireland in 2023 and the projected increase by 2030 (expected to increase from 0.6 GW to 8 GW), we also explore a future scenario with higher PV capacity.
-We generate additional CF time series reflecting the projected PV capacity in 2030 for both the Atlite and C3S-E G models, based on a roadmap published by EirGrid~\cite{eirgrid2023future}.
+Given that PV capacity in Ireland remains relatively low in 2023, and in order to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
+These figures are based on the target energy capacity outlined in the roadmap published by EirGrid~\cite{eirgrid2023future}.
+In this scenario, new PV time series are generated for both the Atlite and C3S-E G models, while the wind CF time series remain the same, as significant changes in wind farm distribution are not expected.
+From there, the results of these new time series are combined using updated weights which reflect a more balanced scenario (9 GW for wind and 8 GW for PV).
 
-The wind capacity is projected to increase to 9 GW in 2030, however as the spatial distribution is not expected to significantly change we use the same CF time series.
-The combination of wind and PV CF is calculated once again using the new PV time series, and by averaging with the future projected capacities of RES farms.
-In total, we evaluate fourteen CF time series across different scenarios and models, as summarised in Table~\ref{tab:datasets}.
+In total, twelve CF time series are analysed in this study, six for individual wind and PV CF (three models for each source) in the existing scenario, and an additional six time series that include the combined CF for both existing and balanced scenarios across the different models.
 
-\begin{table}[h!]
-	\centering
-	\begin{tabular}{l|ccc|ccc|}
-		\cline{2-7}
-		& \multicolumn{3}{c|}{\textbf{Current}}                                   & \multicolumn{3}{c|}{\textbf{Projected}}                                   \\ \cline{2-7} 
-	    & \multicolumn{1}{c|}{Atlite} & \multicolumn{1}{c|}{C3S-E G} & C3S-E N & \multicolumn{1}{c|}{Atlite} & \multicolumn{1}{c|}{C3S-E G} & C3S-E N \\ \hline
-		\multicolumn{1}{|l|}{\textbf{PV}}      & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       &         \\
-		\multicolumn{1}{|l|}{\textbf{Wind}}    & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{}       & \multicolumn{1}{c|}{}        &         \\
-		\multicolumn{1}{|l|}{\textbf{Combine}} & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       & \multicolumn{1}{c|}{x}      & \multicolumn{1}{c|}{x}       & x       \\ \hline
-	\end{tabular}
-	\caption{Summary of the time series compared in this study}
-	\label{tab:datasets}
-\end{table}
+This approach allows for a comparative analysis between a wind-dominated system (the existing scenario) and a more evenly distributed system (the balanced scenario), assessing how the balance of RES capacity affects the occurrence of RES droughts.
+The results for these two scenarios are analysed and discussed in the following sections.
 
-\subsection{RES drought definition}
+\subsection{RES Drought Definition}
 \label{sec:res_drought}
 
-In this study, we define a RES drought event as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-Using the moving average captures fewer, but longer-lasting events than using the regular time series.
+In this study, a RES drought event is defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period longer than 24 hours.
+The choice of this threshold is somewhat arbitrary but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought} .
+By using a 24-hour moving average, fewer but longer-lasting events are captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
+Although other methods exist for identifying RES droughts, a fixed threshold has been chosen in this study to enable consistent inter-comparison across the three datasets.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{data_methods/find_res_droughts}
-	\caption{Time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021.
+	\includegraphics[width=\textwidth]{droughts_methodology}
+	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021.
 CF threshold (dashed black).
-The grey bar shows the period identified as a wind drought.}
+The grey bar shows the period identified as a wind drought under our definition}
 	\label{fig:find_res_droughts}
@@
 Fig.~\ref{fig:find_res_droughts} illustrates the moving average identification method.
-When using the moving average method, this CF time series results in one event being identified, which lasts 161 hours.
-In contrast, using the regular CF series would have resulted in three separate events being identified, with durations of 36, 35, and 40 hours.
+When applying the moving average, this CF time series identifies a single event lasting 161 hours.
+In contrast, using the raw CF series would have resulted in three separate events, with durations of 36, 35, and 40 hours.
+
+The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period.
+This means that a single hour above the threshold does not "break" a drought event if it is surrounded by prolonged low-generation hours.
+As a result, fewer but longer-lasting drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
 
 \section{Results}
-\label{sec:results}
+\label{sec:Results}
 
@@
 
-\subsubsection{Wind}
+\subsubsection{Wind Energy}
 \label{sec:wind_verification}
@@
 The smoothing is done by using a Gaussian filter with different standard deviations that depend on the wind speed.
-We generated a wind CF time series for Ireland for each of these wind turbine power curve options.
-The performance of each wind turbine power curve was calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
+A wind CF time series for Ireland for all wind turbine power curve options and all smoothings is generated.
+The performance of each wind turbine power curve is then, calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
 The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
-Based on these metrics, the most representative power curve for Ireland was the Enercon E112.4500 power curve with the $0.3w$ filter.
+Based on these metrics, the most representative power curve for Ireland was the Enercon E112.4500 power curve with the $0.3w$  smoothing filter.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification/power_curve}
-	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
+	\includegraphics[width=\textwidth]{power_curve}
+	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
 	\label{fig:power_curve}
@@
 
-The smoothing of the wind turbine power curve represents the wake effect, which is important when modelling wind energy on larger scales.
+The smoothing of the wind turbine power curve represents losses due to the wake effect, which is important when modelling wind energy on larger scales.
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind data from EirGrid.
+
+\begin{figure}[!ht]
+	\centering
+	\includegraphics[width=\textwidth]{verification_wind_contour}
+	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
+	\label{fig:wind_verification_contour}
+\end{figure}
+
+The effect of the differences in the power curves are also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data.
+In contrast, the two C3S-E datasets tend to overestimate the EirGrid CF.
+This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.
 
@@
 
-The effect of the differences in the power curves are also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data.
-In contrast, the two C3S-E datasets tend to overestimate the EirGrid CF.
-This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.
+Fig.~\ref{fig:bar_number_events_verification_wind} presents results for the average annual number of wind drought events during the 2014 to 2023 validation period.
+The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
+This pattern is particularly evident for shorter duration events, which are the most frequent.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification/wind_verification_contour}
-	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
-	\label{fig:wind_verification_contour}
-\end{figure}
-
-Fig.~\ref{fig:bar_number_events_verification_wind} presents results for the average annual number of wind drought events during the 2014 to 2023 validation period.
-The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{verification/bar_number_events_verification_wind}
-	\caption{Average annual number of wind drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple) / Change plot}
+	\includegraphics[width=\textwidth]{verification_wind_bar}
+	\caption{Average annual number of wind drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
 	\label{fig:bar_number_events_verification_wind}
@@
 
-\subsubsection{PV}
+\newpage
+\subsubsection{PV Energy}
 \label{sec:pv_verification}
@@
 The C3S-E datasets calculate PV generation multiple times using the same model with varying azimuth and tilt angles.
-The results are then aggregated based on a statistical distribution of module angles.
-Similarly to the wind CF verification, the Atlite model allows the user to select from one of three different PV panel characteristics.
+The results are then aggregated based on a statistical distribution of module angles of Germany and then adapted to other European countries.
+Similarly to the wind CF verification, the Atlite model allows the user to select a given PV panel characteristics.
+In this study we have chosen to look at the three PV panels offered by in the Atlite model.
 Following the same methodology as in the previous section, we compare the three available models using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, we selected the Breyer PV panel model \cite{beyer2004pv}, specifically using the Kaneka Hybrid panel option.
+Based on the best-performing metrics, we selected the Breyer PV panel model \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
 
-Figure \ref{fig:solar_verification_contour} shows that the three datasets have similar correlation when compared to the EirGrid data, which is further supported by the results shown in Table~\ref{tab:pv_skill_scores}.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{verification_pv_contour}
+	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}	
+	\label{fig:solar_verification_contour}
+\end{figure}
+
+Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency of overestimating the CF compared to EirGrid.
+The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
+Atlite shows a slightly lower correlation and a higher RMSE, indicating larger deviations.
+All models show a slight positive bias, with Atlite exhibiting the most overestimation.
+These results reflect the broader challenges in accurately modeling PV, where variability is hard to capture.
+Nonetheless, C3S-E G’s lower RMSE suggests it may be more reliable for PV CF modeling in this context.
 
@@
 	& \textbf{Atlite} & \textbf{C3S-E G} & \textbf{C3S-E N} \\ \hline
-	\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.913            \\ \hline
-	\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.070           & 0.065            & 0.100            \\ \hline
-	\multicolumn{1}{|l|}{\textbf{MBE}}   & 0.002           & -0.012           & -0.013           \\ \hline
+	\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.931            \\ \hline
+	\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.119           & 0.090            & 0.113            \\ \hline
+	\multicolumn{1}{|l|}{\textbf{MBE}}   & 0.046           & 0.027           & 0.021           \\ \hline
 	\end{tabular}
@@
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\textwidth]{verification/solar_verification_contour}
-	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}	
-	\label{fig:solar_verification_contour}
-\end{figure}
-
-Fig.~\ref{fig:bar_number_events_verification_pv} shows results for PV drought events.
-The figure reveals a lack of significant agreement between the three datasets and the EirGrid data.
-The main challenge in validating PV stems from the recent installation of a large proportion of the PV installed capacity in Ireland.
-PV Generation data suffers from large uncertainties for the first few months after each PV farm is connected, making it difficult to know what the actual CF for PV was.
-Given that PV farms installed in 2023 represent more than 60\% of the total capacity, this significantly impacts our ability to perform vigorous verification for PV.
+Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of PV drought events in 2023 across different duration ranges.
+The figure reveals partial agreement between the three datasets and the EirGrid data, with consistent results observed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
+However, discrepancies appear in the other ranges, where the models diverge from the EirGrid data.
+The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity.
+Due to uncertainties in PV generation data during the first few months after each farm is connected, it is difficult to accurately determine the effective capacity.
+With over 60\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
+While the results are reasonably aligned in some ranges, the inherent challenges in PV data highlight the limitations of this verification.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification/bar_number_events_verification_pv}
-	\caption{Average number of PV drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
+	\includegraphics[width=\textwidth]{verification_pv_bar}
+	\caption{Annual number of PV drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple) in 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 \subsection{Analysis}
-\label{sec:analysis}
+\label{sec:Analysis}
 
-In this section, we evaluate RES drought events under two different scenarios with fixed installed capacities: the current scenario, with 89\% of wind capacity(6.7 GW) and 11\% of PV capacity (0.6 GW), and a projected scenario for 2030, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
+In this section, we evaluate RES drought events under two different scenarios with fixed installed capacities: the existing scenario, with 89\% of wind capacity (6.7 GW) and 11\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
 Both scenarios are driven by 45 years of ERA5 data.
@@
 
-The first part of our analysis examines the annual number of RES drought events for each of the three datasets.
-For wind results (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting longer than seven days.
-For PV results (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
-Comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal than those for wind, regardless of duration range or dataset.
+The first part of the analysis examines the annual number of RES drought events across the three datasets.
 
-The bottom panels show the combination of wind and PV using two different capacity scenarios.
-With the current RES installation (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts are similar to those for wind alone, as expected, due to the dominance of installed wind generation.
-However, in the projected scenario (Fig.~\ref{fig:boxplot_number_events}d), there is a distinct reduction in the number of events across all datasets and durations (26\% for Atlite, 20\% for C3S-E G, 18\% for C3S-E N).
+For wind (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting more than seven days.
+For PV (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
+This increase is due to extended low-generation periods occurring from November to March, depending on the dataset.
+When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
+This is due to PV's typically lower CF compared to wind, especially in a region like Ireland where solar potential is limited.
+PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of RES droughts in PV compared to wind.
+
+The bottom panels show the combination of wind and PV under the  two capacity scenarios.
+In the existing RES installation scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
+In contrast, the balanced scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
+This reduction is attributed to the anti-correlation between wind and PV production throughout the year, a phenomenon well-documented in the literature \cite{kaspar2019drought} and discussed in further detail in the later analysis.
 
 The median, first, and third quartiles for the Atlite dataset are consistently greater than or equal to those of the other two datasets, regardless of the duration range or type of renewable energy considered.
+This difference arises from the wind turbine power curve model used in the C3S-E datasets, which tends to overestimate the wind CF (Fig.~\ref{fig:wind_verification_contour}).
+As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{number_events/boxplot_number_events}
+	\includegraphics[width=\textwidth]{droughts_number_events}
 	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
@@
 
-The RES drought events identified over the 45 year period were used to calculate the return periods for different RES drought durations.
+The RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
+A return period refers to the estimated average interval between occurrences of events with a specified duration or intensity.
+Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
+This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the potential impact of RES droughts on energy reliability.
 
 For wind (Fig.~\ref{fig:return_periods}a), the duration of RES droughts increases in a log-linear fashion across the three datasets.
+The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer RES droughts becoming exponentially less likely as duration increases.
 
-For PV (Fig.~\ref{fig:return_periods}b), the duration of RES droughts follows a log-linear increase for events lasting less than 16 days.
-Beyond this threshold, there is a sharp rise in RES drought duration for events up to a one-year return period.
-This sudden jump reflects the effect of winter on PV generation in Ireland, as PV generation does not exceed the CF threshold for a large part of winter.
-The jump for the Atlite dataset is not as large as that for the two C3S-E datasets, due to differences between the datasets in PV generation around this critical 0.1 CF threshold during winter.
-After this jump, the increase in drought duration continues linearly relative to the log of time.
+For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves significantly different than the two C3S-E datasets.
+The Atlite results follow the ones for wind with a log-linear increase, but reach higher values in general with the longest event lasting forty days.
+For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days.
+Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
+This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months.
+The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold.
+Atlite appears to hover slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
+In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
+This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions.
 
-Similar to Fig.~\ref{fig:boxplot_number_events}c, the current installed capacity combination (Fig.~\ref{fig:return_periods}c), shows comparable results as the wind panel (Fig.~\ref{fig:return_periods}a).
-In the projected scenario (Fig.~\ref{fig:return_periods}d) the return periods for RES droughts increase across all durations.
-For example, the return period for a five-day event increases from around half a year to 21 months in the Atlite dataset, and from around a year and a half to almost three years for the two C3S-E datasets.
+For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the current energy mix.
+In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
+For example, the return period for a five-day drought event extends from roughly six months to four years in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
+The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
+This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics.
+Notably, there is a greater difference in how each model represents wind generation than PV, meaning the balanced scenario reduces these wind-related discrepancies, bringing the datasets closer together.
 
-Across Fig.~\ref{fig:return_periods}a,~c,~d, the return periods for the Atlite dataset are consistently higher than those for the two C3S-E datasets.
-For example, in Fig.~\ref{fig:return_periods}c, an one-year event lasts for six months according to Atlite, but only five according to the two C3S-E datasets.
-Additionally, in all four panels, the results for the two C3S-E datasets are closer to each other than they are to Atlite.
+Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
+For instance, in the current scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
+This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influences drought duration estimates.
+Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
 
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{return_periods/return_periods}
+	\includegraphics[width=0.9\textwidth]{droughts_return_periods}
 	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
-The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it}
+The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
+The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset.}
 	\label{fig:return_periods}
@@
 
-The seasonality of RES droughts is analysed by comparing the percentage of hours in a month identified as a RES drought.
-For wind (Fig.~\ref{fig:res_droughts_seasonality}a), the drought percentages are higher during the summer than in winter.
-For Atlite, during summer (JJA), there are on average 24\% of hours which are considered as RES drought, and only 4\% during winter (DJF).
+The seasonality of RES droughts is analysed by comparing the percentage of hours in each month classified as part of a RES drought.
+
+For wind (Fig.~\ref{fig:res_droughts_seasonality}a), RES drought percentages are higher in summer than in winter.
+In the Atlite dataset, for instance, an average of 24\% of hours in summer (JJA) are identified as RES droughts, compared to only 4\% in winter (DJF).
+This seasonal variation is influenced by the wind power curve model used to estimate CF, where the shape of the curve in lower wind speed regions (3-10 m/s) leads to significant differences in CF under low wind conditions.
 In contrast the results for PV (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
@@
 
-Similar to previous results, the current installed capacity wind-PV combination (Fig.~\ref{fig:res_droughts_seasonality}c), shows comparable results to the wind panel (Fig.~\ref{fig:res_droughts_seasonality}a).
-There is a reduction in RES drought frequency for all three datasets when looking at the projected scenario (Fig.~\ref{fig:res_droughts_seasonality}d).
-The reduction is greatest during the summer months, which reduces the overall seasonality.
-Annual reductions in the median values are from 14 to 9\% for Atlite, from 8 to 6\% for C3S-E~G, and 10 to 7\% for C3S-E~N.
+Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the wind panel (Fig.\ref{fig:res_droughts_seasonality}a).
+However, in this panel, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
+This reduction can be explained by the contribution of PV generation during the summer months in the existing scenario, even though it constitutes only 11\% of total capacity.
+Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
+In the projected scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
+The balanced mix of wind and PV in this scenario reduces the seasonal signal, slightly increasing the percentage of RES drought hours in winter months while significantly decreasing it in summer.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{seasonality/percentage_time_events}
-	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\includegraphics[width=\textwidth]{droughts_seasonality}
+	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
@@
 
-In this study, we have considered a threshold of 0.1 in CF to identify a RES drought.
-In order to check how the results may dependent on this threshold, we have calculated the annual number of events across a range of threshold CF values (Fig~\ref{fig:number_days_threshold}).
+In this study, a threshold CF value of 0.1 has been used to identify RES droughts.
+To assess how sensitive the results are to this threshold, the annual number of events was calculated across a range of CF threshold values (Fig.~\ref{fig:number_days_threshold}).
 
-The results for wind (Fig.~\ref{fig:number_days_threshold}a) show an increase of number events when the threshold increases.
-For PV (Fig.~\ref{fig:number_days_threshold}b), the number of events first increases and then start decreasing, as smaller events are continuously merging with each other to produce a smaller number of loner events.
-For the current (\ref{fig:number_days_threshold}c) and projected (\ref{fig:number_days_threshold}d) installed capacity scenario, the annual number of events increase with the threshold used.
+Across all four panels, the results reveal a general trend: as the threshold increases, so does the number of identified events.
+This increase, however, is not uniform across datasets and energy sources, but the overall pattern remains consistent.
+When the threshold increases, smaller events merge into larger, less frequent events.
+Depending on the dataset, the number of events reaches a peak around or after a threshold of 0.2, and then gradually decreases until only a single event encompasses the entire time series.
 
-Overall, the four panels show that regardless of the threshold used, the number of events identified by Atlite is consistently higher than those identified by the two C3S-E datasets.
+Overall, the four panels indicate that regardless of the threshold applied, the Atlite dataset consistently identifies more events than the two C3S-E datasets, except for PV when the threshold is below 0.08.
+This finding suggests that Atlite, likely due to its more accurate representation of wind patterns in Ireland, captures more RES drought events that might be missed with different modeling assumptions in the C3S-E datasets.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{multiple_th/number_days_threshold}
+	\includegraphics[width=\textwidth]{drouhgts_varying_threshold}
 	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
@@
 \section{Discussion and Conclusions}
-\label{sec:DiscConc}
+\label{sec:Conclusion}
 
@@
 
-\begin{comment}
 This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts.
@@
 
-Looking ahead to the projected scenario, our analysis suggests a significant shift in the dynamics of RES droughts due to the anti-correlation between wind and PV generation.
+Looking to the balanced scenario, our analysis suggests a significant shift in the dynamics of RES droughts due to the anti-correlation between wind and PV generation.
 This scenario reveals the importance of seasonality in RES droughts (Fig.~\ref{fig:res_droughts_seasonality}), where a diversified energy portfolio that includes both wind and PV could decrease the frequency and duration of RES droughts.
@@
 However, this expansion would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.
-
-\end{comment}
 
@@
 
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=0.7\textwidth]{data_methods/farm_location_wind_pv}
-	\caption{Wind and PV Farms Location in Ireland. The capacity of a farm is illustrated by the size of the dot, the year of installation is represented by its color}	
-	\label{fig:farm_location_wind_pv}
-\end{figure}
-
-Figure
-
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{l|cc|l|}
-		\cline{2-4}
-		& \textbf{Wind} & \textbf{PV} & \textbf{Total} \\ \hline
-		\multicolumn{1}{|l|}{\textbf{Republic of Ireland}} & 4,730         & 445         & 5,175          \\
-		\multicolumn{1}{|l|}{\textbf{Northern Ireland}}    & 1,364         & 183         & 1,547          \\ \hline
-		\multicolumn{1}{|l|}{\textbf{Total}} & \multicolumn{1}{l}{6,094} & \multicolumn{1}{l|}{628} & \textbf{6,712} \\ \hline
-	\end{tabular}
-	\caption{Installed Capacities of wind and PV farms (MW) as of January 2024! To be modified !}
-	\label{tab:installed_capacities}
-\end{table}
-
-Figure
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=0.9\textwidth]{data_methods/pv_wind_capacity_time}
-	\caption{Installed capacity in Ireland}
-	\label{fig:pv_wind_capacity_time}
-\end{figure}
-
 \end{document}

### 2024-11-14 Article/main.tex
Add comments to discussion

@@
 However, our analysis shows that the Atlite dataset generally performs better, particularly in its wind generation estimates, where it closely follows the EirGrid data.
-This suggests that while the inclusion of RES farm locations is important, the accuracy of the CF modeling used by Atlite may be the determining factor in its superior performance.
+This suggests that while the inclusion of RES farm locations is important, the accuracy of the CF modeling \textcolor{magenta}{I would say power curve model instead of CF modelling} used by Atlite may be the determining factor in its superior performance.\textcolor{magenta}{Superior how?
+Is finding more droughts superior or is it reproducing reality?
+I reckon it's the second, but make it explicit.
+Also, there is no explicit mention of PV, if you mention wind explicitly (which I agree on doing) I would add PV.
+Even though it is replacing that "However" with "Even though this is true for PV, our analysis shows that for wind..."}
 
-When analysing the results, we found that the two C3S-E datasets (C3S-E G and C3S-E N) consistently show similar outcomes across various figures (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods},~\ref{fig:res_droughts_seasonality}), suggesting that their differences in methodology have minimal impact on the results.
-In contrast, Atlite stands out, in its ability to model wind generation, as evidenced by the more accurate correlation with EirGrid data seen in the verification process (Section \ref{sec:wind_verification}).
-This distinction is further underscored in the figures, where Atlite often reports higher or more accurate return periods and a greater number of RES droughts (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods}), particularly in scenarios with increased PV capacity.
+When analysing the results, we found that the two C3S-E datasets (C3S-E G and C3S-E N) consistently show similar outcomes across various figures \textcolor{magenta}{I would not refer to them as figures, but as analysis or something like that.
+But if you do choose the word analysis, do not start the sentence with "When analysing the results..." and instead use something like "The results show..."} (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods},~\ref{fig:res_droughts_seasonality}), suggesting that their differences in methodology have minimal impact on the results.
+In contrast, Atlite stands out in its ability to model wind generation, as evidenced by the more accurate correlation with EirGrid data seen in the verification process (Section \ref{sec:wind_verification}).
+This distinction is further underscored in the figures \textcolor{magenta}{Same as before, avoid referring to the figures.
+We do not discuss the result (that was done in results), we discuss what it implies}, where Atlite often reports higher or more accurate return periods and a greater number of RES droughts (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods}), particularly in scenarios with increased PV capacity.
+\textcolor{magenta}{For me, the point of this paragraph is quite clear but does not really come across at the moment: Atlite is the best one at reproducing reality (as established in the previous paragraph).
+At the same time, a long-term analysis shows more droughts for Atlite than any other model.
+This means that model selection is key to correctly quantify the risk and avoid undersizing the reserve capacity.}
 
-One of the most critical conclusions of this study is that the specific locations of RES farms, while important, have less impact on the accuracy of RES drought modeling than the precision of the CF models themselves.
-Atlite's superior performance suggests that refining these models could lead to better forecasting and planning for RES droughts, which is crucial for energy security.
+One of the most critical conclusions of this study is that the specific locations of RES farms, while important, have a lower impact on the accuracy of RES drought modelling than the precision of the CF models themselves.
+\textcolor{magenta}{Same as before, I would not refer to CF models, but instead wind power curve and PV panel models} Atlite's superior performance suggests that refining these models could lead to better forecasting and planning for RES droughts, which is crucial for energy security.
 
@@
 This scenario reveals the importance of seasonality in RES droughts (Fig.~\ref{fig:res_droughts_seasonality}), where a diversified energy portfolio that includes both wind and PV could decrease the frequency and duration of RES droughts.
-
-For policymakers, the findings of this study highlight the risks of inadequate modeling and planning.
+\textcolor{magenta}{I would merge these two paragraphs into one.
+Here are my thoughts in order: 1.
+We have found how wind and PV together perform better than individually in terms of droughts.
+2.
+A specific mention of seasonality and how it is reduced.
+3.
+Mention how Ireland has a very wind-dependent system and how targets for PV are ambitious in the coming years, almost making wind and PV capacities equal (maybe not the phrasing but you see the point, putting forward the balance again).
+4.
+Link to policymakers in terms of how meeting targets of installed capacity can help improve the energy security thanks to this diversification.
+5.
+Close with the importance of the right model selection for RES drought assessment, especially as more renewable capacity comes into the system.}
+For policymakers, the findings of this study highlight the risks of inadequate modelling and planning.
 Accurate RES drought forecasting is essential for ensuring energy security, especially as Ireland moves toward an ambitious renewable energy target.
@@
 First, integrating future climate data with energy scenarios to better understand how climate change might affect RES droughts.
-Second, expanding the scope of the study to include Europe, would provide a more comprehensive understanding of RES drought in an interconnected energy grid.
+Second, expanding the scope of the study to include Europe would provide a more comprehensive understanding of RES drought in an interconnected energy grid.
 However, this expansion would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.

### 2024-11-14 Article/main.tex
Add the rest of comments to article

@@
 \end{abstract}
+
+\textcolor{magenta}{Some things to keep in mind when reading from Damian's comments:
+	1. The whole current/projected and existing/balanced thing. Pick one and stick to it, including the figures (also fix the a/c level in fig 10)
+	2. See if you are happy with the section titles throughout the document
+	3. Check if when it is EirGrid/SONI it is clear, both in the text and references
+	4. Is the Ireland / ROI / NI distinction clear enough? I did not see the line where we say that we call ROI+NI=Ireland (may have missed it) but I am alright with how it is written now either way}
 
@@
 
-Detailed data analysis, supported by access to additional data from EirGrid, revealed that the installed capacity alone does not accurately reflect the true capacity used by the grid.
-Instead, the Maximum Export Capacity (MEC) values from the spreadsheet are used as a basis.
-To adjust for actual grid conditions, these MEC values are further multiplied by a factor of 1.4, which provides a more realistic representation of the PV generation potential.
+Detailed data analysis, supported by access to additional data from EirGrid, revealed that the PV installed capacity available on the spreadsheets is the Maximum Export Capacity (MEC) and does not accurately reflect the true generating capacity.
+Instead, MEC values from the spreadsheet are used as a basis.
+To adjust for actual grid conditions, installed capacities are assumed to be 1.4 times the MEC values.
+Whenever farm generation is calculated, this scaled capacity is used and any farm output is capped to the MEC.
 
@@
 
-Atlite and C3S-E datasets are driven by the ERA5 reanalysis dataset~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present.
@@
 PV generation is calculated multiple times using the same model with different azimuth and tilt angles.
-The results are aggregated based on a statistical distribution of the module angles found in Germany and replicated in European countries~\cite{saintdrenan2018solar}.
+The results are aggregated based on a statistical distribution of the module angles based on the geographical location~\cite{saintdrenan2018solar}.
 
@@
 From there, the results of these new time series are combined using updated weights which reflect a more balanced scenario (9 GW for wind and 8 GW for PV).
+\textcolor{magenta}{Is the specific distribution of the capacities mentioned somewhere?
+Same locations but scaled for wind, change in locations for PV}
 
@@
 	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021.
-CF threshold (dashed black).
+The black dashed line indicates the CF threshold.
 The grey bar shows the period identified as a wind drought under our definition}
@@
 
-The C3S-E datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve} a).
+The C3S-E datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve}a).
 The Atlite model allows the user to specify the power curve.
@@
 A wind CF time series for Ireland for all wind turbine power curve options and all smoothings is generated.
-The performance of each wind turbine power curve is then, calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
+The performance of each wind turbine power curve is then calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
 The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
@@
 
-The smoothing of the wind turbine power curve represents losses due to the wake effect, which is important when modelling wind energy on larger scales.
-The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind data from EirGrid.
+The smoothing of the wind turbine power curve represents losses due to the wake effect \textcolor{magenta}{yes, but not only.
+we do not want someone to be able to suck us into what the smoothing means rabbithole.
+My proposal: The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines}, which is important when modelling wind energy on larger scales.
+The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
 
@@
 
-The effect of the differences in the power curves are also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data.
+The effect of the differences in the power curves is also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data.
 In contrast, the two C3S-E datasets tend to overestimate the EirGrid CF.
 This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.
+
+\textcolor{magenta}{I was reading Damian's comment about the tables being "a lot of numbers".
+I think the way it is written now is clear but please review it and see if you are happy with it too.
+I though about maybe explaining that RMSE and MBE are better closer to 0 and CC closer to 1, but honestly I think that the reader will know that, it is quite a common method and not a complex selection of skill scores.}
 
@@
 The C3S-E datasets calculate PV generation multiple times using the same model with varying azimuth and tilt angles.
-The results are then aggregated based on a statistical distribution of module angles of Germany and then adapted to other European countries.
+The results are then aggregated based on a statistical distribution linked to each geographical location.
 Similarly to the wind CF verification, the Atlite model allows the user to select a given PV panel characteristics.
@@
 
-Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency of overestimating the CF compared to EirGrid.
+Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency of overestimating the CF compared to EirGrid, especially for high CF values.
 The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
@@
 However, discrepancies appear in the other ranges, where the models diverge from the EirGrid data.
-The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity.
-Due to uncertainties in PV generation data during the first few months after each farm is connected, it is difficult to accurately determine the effective capacity.
+The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, leading to uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
 With over 60\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
@@
 As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite.
+\textcolor{magenta}{Sorry to add this here again, but where are we on the current/projected/balanced thing?
+Just please make sure it is consistent, we can figure it out once Conor has gone over it, it's not crucial to the reading but I just remembered.}
 
@@
 This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions.
+\textcolor{magenta}{Not sure if it is worth a mention, but in case you had not made the connection: this is consistent with atlite having the highest MBE out of the three in the verification}
 
@@
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
+\textcolor{magenta}{This is a good point to mention the seasonality of the demand in Ireland that Damian brought up, although it could also go in the conclusions}
 
@@
 This finding suggests that Atlite, likely due to its more accurate representation of wind patterns in Ireland, captures more RES drought events that might be missed with different modeling assumptions in the C3S-E datasets.
+
+\textcolor{magenta}{So, here the panel names are different as in the other ones, with existing and balanced.
+this is what I meant by consistency.
+Also, note the panel a in the figure is labelled as c.
+that should be changed.
+}
 

### 2024-11-15 Article/main.tex
Update Article

@@
 	1. The whole current/projected and existing/balanced thing. Pick one and stick to it, including the figures (also fix the a/c level in fig 10)
-	2. See if you are happy with the section titles throughout the document
-	3. Check if when it is EirGrid/SONI it is clear, both in the text and references
-	4. Is the Ireland / ROI / NI distinction clear enough? I did not see the line where we say that we call ROI+NI=Ireland (may have missed it) but I am alright with how it is written now either way}
+	2. See if you are happy with the section titles throughout the document}
 
@@
 This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
-The first represents the current energy mix, predominantly driven by wind, while the second is based on projected energy scenarios, featuring a more balanced share between wind and PV.
+The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected energy scenarios, featuring a more balanced share between wind and PV.
 A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.
@@
 This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
-The primary data sources include EirGrid, the transmission system operator for the Republic of Ireland, and the ERA5 reanalysis dataset.
+The primary data sources include EirGrid, the transmission system operator for the ic of Ireland, and the ERA5 reanalysis dataset.
 
@@
 
-For the second dataset, we use the gridded dataset from C3S-E to create a CF time series which is representative of the actual RES farm distribution.
-Using the farms' coordinates, we identify the nearest grid point on the C3S-E grid for each farm.
-We retrieve the CF values from the C3S-E dataset for its corresponding grid point.
-We then average the CF for each farm using the installed capacity as weights.
-This process results in a CF time series for each renewable energy source (wind energy and PV energy), which is representative of the location of the RES farms in Ireland.
+EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
+These datasets include critical information such as each farm’s installed capacity, name, and connection date.
+To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
+For simplicity, this data will be referred to as originating from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
 
@@
 In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
-This results in a total of nine CF time series for the existing scenario, which represents the current distribution of RES installations in Ireland.
+This results in a total of nine CF time series for the existing scenario, which represents the existing distribution of RES installations in Ireland.
 
-Given that PV capacity in Ireland remains relatively low in 2023, and in order to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
-These figures are based on the target energy capacity outlined in the roadmap published by EirGrid~\cite{eirgrid2023future}.
-In this scenario, new PV time series are generated for both the Atlite and C3S-E G models, while the wind CF time series remain the same, as significant changes in wind farm distribution are not expected.
-From there, the results of these new time series are combined using updated weights which reflect a more balanced scenario (9 GW for wind and 8 GW for PV).
-\textcolor{magenta}{Is the specific distribution of the capacities mentioned somewhere?
-Same locations but scaled for wind, change in locations for PV}
+Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario.
+In this scenario, the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW, based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
+For PV, new time series are generated for both the Atlite and C3S-E G models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+For wind, the CF time series remain unchanged, as significant shifts in wind farm distribution are not expected.
+The PV and wind time series are then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{droughts_methodology}
+	\includegraphics{droughts_methodology}
 	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021.
@@
 	\centering
-	\includegraphics[width=\textwidth]{power_curve}
+	\includegraphics[width=\textwidth]{verification_power_curve}
 	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
@@
 
-The smoothing of the wind turbine power curve represents losses due to the wake effect \textcolor{magenta}{yes, but not only.
-we do not want someone to be able to suck us into what the smoothing means rabbithole.
-My proposal: The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines}, which is important when modelling wind energy on larger scales.
+The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines, which is important when modelling wind energy on larger scales.
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
@@
 This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.
-
-\textcolor{magenta}{I was reading Damian's comment about the tables being "a lot of numbers".
-I think the way it is written now is clear but please review it and see if you are happy with it too.
-I though about maybe explaining that RMSE and MBE are better closer to 0 and CC closer to 1, but honestly I think that the reader will know that, it is quite a common method and not a complex selection of skill scores.}
 
@@
 As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite.
-\textcolor{magenta}{Sorry to add this here again, but where are we on the current/projected/balanced thing?
-Just please make sure it is consistent, we can figure it out once Conor has gone over it, it's not crucial to the reading but I just remembered.}
 
@@
 	\includegraphics[width=\textwidth]{droughts_number_events}
-	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions.
-\textcolor{magenta}{Not sure if it is worth a mention, but in case you had not made the connection: this is consistent with atlite having the highest MBE out of the three in the verification}
 
-For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the current energy mix.
+For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the existing energy mix.
 In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
@@
 Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
-For instance, in the current scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
+For instance, in the existing scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
 This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influences drought duration estimates.
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{droughts_return_periods}
-	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\includegraphics[width=\textwidth]{droughts_return_periods}
+	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
-In the projected scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
+In the balanced scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
 The balanced mix of wind and PV in this scenario reduces the seasonal signal, slightly increasing the percentage of RES drought hours in winter months while significantly decreasing it in summer.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-\textcolor{magenta}{This is a good point to mention the seasonality of the demand in Ireland that Damian brought up, although it could also go in the conclusions}
+This shift is especially relevant, as energy demand is generally higher in winter than in summer, making winter droughts more critical in terms of grid security.
+By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns.
 
@@
 
-\textcolor{magenta}{So, here the panel names are different as in the other ones, with existing and balanced.
-this is what I meant by consistency.
-Also, note the panel a in the figure is labelled as c.
-that should be changed.
-}
-
 \begin{figure}[!ht]
 	\centering
-	\includegraphics[width=\textwidth]{drouhgts_varying_threshold}
-	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\includegraphics[width=\textwidth]{droughts_varying_threshold}
+	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment.
@@
 
-\begin{itemize}
-	\item Remind differences between three datasets
-	\begin{itemize}
-		\item Atlite and C3S-E G: incorporate locations of RES farms
-		\item Atlite models better follows EirGrid data (for Wind)
-	\end{itemize}
-	\item Point Arguments with examples from the figures
-	\begin{itemize}
-		\item The two C3S-E datasets show similar results
-		\item Atlite stands out of these two datasets
-	\end{itemize}
-	\item Conclusion: Location of RES farms has less impact than the accuracy of modelling the CF!
-	\\
-	\item Impact of the 2030 scenario (Anti correlation between wind and PV) | Seasonality
-	\\
-	\item Risk for policymakers. The more accurate Atlite model produces more RES drought events and longer RES drought durations, an important consideration for risk-adverse decision making.
-	\\
-	\item Where to go next ?
-	\begin{itemize}
-		\item Looking at future climate data (Combine Climate and Energy scenarios)
-		\item Studying broader region (Europe), which is more relevant as everything will be interconnected but more difficult as we will need to repeat verification for all EU countries
-	\end{itemize}
-\end{itemize}
-
 This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts.
 One of the most significant differences is how each dataset incorporates the specific locations of RES farms.
-Both Atlite and C3S-E G take into account the exact locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
-However, our analysis shows that the Atlite dataset generally performs better, particularly in its wind generation estimates, where it closely follows the EirGrid data.
-This suggests that while the inclusion of RES farm locations is important, the accuracy of the CF modeling \textcolor{magenta}{I would say power curve model instead of CF modelling} used by Atlite may be the determining factor in its superior performance.\textcolor{magenta}{Superior how?
-Is finding more droughts superior or is it reproducing reality?
-I reckon it's the second, but make it explicit.
-Also, there is no explicit mention of PV, if you mention wind explicitly (which I agree on doing) I would add PV.
-Even though it is replacing that "However" with "Even though this is true for PV, our analysis shows that for wind..."}
+Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
+While this approach improves PV drought models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with EirGrid data for wind generation estimates.
+This finding suggests that although the inclusion of RES farm locations is beneficial, particularly for PV, the accuracy of the power curve model in Atlite may be the critical factor in producing realistic time series for low-generation events.
 
-When analysing the results, we found that the two C3S-E datasets (C3S-E G and C3S-E N) consistently show similar outcomes across various figures \textcolor{magenta}{I would not refer to them as figures, but as analysis or something like that.
-But if you do choose the word analysis, do not start the sentence with "When analysing the results..." and instead use something like "The results show..."} (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods},~\ref{fig:res_droughts_seasonality}), suggesting that their differences in methodology have minimal impact on the results.
-In contrast, Atlite stands out in its ability to model wind generation, as evidenced by the more accurate correlation with EirGrid data seen in the verification process (Section \ref{sec:wind_verification}).
-This distinction is further underscored in the figures \textcolor{magenta}{Same as before, avoid referring to the figures.
-We do not discuss the result (that was done in results), we discuss what it implies}, where Atlite often reports higher or more accurate return periods and a greater number of RES droughts (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods}), particularly in scenarios with increased PV capacity.
-\textcolor{magenta}{For me, the point of this paragraph is quite clear but does not really come across at the moment: Atlite is the best one at reproducing reality (as established in the previous paragraph).
-At the same time, a long-term analysis shows more droughts for Atlite than any other model.
-This means that model selection is key to correctly quantify the risk and avoid undersizing the reserve capacity.}
+The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
+In contrast, Atlite stands out for its closer alignment with EirGrid data, particularly in wind generation, as seen in the verification process (Section \ref{sec:wind_verification}).
+This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with balanced share of RES.
+One of the study's most critical findings is that the accuracy of RES drought modeling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
+Atlite’s superior performance highlights the importance of selecting accurate models for assessing RES drought risks.
+This careful model selection can better quantify risks, support effective planning, and avoid the potential underestimation of capacity needs, which is essential for ensuring energy security.
 
-One of the most critical conclusions of this study is that the specific locations of RES farms, while important, have a lower impact on the accuracy of RES drought modelling than the precision of the CF models themselves.
-\textcolor{magenta}{Same as before, I would not refer to CF models, but instead wind power curve and PV panel models} Atlite's superior performance suggests that refining these models could lead to better forecasting and planning for RES droughts, which is crucial for energy security.
-
-Looking to the balanced scenario, our analysis suggests a significant shift in the dynamics of RES droughts due to the anti-correlation between wind and PV generation.
-This scenario reveals the importance of seasonality in RES droughts (Fig.~\ref{fig:res_droughts_seasonality}), where a diversified energy portfolio that includes both wind and PV could decrease the frequency and duration of RES droughts.
-\textcolor{magenta}{I would merge these two paragraphs into one.
-Here are my thoughts in order: 1.
-We have found how wind and PV together perform better than individually in terms of droughts.
-2.
-A specific mention of seasonality and how it is reduced.
-3.
-Mention how Ireland has a very wind-dependent system and how targets for PV are ambitious in the coming years, almost making wind and PV capacities equal (maybe not the phrasing but you see the point, putting forward the balance again).
-4.
-Link to policymakers in terms of how meeting targets of installed capacity can help improve the energy security thanks to this diversification.
-5.
-Close with the importance of the right model selection for RES drought assessment, especially as more renewable capacity comes into the system.}
-For policymakers, the findings of this study highlight the risks of inadequate modelling and planning.
-Accurate RES drought forecasting is essential for ensuring energy security, especially as Ireland moves toward an ambitious renewable energy target.
+Looking at the balanced scenario, the analysis indicates a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
+Wind and PV together perform better in terms of drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
+This diversification reduces the seasonal impact on RES droughts, as PV generation peaks in the summer and wind generation is more consistent in winter.
+Ireland currently has a highly wind-dependent energy system, but with ambitious targets for PV installations in the coming years, the energy mix is expected to approach a balance between wind and PV capacity.
+This balanced approach offers a more stable and secure energy supply by mitigating RES drought risks.
+For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.
+Additionally, the choice of model for RES drought assessment becomes increasingly critical as more renewable capacity is integrated into the system.
+Accurate forecasting and careful modelling are essential to support reliable energy planning in Ireland’s renewable future.
 

### 2024-11-21 Article/main.tex
Update Article
Minor changes made, larger changes flagged.

@@
 %opening
-\title{The Impact of Model Assumptions on RES Drought Analysis}
+\title{ \textcolor{blue}{Change Title} The Impact of Model Assumptions on RES Drought Analysis}
 \author{Boris Morin, Damian Flynn, Aina Maimó Far, Conor Sweeney}
@@
 
+\textcolor{blue}{Add Abstract}
+
 Keywords: 
 \end{abstract}
-
-\textcolor{magenta}{Some things to keep in mind when reading from Damian's comments:
-	1. The whole current/projected and existing/balanced thing. Pick one and stick to it, including the figures (also fix the a/c level in fig 10)
-	2. See if you are happy with the section titles throughout the document}
 
@@
 
-As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy.
+\textcolor{blue}{Get rid of ``we" everywhere.}
+
+\textcolor{blue}{Make sure past tense is used everywhere, e.g. ``was", not ``is"}
+
+\textcolor{blue}{Merge and shorten first two paragraphs} As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy.
 Under the revised European Green Deal \cite{greendeal2023report}, the EU has set ambitious targets to achieve at least 69\% of electricity generation from renewable energy sources (RES) by 2030.
@@
 
-However, as reliance on RES grows, so does the challenge of managing the variability of weather-dependent energy sources like wind and PhotoVoltaic (PV) power.
+However, as reliance on RES grows, so does the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
 This challenge is further intensified by the increasing electrification of energy sectors which places additional demand on the power system~\cite{bloomfield2021}.
 As more sectors shift from fossil fuels to electricity, the power grid becomes increasingly sensitive to meteorological conditions~\cite{bloomfield2016, vanderWiel2019drought}.
-In this context, periods of low renewable generation often referred to as \textit{Dunkelflaute} or RES droughts, pose a greater risk to grid stability and energy security.
+In this context, periods of low renewable generation, often referred to as \textit{Dunkelflaute} or RES droughts, pose a greater risk to grid stability and energy security.
 Ensuring a reliable electricity supply during these RES droughts is crucial for maintaining a resilient energy system capable of meeting both the growing demand for electricity and the EU’s decarbonization targets.
 
-For this study, a RES drought event is defined when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology established in previous research~\cite{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
-Other studies have employed varying methodologies for defining RES droughts.
-One approach uses relative thresholds that change over the course of the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
-Another common method relies on quantile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\cite{bracken2024drought, allen2023drought}.
+\textcolor{blue}{Insert sales pitch paragraph here}
+
+For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in other research~\cite{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
+Alternative methodologies exist for defining RES droughts.
+One approach uses relative CF thresholds that change over the course of the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
+Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\cite{bracken2024drought, allen2023drought}.
 Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\cite{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
-In this paper, the focus is exclusively on energy generation, using a fixed threshold approach to define RES droughts.
+In this paper, the focus is exclusively on energy generation, and a fixed threshold approach to define RES droughts is used.
 
-RES drought events can be identified from wind and PV CF time series.
-In this study, we use three different datasets, all of them driven by ERA5 data~\cite{hersbach2020era5}.
-The first two datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy} (C3S).
-One of the C3S-E datasets provides CF time series aggregated at the national scale.
-Another C3S-E dataset provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset is produced by the authors using the Atlite model~\cite{hofman2021atlite}, which converts the  ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+RES drought are identified using wind and PV CF time series.
+In this study, three different datasets are used, all of which are driven by ERA5 data~\cite{hersbach2020era5}.
+Two of the datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy} (C3S).
+One of the C3S-E datasets used provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
+The third dataset used was generated using the Atlite model~\cite{hofman2021atlite}, which converts the  ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite}, and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
 
 This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
-The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected energy scenarios, featuring a more balanced share between wind and PV.
-A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.
+The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected future energy scenario, featuring a more balanced share between wind and PV.
+A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, duration and seasonality are presented.
 
@@
 This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
-The primary data sources include EirGrid, the transmission system operator for the ic of Ireland, and the ERA5 reanalysis dataset.
+The primary data sources include: EirGrid, the transmission system operator for the island of Ireland; the ERA5 reanalysis dataset; and the C3S-E datasets.
 
@@
 EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
-These datasets include critical information such as each farm’s installed capacity, name, and connection date.
+These datasets include information such as each farm’s installed capacity, name, and connection date.
 To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
 
-The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
+\textcolor{blue}{Rewrite this paragraph:} The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
 For this study, availability data, which indicates the amount of energy being delivered to the grid, is used.
@@
 
-Detailed data analysis, supported by access to additional data from EirGrid, revealed that the PV installed capacity available on the spreadsheets is the Maximum Export Capacity (MEC) and does not accurately reflect the true generating capacity.
-Instead, MEC values from the spreadsheet are used as a basis.
-To adjust for actual grid conditions, installed capacities are assumed to be 1.4 times the MEC values.
-Whenever farm generation is calculated, this scaled capacity is used and any farm output is capped to the MEC.
+The PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed PV capacity.
+To adjust for actual PV generation potential, installed capacities are estimated to be 1.4 times the MEC values.
+This installed capacity is used for PV genreation, with farm output capped to the MEC value.
 
@@
 It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
-Table~\ref{tab:var_name} lists the ERA5 variables used in the models to generate the Atlite and C3S-Energy datasets.
+Table~\ref{tab:var_name} lists the ERA5 variables which have been used by Atlite and C3S-Energy.
 
@@
 	\end{tabular}
-	\caption{Variables retrieved from ERA5 dataset}
+	\caption{ERA5 variables used to calculate wind and PV generation.}
 	\label{tab:var_name}
@@
 The C3S-E dataset estimates wind energy using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 m.
-This choice is based on expert advice and actual trend in wind turbine installation.
-The PV generation model used by C3S-E uses two ERA5 variables: global horizontal irradiance (referred to here as $ssrd$) and air temperature ($t2m$).
-PV generation is calculated multiple times using the same model with different azimuth and tilt angles.
+This choice is based on expert advice and the trend in wind turbine installation.
+The PV generation model used by C3S-E uses two ERA5 variables: Surface solar radiation downwards ($ssrd$) and air temperature ($t2m$).
+PV generation is calculated multiple times, using the same model with different azimuth and tilt angles.
 The results are aggregated based on a statistical distribution of the module angles based on the geographical location~\cite{saintdrenan2018solar}.
@@
 
-In this study, we use three datasets to analyse RES droughts.
-Data downloaded from C3S-E are used to obtain two datasets: one based on national-level (C3S-E N), and another on grid-level (C3S-E G).
+In this study, we use three datasets to analyse RES droughts for teh island of Ireland.
+Data downloaded from C3S-E are used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
 The third dataset is computed using the Atlite model (Atlite).
@@
 
-For national-level analyses, we use the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+\textcolor{blue}{Rewrite:} For national-level analyses, we use the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
 We compute a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland.
@@
 
-Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite}, and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
-This model transforms weather data into energy data, using the locations of existing RES farms as described in C3S-E G.
+Atlite transforms weather data into energy data, using the wind speed at 100 metres ($u_{100}$, $v_{100}$), and the locations of existing RES farms as described in C3S-E G.
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
 
-For wind power estimation, Atlite uses the wind speed at 100 metres ($u_{100}$, $v_{100}$).
-We provide Atlite with a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake} at a hub height of 100 metres, which it uses when converting wind speed to CF.
+This study uses a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake}.
 For PV power estimation, Atlite uses four irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) as well as the air temperature ($t2m$).
@@
 
-This study investigates wind and PV capacity factors (CF) using three different models, resulting in six individual CF time series based on the installed capacities of 2023.
 In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
-This results in a total of nine CF time series for the existing scenario, which represents the existing distribution of RES installations in Ireland.
-
-Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario.
-In this scenario, the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW, based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
-For PV, new time series are generated for both the Atlite and C3S-E G models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
+These values are based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
+New time series are generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind, the CF time series remain unchanged, as significant shifts in wind farm distribution are not expected.
@@
 This approach allows for a comparative analysis between a wind-dominated system (the existing scenario) and a more evenly distributed system (the balanced scenario), assessing how the balance of RES capacity affects the occurrence of RES droughts.
-The results for these two scenarios are analysed and discussed in the following sections.
 
@@
 
-In this study, a RES drought event is defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period longer than 24 hours.
-The choice of this threshold is somewhat arbitrary but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought} .
+In this study, a RES drought event is defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
+The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought} .
 By using a 24-hour moving average, fewer but longer-lasting events are captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
-Although other methods exist for identifying RES droughts, a fixed threshold has been chosen in this study to enable consistent inter-comparison across the three datasets.
+A fixed threshold approach has been chosen in this study to enable consistent inter-comparison between datasets.
 
@@
 
-Fig.~\ref{fig:find_res_droughts} illustrates the moving average identification method.
-When applying the moving average, this CF time series identifies a single event lasting 161 hours.
-In contrast, using the raw CF series would have resulted in three separate events, with durations of 36, 35, and 40 hours.
-
-The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period.
+The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period (Fig.~\ref{fig:find_res_droughts}).
 This means that a single hour above the threshold does not "break" a drought event if it is surrounded by prolonged low-generation hours.
@@
 
-Before going into the analysis of RES droughts, it is important to first verify the accuracy of the datasets used in this study.
+The accuracy of the datasets used in this study are verified, before continuing to the analysis of RES droughts.
 For the verification process, we use time-varying values of installed capacity to account for changes in RES development over the verification period.
@@
 The smoothing is done by using a Gaussian filter with different standard deviations that depend on the wind speed.
-A wind CF time series for Ireland for all wind turbine power curve options and all smoothings is generated.
-The performance of each wind turbine power curve is then calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
+A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
+The performance of each CF time series was then calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
 The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
@@
 
-The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines, which is important when modelling wind energy on larger scales.
+The smoothing of the wind turbine power curve represents losses associated with each individual turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour}
-	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
+	\caption{\textcolor{blue}{Save as png?
+Investigate vertical blobs near zero in (c)}Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
 	\label{fig:wind_verification_contour}
@@
 
-The effect of the differences in the power curves is also visible in Fig.~\ref{fig:wind_verification_contour} where the Atlite model shows good agreement with the EirGrid data.
-In contrast, the two C3S-E datasets tend to overestimate the EirGrid CF.
-This is confirmed by Table~\ref{tab:wind_skill_scores}, which shows that Atlite performs better than the C3S-E datasets for all three skill scores.
+The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
+The two C3S-E datasets are shown to overestimate the EirGrid CF, whereas the Atlite model is in good agreement with the EirGrid data.
+The skill scores presented in Table~\ref{tab:wind_skill_scores} show that Atlite performs better than the C3S-E datasets for all of the skill scores.
 
@@
 
-The C3S-E datasets calculate PV generation multiple times using the same model with varying azimuth and tilt angles.
-The results are then aggregated based on a statistical distribution linked to each geographical location.
-Similarly to the wind CF verification, the Atlite model allows the user to select a given PV panel characteristics.
-In this study we have chosen to look at the three PV panels offered by in the Atlite model.
+The Atlite model allows the user to select certain PV panel characteristics.
+In this study we consider the three PV panels types available in the Atlite model.
 Following the same methodology as in the previous section, we compare the three available models using four skill scores (CC, RMSE, MB, and area under the curve).
@@
 With over 60\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
-While the results are reasonably aligned in some ranges, the inherent challenges in PV data highlight the limitations of this verification.
 
@@
 Both scenarios are driven by 45 years of ERA5 data.
-Using the RES drought identification process described earlier (Section~\ref{sec:res_drought}), we first analyse wind and PV droughts separately before presenting the results for combined (wind + PV) RES droughts.
+Using the RES drought identification process described in Section~\ref{sec:res_drought}, we first analyse wind and PV droughts separately before presenting the results for combined (wind + PV) RES droughts.
 
@@
 In contrast, the balanced scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
-This reduction is attributed to the anti-correlation between wind and PV production throughout the year, a phenomenon well-documented in the literature \cite{kaspar2019drought} and discussed in further detail in the later analysis.
+This reduction is attributed to the anti-correlation between wind and PV production throughout the year, a phenomenon well-documented in the literature \cite{kaspar2019drought}.
 
@@
 Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
-This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the potential impact of RES droughts on energy reliability.
+This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the potential impact of RES droughts on energy reliability and security of supply.
 
@@
 
-For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves significantly different than the two C3S-E datasets.
-The Atlite results follow the ones for wind with a log-linear increase, but reach higher values in general with the longest event lasting forty days.
+For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
+The Atlite results show a log-linear increase, but reach higher values in general with the longest event lasting forty days.
 For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days.
@@
 This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months.
-The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold.
-Atlite appears to hover slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
+The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold of 0.1 CF.
+Atlite remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
 In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
@@
 
-For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the existing energy mix.
+For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), due to the low levels of installed PV.
 In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
-For example, the return period for a five-day drought event extends from roughly six months to four years in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
-The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
+For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the existing scenario, to four years for the balanced scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
+\textcolor{blue}{Shorten or remove: }The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
 This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics.
@@
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-The balanced mix of wind and PV in this scenario reduces the seasonal signal, slightly increasing the percentage of RES drought hours in winter months while significantly decreasing it in summer.
-This shift is especially relevant, as energy demand is generally higher in winter than in summer, making winter droughts more critical in terms of grid security.
+The balanced mix of wind and PV in this scenario reduces the seasonal signal overall, and significantly decreases the percentage of RES drought hours in the summer.
 By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns.
@@
 Overall, the four panels indicate that regardless of the threshold applied, the Atlite dataset consistently identifies more events than the two C3S-E datasets, except for PV when the threshold is below 0.08.
-This finding suggests that Atlite, likely due to its more accurate representation of wind patterns in Ireland, captures more RES drought events that might be missed with different modeling assumptions in the C3S-E datasets.
+This finding suggests that Atlite, likely due to its more accurate representation of wind generation in Ireland, captures more RES drought events that might be missed with different modeling assumptions in the C3S-E datasets.
 
@@
 Additionally, the choice of model for RES drought assessment becomes increasingly critical as more renewable capacity is integrated into the system.
-Accurate forecasting and careful modelling are essential to support reliable energy planning in Ireland’s renewable future.
 

### 2024-11-25 Article/main.tex
Update Article

@@
 
-Keywords: 
+Keywords: Energy Droughts, Wind Power, Solar PV Power, Renewable Energy Sources, Seasonality, Return Periods, 
 \end{abstract}
@@
 
-\textcolor{blue}{Merge and shorten first two paragraphs} As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy.
+\textcolor{blue}{Merge and shorten first two paragraphs}
+
+As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy.
 Under the revised European Green Deal \cite{greendeal2023report}, the EU has set ambitious targets to achieve at least 69\% of electricity generation from renewable energy sources (RES) by 2030.
@@
 For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in other research~\cite{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
-Alternative methodologies exist for defining RES droughts.
-One approach uses relative CF thresholds that change over the course of the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
+Alternative methods exist for defining RES droughts.
+One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
 Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\cite{bracken2024drought, allen2023drought}.
@@
 
-RES drought are identified using wind and PV CF time series.
+RES droughts are identified using wind and PV CF time series.
 In this study, three different datasets are used, all of which are driven by ERA5 data~\cite{hersbach2020era5}.
 Two of the datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy} (C3S).
-One of the C3S-E datasets used provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset used was generated using the Atlite model~\cite{hofman2021atlite}, which converts the  ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite}, and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
+One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
+The third dataset was generated using the Atlite model~\cite{hofman2021atlite}, which converts the  ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite} and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
 
@@
 To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
+For simplicity, this data will be referred to as originating from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
 
-\textcolor{blue}{Rewrite this paragraph:} The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
-For this study, availability data, which indicates the amount of energy being delivered to the grid, is used.
+The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
+Availability refers to the total energy generated by the RES farm, including any energy that is intentionally reduced by EirGrid or SONI to maintain grid stability.
 This value is combined with installed capacity to compute the CF, the main metric for assessing renewable energy output.
@@
 To adjust for actual PV generation potential, installed capacities are estimated to be 1.4 times the MEC values.
-This installed capacity is used for PV genreation, with farm output capped to the MEC value.
+This installed capacity is used for PV generation, with farm output capped to the MEC value.
 
@@
 
-In this study, we use three datasets to analyse RES droughts for teh island of Ireland.
+This study relies on three datasets to analyse RES droughts across the island of Ireland.
 Data downloaded from C3S-E are used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
@@
 
-\textcolor{blue}{Rewrite:} For national-level analyses, we use the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+For national-level analyses, the aggregated CF time series provided by C3S-E are used at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
 We compute a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland.
@@
 
-EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
-These datasets include critical information such as each farm’s installed capacity, name, and connection date.
-To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
-For simplicity, this data will be referred to as originating from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
+For grid-level analyses, we use the gridded dataset from C3S-E to create a representative time series for Ireland's RES farms.
+We compile a comprehensive list of RES farms from the EirGrid website, which includes each farm’s GPS coordinates and installed capacity.
+Using these coordinates, we identify the nearest grid point on the C3S-E grid for each farm.
+For each farm, we retrieve the CF values from the C3S-E dataset for its corresponding grid point.
+To construct the final CF time series, we calculate a weighted average of the CF values, using the installed capacity of each farm as the weight.
+This process results in a single time series for each energy source (wind and PV), providing an accurate representation of the RES generation in Ireland at a grid level.
+
+For grid-level analyses, the gridded dataset from C3S-E is used to create a representative time series for Ireland's RES farms.
+A comprehensive list of RES farms, including each farm’s GPS coordinates and installed capacity, is compiled from the EirGrid website.
+Using these coordinates, the nearest grid point on the C3S-E grid is identified for each farm.
+The CF values from the C3S-E dataset corresponding to these grid points are retrieved for each farm.
+To construct the final CF time series, a weighted average of the CF values is calculated, with the installed capacity of each farm serving as the weight.
+This process results in a single time series for each energy source (wind and PV), providing an accurate representation of RES generation in Ireland at a grid level.
 
@@
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
-
 This study uses a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake}.
 For PV power estimation, Atlite uses four irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) as well as the air temperature ($t2m$).
-We use a single PV model~\cite{beyer2004pv}, with the Kaneka Hybrid panel option.
-The azimuth angle is fixed at 180\textdegree (due south) and we select the optimal tilt angle option.
-
-The selection of a specific wind turbine and PV panel characteristics, are further discussed and explained in section \ref{sec:verification}.
+A single PV model~\cite{beyer2004pv}, with the Kaneka Hybrid panel option, is used.
+The azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
+The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
 
@@
 New time series are generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
-For wind, the CF time series remain unchanged, as significant shifts in wind farm distribution are not expected.
+For wind, the CF time series remains unchanged, as significant shifts in wind farm distribution are not expected.
 The PV and wind time series are then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
@@
 In this study, a RES drought event is defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought} .
+The choice of this threshold is somewhat arbitrary but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought}.
 By using a 24-hour moving average, fewer but longer-lasting events are captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
@@
 
-The accuracy of the datasets used in this study are verified, before continuing to the analysis of RES droughts.
+The accuracy of the datasets used in this study is verified, before continuing to the analysis of RES droughts.
 For the verification process, we use time-varying values of installed capacity to account for changes in RES development over the verification period.
@@
 We considered the 121 power curves available for download from Renewables.ninja~\cite{staffell2016wake}.
-For each power curve, Renewables.ninja also provide four associated smoothed power curves.
-The smoothing is done by using a Gaussian filter with different standard deviations that depend on the wind speed.
+For each power curve, Renewables.ninja also provides four associated smoothed power curves.
+The smoothing is done using a Gaussian filter with different standard deviations that depend on the wind speed.
 A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
@@
 
-The smoothing of the wind turbine power curve represents losses associated with each individual turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
+The smoothing of the wind turbine power curve represents losses associated with each turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
@@
 The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
-This pattern is particularly evident for shorter duration events, which are the most frequent.
+This pattern is particularly evident for shorter-duration events, which are the most frequent.
 
@@
 The Atlite model allows the user to select certain PV panel characteristics.
-In this study we consider the three PV panels types available in the Atlite model.
-Following the same methodology as in the previous section, we compare the three available models using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, we selected the Breyer PV panel model \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
+In this study, we consider the three PV panel types available in the Atlite model.
+Following the same methodology as in the previous section, the three available models are being compared using four skill scores (CC, RMSE, MB, and area under the curve).
+Based on the best-performing metrics, the Breyer PV panel model is selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
 
@@
 
-Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency of overestimating the CF compared to EirGrid, especially for high CF values.
+Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to EirGrid, especially for high CF values.
 The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
@@
 All models show a slight positive bias, with Atlite exhibiting the most overestimation.
-These results reflect the broader challenges in accurately modeling PV, where variability is hard to capture.
-Nonetheless, C3S-E G’s lower RMSE suggests it may be more reliable for PV CF modeling in this context.
+These results reflect the broader challenges in accurately modelling PV, where variability is hard to capture.
+Nonetheless, C3S-E G’s lower RMSE suggests it may be more reliable for PV CF modelling in this context.
 
@@
 
-In this section, we evaluate RES drought events under two different scenarios with fixed installed capacities: the existing scenario, with 89\% of wind capacity (6.7 GW) and 11\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 89\% of wind capacity (6.7 GW) and 11\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
 Both scenarios are driven by 45 years of ERA5 data.
-Using the RES drought identification process described in Section~\ref{sec:res_drought}, we first analyse wind and PV droughts separately before presenting the results for combined (wind + PV) RES droughts.
+Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and PV droughts are first analysed separately before presenting the results for combined (wind + PV) RES droughts.
 
@@
 For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
-The Atlite results show a log-linear increase, but reach higher values in general with the longest event lasting forty days.
+The Atlite results show a log-linear increase but reach higher values in general with the longest event lasting forty days.
 For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days.
@@
 For instance, in the existing scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
-This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influences drought duration estimates.
+This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influence drought duration estimates.
 Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
@@
 This seasonal variation is influenced by the wind power curve model used to estimate CF, where the shape of the curve in lower wind speed regions (3-10 m/s) leads to significant differences in CF under low wind conditions.
-In contrast the results for PV (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
+In contrast, the results for PV (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
 The Atlite results show a higher percentage of RES drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.
@@
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-The balanced mix of wind and PV in this scenario reduces the seasonal signal overall, and significantly decreases the percentage of RES drought hours in the summer.
+The balanced mix of wind and PV in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
 By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns.
@@
 Overall, the four panels indicate that regardless of the threshold applied, the Atlite dataset consistently identifies more events than the two C3S-E datasets, except for PV when the threshold is below 0.08.
-This finding suggests that Atlite, likely due to its more accurate representation of wind generation in Ireland, captures more RES drought events that might be missed with different modeling assumptions in the C3S-E datasets.
+This finding suggests that Atlite, likely due to its more accurate representation of wind generation in Ireland, captures more RES drought events that might be missed with different modelling assumptions in the C3S-E datasets.
 
@@
 In contrast, Atlite stands out for its closer alignment with EirGrid data, particularly in wind generation, as seen in the verification process (Section \ref{sec:wind_verification}).
-This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with balanced share of RES.
-One of the study's most critical findings is that the accuracy of RES drought modeling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
+This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
+One of the study's most critical findings is that the accuracy of RES drought modelling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
 Atlite’s superior performance highlights the importance of selecting accurate models for assessing RES drought risks.

### 2024-11-26 Article/main.tex
Update Article

@@
 
-\textcolor{blue}{Add Abstract}
+As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, managing their variability becomes a critical challenge for energy systems.
+This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
+Using three datasets RES droughts are analysed over 45 years to assess their frequency, duration, and return periods.
+The methodology involves comparing modelled capacity factors against observed data to evaluate the accuracy of each dataset and applying a fixed threshold to identify RES drought events.
+Results show that adding PV capacity significantly reduces the frequency and duration of RES droughts, with a more balanced energy mix minimising seasonal drought patterns.
+These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
+\\
 
-Keywords: Energy Droughts, Wind Power, Solar PV Power, Renewable Energy Sources, Seasonality, Return Periods, 
+\textbf{Keywords}: Energy Droughts, Wind Power, Solar PV Power, Renewable Energy Sources, Seasonality, Return Periods, 
 \end{abstract}
@@
 
-\textcolor{blue}{Get rid of ``we" everywhere.}
-
 \textcolor{blue}{Make sure past tense is used everywhere, e.g. ``was", not ``is"}
 
-\textcolor{blue}{Merge and shorten first two paragraphs}
-
-As the global shift towards renewable energy accelerates, the European Union (EU) has positioned itself as a leader in climate and energy policy.
-Under the revised European Green Deal \cite{greendeal2023report}, the EU has set ambitious targets to achieve at least 69\% of electricity generation from renewable energy sources (RES) by 2030.
-This is a significant increase from 2022, when RES accounted for over 41\% of the EU’s electricity generation \cite{europe2023stat}.
-Meeting this goal is a key step in reducing greenhouse gas emissions and ensuring a sustainable energy future.
-
-However, as reliance on RES grows, so does the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
-This challenge is further intensified by the increasing electrification of energy sectors which places additional demand on the power system~\cite{bloomfield2021}.
-As more sectors shift from fossil fuels to electricity, the power grid becomes increasingly sensitive to meteorological conditions~\cite{bloomfield2016, vanderWiel2019drought}.
-In this context, periods of low renewable generation, often referred to as \textit{Dunkelflaute} or RES droughts, pose a greater risk to grid stability and energy security.
-Ensuring a reliable electricity supply during these RES droughts is crucial for maintaining a resilient energy system capable of meeting both the growing demand for electricity and the EU’s decarbonization targets.
+As the global shift towards renewable energy accelerates, the European Union (EU) has emerged as a leader in climate and energy policy.
+Under the revised European Green Deal \cite{greendeal2023report}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{europe2023stat} \textcolor{cyan}{Can't seem to find a proper reference}.
+While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
+This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\cite{bloomfield2021, bloomfield2016, vanderWiel2019drought}.
+Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to grid stability and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
 
 \textcolor{blue}{Insert sales pitch paragraph here}
+\begin{comment}
+This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
+The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected future energy scenario, featuring a more balanced share between wind and PV.
+A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, duration and seasonality are presented.
+\end{comment}
+
+This study focuses on Ireland, a region with a strong reliance on wind energy and ambitious targets for PV expansion, to provide valuable insights into the potential benefits of diversifying the renewable energy mix.
+This work compares the performance of energy models and highlights the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
 
@@
 Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite} and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
-
-This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
-The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected future energy scenario, featuring a more balanced share between wind and PV.
-A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, duration and seasonality are presented.
 
@@
 
-For grid-level analyses, we use the gridded dataset from C3S-E to create a representative time series for Ireland's RES farms.
-We compile a comprehensive list of RES farms from the EirGrid website, which includes each farm’s GPS coordinates and installed capacity.
-Using these coordinates, we identify the nearest grid point on the C3S-E grid for each farm.
-For each farm, we retrieve the CF values from the C3S-E dataset for its corresponding grid point.
-To construct the final CF time series, we calculate a weighted average of the CF values, using the installed capacity of each farm as the weight.
-This process results in a single time series for each energy source (wind and PV), providing an accurate representation of the RES generation in Ireland at a grid level.
-
 For grid-level analyses, the gridded dataset from C3S-E is used to create a representative time series for Ireland's RES farms.
@@
 The accuracy of the datasets used in this study is verified, before continuing to the analysis of RES droughts.
-For the verification process, we use time-varying values of installed capacity to account for changes in RES development over the verification period.
+For the verification process, time-varying values of installed capacity is used to account for changes in RES development over the verification period.
 This step allows us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour}
-	\caption{\textcolor{blue}{Save as png?
-Investigate vertical blobs near zero in (c)}Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
+	\caption{\textcolor{blue}{Investigate vertical blobs near zero in (c)}Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
 	\label{fig:wind_verification_contour}
@@
 The Atlite model allows the user to select certain PV panel characteristics.
-In this study, we consider the three PV panel types available in the Atlite model.
+In this study, the three PV panel types available are considered in the Atlite model.
 Following the same methodology as in the previous section, the three available models are being compared using four skill scores (CC, RMSE, MB, and area under the curve).

### 2024-11-27 Article/main.tex
Add changes to the article

@@
 %opening
-\title{ \textcolor{blue}{Change Title} The Impact of Model Assumptions on RES Drought Analysis}
+%\title{ \textcolor{blue}{Change Title} The Impact of Model Assumptions on RES Drought Analysis}
+\title{Reducing RES Droughts through the integration of wind and PV}
 \author{Boris Morin, Damian Flynn, Aina Maimó Far, Conor Sweeney}
@@
 
-As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, managing their variability becomes a critical challenge for energy systems.
-This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
+As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
+This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
+Three different modelled datasets are used, and their capacity factors are compared to observed data to quantify their overall skill.
+Then, using a RES drought definition with a fixed threshold, their skill at representing these events is assessed.
+These modelled capacity factors allow for an expansion over a 45-year period, allowing for an analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
+Results show the importance of model selection for RES drought risk assessment.
+On a similar note, the addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, with a more balanced energy mix minimising seasonal drought patterns.
+These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
+\textcolor{magenta}{I rewrote this and I still do not like it}
+
+% As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
+This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
 Using three datasets RES droughts are analysed over 45 years to assess their frequency, duration, and return periods.
@@
 
-\textcolor{blue}{Make sure past tense is used everywhere, e.g. ``was", not ``is"}
-
 As the global shift towards renewable energy accelerates, the European Union (EU) has emerged as a leader in climate and energy policy.
-Under the revised European Green Deal \cite{greendeal2023report}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{europe2023stat} \textcolor{cyan}{Can't seem to find a proper reference}.
+Under the revised European Green Deal \cite{greendeal2023report} \url{https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=SWD%3A2022%3A230%3AFIN&qid=1653033922121}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{europe2023stat} \url{https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Renewable_energy_statistics}.
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
@@
 
-\textcolor{blue}{Insert sales pitch paragraph here}
 \begin{comment}
@@
 
+This study focuses on Ireland, a region with a strong reliance on wind energy and ambitious targets for PV expansion.
+This makes it the perfect case study to provide valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
+This work also compares the performance of different energy models, revealing how modelling decisions impact the representation of RES droughts.
+The combination of these two offers insights about the effect of system planning decisions on RES droughts and the responses required to minimise their impact.
+
+\begin{comment}
 This study focuses on Ireland, a region with a strong reliance on wind energy and ambitious targets for PV expansion, to provide valuable insights into the potential benefits of diversifying the renewable energy mix.
 This work compares the performance of energy models and highlights the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
+\end{comment}
 
@@
 In this study, three different datasets are used, all of which are driven by ERA5 data~\cite{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy} (C3S).
+Two of the datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy}.
 One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
@@
 This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
-The primary data sources include: EirGrid, the transmission system operator for the island of Ireland; the ERA5 reanalysis dataset; and the C3S-E datasets.
+The primary data sources include: EirGrid, the transmission system operator (TSO) for the Republic of Ireland; the ERA5 reanalysis dataset; and the C3S-E datasets.
 
@@
 
-EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
+EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
 These datasets include information such as each farm’s installed capacity, name, and connection date.
 To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
-For simplicity, this data will be referred to as originating from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
+For simplicity, this data will be referred to as originating from EirGrid, as all-island data was directly obtained from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
 
-The spreadsheet available from the EirGrid website contains two key variables: generation and availability values for wind (from 2014 onward) and PV (from 2018 onward).
-Availability refers to the total energy generated by the RES farm, including any energy that is intentionally reduced by EirGrid or SONI to maintain grid stability.
-This value is combined with installed capacity to compute the CF, the main metric for assessing renewable energy output.
+The spreadsheet available from the EirGrid website contains two key variables: generation and availability.
+Generation is the energy that a RES farm actually contributed to the grid, which may include limitations introduced by the TSO to maintain grid stability, most commonly in the form of curtailment.
+Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
+This is then combined with installed capacity to compute the CF, the main variable for assessing renewable energy output.
+Generation and availability values are available from 2014 onward for wind and from 2018 onward for PV, although PV only became present in the Republic of Ireland data in 2023.
 
 The PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed PV capacity.
-To adjust for actual PV generation potential, installed capacities are estimated to be 1.4 times the MEC values.
-This installed capacity is used for PV generation, with farm output capped to the MEC value.
+To adjust for actual PV generation potential, installed capacities were estimated to be 1.4 times the MEC values.
+This installed capacity was used for PV generation, with farm output capped to the MEC value.
 
@@
 It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
-Table~\ref{tab:var_name} lists the ERA5 variables which have been used by Atlite and C3S-Energy.
+Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.
 
@@
 
-C3S has developed a renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
+The EU Copernicus Climate Change Service developed the C3S-E renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
 This dataset provides hourly CF for wind and PV energy from 1979 to the present.
@@
 This study relies on three datasets to analyse RES droughts across the island of Ireland.
-Data downloaded from C3S-E are used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
-The third dataset is computed using the Atlite model (Atlite).
+Data downloaded from C3S-E were used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
+The third dataset was computed using the Atlite model (Atlite).
 
@@
 
-For national-level analyses, the aggregated CF time series provided by C3S-E are used at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
-We compute a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland.
+For national-level analyses, the aggregated CF time series provided by C3S-E were used at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+We computed a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland.
 This CF time series for Ireland is based on the underlying C3S-E assumption of RES generation at every 0.25\textdegree grid point in Ireland.
@@
 
-For grid-level analyses, the gridded dataset from C3S-E is used to create a representative time series for Ireland's RES farms.
-A comprehensive list of RES farms, including each farm’s GPS coordinates and installed capacity, is compiled from the EirGrid website.
-Using these coordinates, the nearest grid point on the C3S-E grid is identified for each farm.
-The CF values from the C3S-E dataset corresponding to these grid points are retrieved for each farm.
-To construct the final CF time series, a weighted average of the CF values is calculated, with the installed capacity of each farm serving as the weight.
-This process results in a single time series for each energy source (wind and PV), providing an accurate representation of RES generation in Ireland at a grid level.
+For grid-level analyses, the gridded dataset from C3S-E was used to create a representative time series for Ireland's RES farms.
+A comprehensive list of RES farms, including each farm’s GPS coordinates and installed capacity, was compiled from the EirGrid website.
+Using these coordinates, the nearest grid point on the C3S-E grid was identified for each farm.
+The CF values from the C3S-E dataset corresponding to these grid points were retrieved for each farm.
+To construct the final CF time series, a weighted average of the CF values was calculated, with the installed capacity of each farm serving as the weight.
+This process resulted in a single time series for each energy source (wind and PV), providing an accurate representation of RES generation in Ireland at a grid level.
 
@@
 
-In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
-Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
+In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
+Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
 These values are based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
-New time series are generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind, the CF time series remains unchanged, as significant shifts in wind farm distribution are not expected.
-The PV and wind time series are then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
+The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
 
-In total, twelve CF time series are analysed in this study, six for individual wind and PV CF (three models for each source) in the existing scenario, and an additional six time series that include the combined CF for both existing and balanced scenarios across the different models.
+In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three models for each source) in the existing scenario, and an additional six time series that include the combined CF for both existing and balanced scenarios across the different models.
 
@@
 
-In this study, a RES drought event is defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
+In this study, a RES drought event was defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
 The choice of this threshold is somewhat arbitrary but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought}.
-By using a 24-hour moving average, fewer but longer-lasting events are captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
-A fixed threshold approach has been chosen in this study to enable consistent inter-comparison between datasets.
+By using a 24-hour moving average, fewer but longer-lasting events were captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
+A fixed threshold approach was chosen in this study to enable consistent inter-comparison between datasets.
 
@@
 
-The accuracy of the datasets used in this study is verified, before continuing to the analysis of RES droughts.
-For the verification process, time-varying values of installed capacity is used to account for changes in RES development over the verification period.
-This step allows us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
+The accuracy of the datasets used in this study was verified, before continuing to the analysis of RES droughts.
+For the verification process, time-varying values of installed capacity were used to account for changes in RES development over the verification period.
+This step allowed us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
 
@@
 A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
-The performance of each CF time series was then calculated based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
+The performance of each CF time series was then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
 The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
@@
 The Atlite model allows the user to select certain PV panel characteristics.
-In this study, the three PV panel types available are considered in the Atlite model.
-Following the same methodology as in the previous section, the three available models are being compared using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, the Breyer PV panel model is selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
+In this study, the three PV panel types available in the Atlite model were considered.
+Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and area under the curve).
+Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
 
@@
 The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, leading to uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
-With over 60\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
+With over 65\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
 
@@
 In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 89\% of wind capacity (6.7 GW) and 11\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
-Both scenarios are driven by 45 years of ERA5 data.
-Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and PV droughts are first analysed separately before presenting the results for combined (wind + PV) RES droughts.
+Both scenarios were driven by 45 years of ERA5 data.
+Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and PV droughts are first analysed separately before presenting the results for combined (wind + PV) RES droughts under both scenarios.
 
@@
 When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
-This is due to PV's typically lower CF compared to wind, especially in a region like Ireland where solar potential is limited.
+This is due to the typically lower CF of PV compared to wind, especially in a region like Ireland where solar potential is limited.
 PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of RES droughts in PV compared to wind.
@@
 This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months.
-The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold of 0.1 CF.
+The difference between Atlite and the C3S-E datasets arises from how each model handles PV output near the threshold of 0.1 CF.
 Atlite remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
@@
 For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the existing scenario, to four years for the balanced scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
-\textcolor{blue}{Shorten or remove: }The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
+% The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
 This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics.
@@
 
-The seasonality of RES droughts is analysed by comparing the percentage of hours in each month classified as part of a RES drought.
+The seasonality of RES droughts was analysed by comparing the percentage of hours in each month classified as part of a RES drought.
 
@@
 
-In this study, a threshold CF value of 0.1 has been used to identify RES droughts.
+In this study, a threshold CF value of 0.1 was used to identify RES droughts.
 To assess how sensitive the results are to this threshold, the annual number of events was calculated across a range of CF threshold values (Fig.~\ref{fig:number_days_threshold}).
@@
 
-This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts.
+This study highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts.
 One of the most significant differences is how each dataset incorporates the specific locations of RES farms.
@@
 
-The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
+The results showed that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
 In contrast, Atlite stands out for its closer alignment with EirGrid data, particularly in wind generation, as seen in the verification process (Section \ref{sec:wind_verification}).
-This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
+This distinction was also evident in the analysis, where Atlite reported higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
 One of the study's most critical findings is that the accuracy of RES drought modelling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
@@
 
-Looking at the balanced scenario, the analysis indicates a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
+Looking at the balanced scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
 Wind and PV together perform better in terms of drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
@@
 In terms of future research, there are two main topics to explore.
-First, integrating future climate data with energy scenarios to better understand how climate change might affect RES droughts.
-Second, expanding the scope of the study to include Europe would provide a more comprehensive understanding of RES drought in an interconnected energy grid.
+First, future climate data could be integrated with energy scenarios to better understand how climate change might affect RES droughts.
+Second, expanding the scope of the study to include Europe would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
 However, this expansion would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.
@@
 \newpage
-\section{Acknowledgment}
+\section{Acknowledgments}
 Using the Climate Data Store. Thank EirGrid for the data and add a reference to the funding project or projects

### 2024-11-27 Article/main.tex
Update Article

@@
 \usepackage{amsmath} % Math equation
+
+\usepackage{url}
 
@@
 
-As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
+\noindent As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
 This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
@@
 As the global shift towards renewable energy accelerates, the European Union (EU) has emerged as a leader in climate and energy policy.
-Under the revised European Green Deal \cite{greendeal2023report} \url{https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=SWD%3A2022%3A230%3AFIN&qid=1653033922121}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{europe2023stat} \url{https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Renewable_energy_statistics}.
+Under the revised European Green Deal \cite{greendeal2022report}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{eurostat2023share}.
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
@@
 
-EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}.
+EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid2023spreadsheet}.
 These datasets include information such as each farm’s installed capacity, name, and connection date.
@@
 	\includegraphics{droughts_methodology}
-	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021.
+	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021.
 The black dashed line indicates the CF threshold.
@@
 	\includegraphics[width=\textwidth]{verification_power_curve}
-	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
+	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b/ Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid (shaded)}
 	\label{fig:power_curve}
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour}
-	\caption{\textcolor{blue}{Investigate vertical blobs near zero in (c)}Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
+	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
 	\label{fig:wind_verification_contour}

### 2024-11-28 Article/main.tex
Add some changes

@@
 \noindent As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
-This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
-Three different modelled datasets are used, and their capacity factors are compared to observed data to quantify their overall skill.
-Then, using a RES drought definition with a fixed threshold, their skill at representing these events is assessed.
-These modelled capacity factors allow for an expansion over a 45-year period, allowing for an analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
+This study investigates periods of low renewable generation, known as RES droughts, with a focus on Ireland, its current wind-dominated energy mix and the more balanced target distribution of wind and PV capacity.
+Three different models are used to estimate the capacity factors of the different scenarios, and their overall skill is quantified by comparing them to observed data.
+Their skill at representing RES droughts is also calculated by using a RES drought definition with a fixed threshold.
+These models allow for an expansion of the capacity factor series over a 45-year period, provide the opportunity for an analysis of the frequency, duration and return periods of RES droughts at a climatological scale, identifying risks that may not be explicit at the short observed time scales.
 Results show the importance of model selection for RES drought risk assessment.

### 2024-11-29 Article/main.tex
Updated Figures
Update Figures
CS edits up to Sec 4.

--- 2024-11-27 Article/main.tex
@@
 
-\noindent As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
-This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
-Three different modelled datasets are used, and their capacity factors are compared to observed data to quantify their overall skill.
-Then, using a RES drought definition with a fixed threshold, their skill at representing these events is assessed.
-These modelled capacity factors allow for an expansion over a 45-year period, allowing for an analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
-Results show the importance of model selection for RES drought risk assessment.
-On a similar note, the addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, with a more balanced energy mix minimising seasonal drought patterns.
+\noindent The dependence of renewable energy sources (RES) such as wind and photovoltaic (PV) systems on the weather poses a critical challenge for energy systems.
+This study investigates the impact of targeting a balanced distribution of wind and PV capacity on reducing periods of low renewable generation, known as RES droughts.
+Three different RES models are used to estimate the capacity factors for different installed capacities of wind and PV.
+The skill of the RES models is quantified by comparing capacity factor time series to observed data.
+Their skill at representing RES droughts is also quantified.
+The RES models are used to generate a 45-year hourly time series of RES generation, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
+Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
+The addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, as well as reducing seasonal drought patterns.
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
-\textcolor{magenta}{I rewrote this and I still do not like it}
 
-% As the share of renewable energy sources (RES) like wind and photovoltaic (PV) power grows, their dependence on weather poses a critical challenge for energy systems.
-This study investigates periods of low renewable generation, known as RES droughts, focusing on Ireland's current wind-dominated energy mix and a projected scenario with a more balanced distribution of wind and PV capacity.
-Using three datasets RES droughts are analysed over 45 years to assess their frequency, duration, and return periods.
-The methodology involves comparing modelled capacity factors against observed data to evaluate the accuracy of each dataset and applying a fixed threshold to identify RES drought events.
-Results show that adding PV capacity significantly reduces the frequency and duration of RES droughts, with a more balanced energy mix minimising seasonal drought patterns.
-These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
-\\
-
-\textbf{Keywords}: Energy Droughts, Wind Power, Solar PV Power, Renewable Energy Sources, Seasonality, Return Periods, 
+\textbf{Keywords}: RES Drought, Wind Power, Solar PV Power, Renewable Energy Sources, Return Periods
 \end{abstract}
@@
 
-As the global shift towards renewable energy accelerates, the European Union (EU) has emerged as a leader in climate and energy policy.
-Under the revised European Green Deal \cite{greendeal2022report}, the EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{eurostat2023share}.
+The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{eurostat2023share}.
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
@@
 
-\begin{comment}
-This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios.
-The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected future energy scenario, featuring a more balanced share between wind and PV.
-A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, duration and seasonality are presented.
-\end{comment}
-
-This study focuses on Ireland, a region with a strong reliance on wind energy and ambitious targets for PV expansion.
-This makes it the perfect case study to provide valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
-This work also compares the performance of different energy models, revealing how modelling decisions impact the representation of RES droughts.
-The combination of these two offers insights about the effect of system planning decisions on RES droughts and the responses required to minimise their impact.
-
-\begin{comment}
-This study focuses on Ireland, a region with a strong reliance on wind energy and ambitious targets for PV expansion, to provide valuable insights into the potential benefits of diversifying the renewable energy mix.
-This work compares the performance of energy models and highlights the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
-\end{comment}
+This study focuses on Ireland, a region with a strong reliance on wind energy, which has ambitious targets for PV expansion.
+This case study provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
+The performance of different RES models are compared, and a 45-year time series of RES generation is produced.
+The results highlight the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
 
@@
 Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\cite{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
-In this paper, the focus is exclusively on energy generation, and a fixed threshold approach to define RES droughts is used.
+In this paper, the focus is exclusively on energy generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
 
@@
 
-The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their sources, characteristics, and relevance for evaluating RES droughts.
-Section~\ref{sec:Methods} outlines the energy models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:Results}, the models are first verified against actual energy data to assess their accuracy, followed by an analysis of RES drought occurrences across the two energy scenarios, highlighting key differences across the datasets and the effects of varying energy capacities.
+The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their characteristics and relevance for evaluating RES droughts.
+Section~\ref{sec:Methods} outlines the RES models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
+In section~\ref{sec:Results}, the models are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
 Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
@@
 The spreadsheet available from the EirGrid website contains two key variables: generation and availability.
-Generation is the energy that a RES farm actually contributed to the grid, which may include limitations introduced by the TSO to maintain grid stability, most commonly in the form of curtailment.
+Generation is the energy that a RES farm actually contributed to the grid, which may include limitations introduced by the TSO to maintain grid stability, such as constraints and curtailment.
 Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-This is then combined with installed capacity to compute the CF, the main variable for assessing renewable energy output.
 Generation and availability values are available from 2014 onward for wind and from 2018 onward for PV, although PV only became present in the Republic of Ireland data in 2023.
-
-The PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed PV capacity.
-To adjust for actual PV generation potential, installed capacities were estimated to be 1.4 times the MEC values.
-This installed capacity was used for PV generation, with farm output capped to the MEC value.
+Availability values are used to calculate the skill of the RES models.
 
@@
 Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
-This global gridded dataset provides hourly atmospheric variables from 1940 to the present.
+This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree, which is approximately every 27\,km.
 It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
@@
 
-This study relies on three datasets to analyse RES droughts across the island of Ireland.
+This study uses three datasets to analyse RES droughts across the island of Ireland.
 Data downloaded from C3S-E were used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
@@
 For national-level analyses, the aggregated CF time series provided by C3S-E were used at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
-We computed a weighted average of these, based on the installed capacity of each one, to represent the total CF of Ireland.
-This CF time series for Ireland is based on the underlying C3S-E assumption of RES generation at every 0.25\textdegree grid point in Ireland.
+These are based on the assumption by C3S-E that RES generation occurs at every ERA5 grid point in Ireland.
+We computed a weighted average of these, based on the installed capacity of each one, to represent the total CF for Ireland.
 
@@
 
-For grid-level analyses, the gridded dataset from C3S-E was used to create a representative time series for Ireland's RES farms.
-A comprehensive list of RES farms, including each farm’s GPS coordinates and installed capacity, was compiled from the EirGrid website.
+The gridded dataset from C3S-E was used to create CF datasets which account for the location of RES farms in Ireland.
+A list of the RES farms in Ireland was compiled, including each farm’s latitude, longitude and installed capacity.
 Using these coordinates, the nearest grid point on the C3S-E grid was identified for each farm.
-The CF values from the C3S-E dataset corresponding to these grid points were retrieved for each farm.
-To construct the final CF time series, a weighted average of the CF values was calculated, with the installed capacity of each farm serving as the weight.
-This process resulted in a single time series for each energy source (wind and PV), providing an accurate representation of RES generation in Ireland at a grid level.
+The CF values from the C3S-E dataset corresponding to these grid points were retrieved.
+A weighted average of the CF values was calculated, with the installed capacity of each farm serving as the weight, to construct the CF time series for Ireland.
+This process resulted in a time series of RES generation for each energy source (wind and PV) for Ireland, which takes the location of the RES farms into account.
 
@@
 
-Atlite transforms weather data into energy data, using the wind speed at 100 metres ($u_{100}$, $v_{100}$), and the locations of existing RES farms as described in C3S-E G.
+Atlite transforms weather data into energy data using the gridded ERA5 data and the locations of existing RES farms, as described in C3S-E G.
+ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate PV generation.
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
-This study uses a wind turbine power curve generated by Renewables.ninja: Enercon E112.4500 with 0.30w smoothing~\cite{staffell2016wake}.
-For PV power estimation, Atlite uses four irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) as well as the air temperature ($t2m$).
-A single PV model~\cite{beyer2004pv}, with the Kaneka Hybrid panel option, is used.
-The azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
+This study identifes the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\cite{staffell2016wake}.
 The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
@@
 In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
-Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
+Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
 These values are based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
 New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
-For wind, the CF time series remains unchanged, as significant shifts in wind farm distribution are not expected.
+For wind, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
 The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
@@
 In this study, a RES drought event was defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-The choice of this threshold is somewhat arbitrary but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought}.
+The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought}.
 By using a 24-hour moving average, fewer but longer-lasting events were captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
@@
 The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period (Fig.~\ref{fig:find_res_droughts}).
-This means that a single hour above the threshold does not "break" a drought event if it is surrounded by prolonged low-generation hours.
+This means that a single hour above the threshold does not``break" a drought event if it is surrounded by prolonged low-generation hours.
 As a result, fewer but longer-lasting drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
@@
 Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
+For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
+
+The PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed PV capacity.
+To enable actual PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
+This installed capacity was used for PV generation, with farm output capped to the MEC value.
 
@@
 
-This study highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts.
+This study highlighted the key differences between three RES models: Atlite, C3S-E G, and C3S-E N, in their ability to represent RES droughts.
 One of the most significant differences is how each dataset incorporates the specific locations of RES farms.
@@
 \section{Acknowledgments}
+\begin{comment}
 Using the Climate Data Store. Thank EirGrid for the data and add a reference to the funding project or projects
+\end{comment}
 

### 2024-11-29 Article/main.tex
CS edit complete

@@
 	\includegraphics[width=\textwidth]{verification_wind_bar}
-	\caption{Average annual number of wind drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
+	\caption{Average annual number of wind drought events for EirGrid (black outline), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
 	\label{fig:bar_number_events_verification_wind}
@@
 Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to EirGrid, especially for high CF values.
-The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
-Atlite shows a slightly lower correlation and a higher RMSE, indicating larger deviations.
-All models show a slight positive bias, with Atlite exhibiting the most overestimation.
+The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E~G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
+All models show a slight positive bias, with Atlite exhibiting a slightly lower correlation and higher RMSE.
 These results reflect the broader challenges in accurately modelling PV, where variability is hard to capture.
-Nonetheless, C3S-E G’s lower RMSE suggests it may be more reliable for PV CF modelling in this context.
 
@@
 	\includegraphics[width=\textwidth]{verification_pv_bar}
-	\caption{Annual number of PV drought events for EirGrid (black), Atlite (red), C3S-E G (blue), and C3S-E N (purple) in 2023}
+	\caption{Annual number of PV drought events for EirGrid (black outline), Atlite (red), C3S-E G (blue), and C3S-E N (purple) in 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 The first part of the analysis examines the annual number of RES drought events across the three datasets.
-
 For wind (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting more than seven days.
@@
 In contrast, the balanced scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
-This reduction is attributed to the anti-correlation between wind and PV production throughout the year, a phenomenon well-documented in the literature \cite{kaspar2019drought}.
+This reduction is attributed to the anti-correlation between wind and PV production.
 
@@
 The RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
-A return period refers to the estimated average interval between occurrences of events with a specified duration or intensity.
+A return period refers to the estimated expected interval between occurrences of events with a specified duration or intensity.
 Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
@@
 This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months.
-The difference between Atlite and the C3S-E datasets arises from how each model handles PV output near the threshold of 0.1 CF.
+The difference between Atlite and the C3S-E results arises from differences in the datasets near the threshold of 0.1 CF.
 Atlite remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
@@
 In this study, a threshold CF value of 0.1 was used to identify RES droughts.
-To assess how sensitive the results are to this threshold, the annual number of events was calculated across a range of CF threshold values (Fig.~\ref{fig:number_days_threshold}).
-
-Across all four panels, the results reveal a general trend: as the threshold increases, so does the number of identified events.
-This increase, however, is not uniform across datasets and energy sources, but the overall pattern remains consistent.
+To assess how sensitive the results are to this threshold, the annual number of RES drought events was calculated for a range of CF threshold values (Fig.~\ref{fig:number_days_threshold}).
+The results reveal a general trend: as the threshold increases, so too does the number of identified events.
+This increase is not uniform across datasets and energy sources, but the overall pattern remains consistent.
 When the threshold increases, smaller events merge into larger, less frequent events.
@@
 
-This study highlighted the key differences between three RES models: Atlite, C3S-E G, and C3S-E N, in their ability to represent RES droughts.
-One of the most significant differences is how each dataset incorporates the specific locations of RES farms.
+This study has investigated the ability of three RES models to represent RES droughts: Atlite, C3S-E G, and C3S-E N.
+One of the most evident differences is how each dataset incorporates the specific locations of RES farms.
 Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
-While this approach improves PV drought models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with EirGrid data for wind generation estimates.
-This finding suggests that although the inclusion of RES farm locations is beneficial, particularly for PV, the accuracy of the power curve model in Atlite may be the critical factor in producing realistic time series for low-generation events.
+While this approach slightly improves PV models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with EirGrid data for wind generation estimates.
+This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES model is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
 
-The results showed that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
-In contrast, Atlite stands out for its closer alignment with EirGrid data, particularly in wind generation, as seen in the verification process (Section \ref{sec:wind_verification}).
+Atlite shows the best alignment with EirGrid data for wind generation.
+Differences between the models are smaller for PV, with C3S-G performing marginally better than the other two.
+The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
 This distinction was also evident in the analysis, where Atlite reported higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
-One of the study's most critical findings is that the accuracy of RES drought modelling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
-Atlite’s superior performance highlights the importance of selecting accurate models for assessing RES drought risks.
+Again, the results from RES drought modelling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
+Atlite’s superior performance highlights the importance of selecting validated models for assessing RES drought risks.
 This careful model selection can better quantify risks, support effective planning, and avoid the potential underestimation of capacity needs, which is essential for ensuring energy security.
@@
 Looking at the balanced scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
-Wind and PV together perform better in terms of drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
+Wind and PV together perform better in terms of reducing drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
 This diversification reduces the seasonal impact on RES droughts, as PV generation peaks in the summer and wind generation is more consistent in winter.
@@
 
-In terms of future research, there are two main topics to explore.
-First, future climate data could be integrated with energy scenarios to better understand how climate change might affect RES droughts.
-Second, expanding the scope of the study to include Europe would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
-However, this expansion would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.
+Future work is planned to extend the current analysis.
+First, climate projection data will be integrated with energy scenarios to better understand how climate change might affect RES droughts.
+Second, expanding the geographic domain of the study to include the EU would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
+This would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.
 

### 2024-12-04 Article/main.tex
Update Article

@@
-\documentclass[a4paper, 11pt]{article}
+\documentclass[a4paper, 11p1t]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
-\graphicspath{ {../Figures/} }
+\graphicspath{ {../Scripts/} }
 
@@
 	\centering
-	\includegraphics{droughts_methodology}
+	\includegraphics{droughts_methodology.pdf}
 	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021.
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification_power_curve}
-	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b/ Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid (shaded)}
+	\includegraphics{verification_power_curve.pdf}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid (shaded)}
 	\label{fig:power_curve}
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification_wind_contour}
+	\includegraphics{verification_wind_contour.png}
 	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification_wind_bar}
-	\caption{Average annual number of wind drought events for EirGrid (black outline), Atlite (red), C3S-E G (blue), and C3S-E N (purple)}
+	\includegraphics{verification_wind_number_events.pdf}
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and EirGrid (black outline)}
 	\label{fig:bar_number_events_verification_wind}
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification_pv_contour}
+	\includegraphics{verification_pv_contour.png}
 	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}	
@@
 	\centering
-	\includegraphics[width=\textwidth]{verification_pv_bar}
-	\caption{Annual number of PV drought events for EirGrid (black outline), Atlite (red), C3S-E G (blue), and C3S-E N (purple) in 2023}
+	\includegraphics{verification_pv_number_events.pdf}
+	\caption{Annual number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and EirGrid (black outline) in 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 	\centering
-	\includegraphics[width=\textwidth]{droughts_number_events}
+	\includegraphics{droughts_number_events.pdf}
 	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
@@
 
+\begin{figure}[!ht]
+	\centering
+	\includegraphics{droughts_return_periods.pdf}
+	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
+The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
+	\label{fig:return_periods}
+\end{figure}
+
 Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
@@
 Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_return_periods}
-	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
-The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
-The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset.}
-	\label{fig:return_periods}
-\end{figure}
 
@@
 
+\begin{figure}[!ht]
+	\centering
+	\includegraphics{droughts_seasonality.pdf}
+	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
+Lines correspond to the median values and the area between the first and third quartiles is shaded.}
+	\label{fig:res_droughts_seasonality}
+\end{figure}
+
 Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the wind panel (Fig.\ref{fig:res_droughts_seasonality}a).
@@
 
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_seasonality}
-	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
-The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
-Lines correspond to the median values and the area between the first and third quartiles is shaded.}
-	\label{fig:res_droughts_seasonality}
-\end{figure}
-
-\newpage
+\begin{comment}
 \subsubsection{Change of Threshold}
@@
 	\includegraphics[width=\textwidth]{droughts_varying_threshold}
-	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
-The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment.
+	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid) The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment.
 The y-axis indicates the annual number of events.
@@
 \end{figure}
+\end{comment}
 

### 2024-12-09 Article/main.tex
Update Article

@@
 The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{eurostat2023share}.
-While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources like wind and photovoltaic (PV) power.
+While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
 This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\cite{bloomfield2021, bloomfield2016, vanderWiel2019drought}.
-Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to grid stability and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
+Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
 
@@
 
-The C3S-E dataset estimates wind energy using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 m.
+The C3S-E dataset estimates wind energy using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
 This choice is based on expert advice and the trend in wind turbine installation.
@@
 Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
-These values are based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}.
+These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
 New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
@@
 The Atlite model allows the user to select certain PV panel characteristics.
-In this study, the three PV panel types available in the Atlite model were considered.
+In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kanena).
 Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
+Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kanena Hybrid panel option.
 For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
@@
 	\includegraphics{verification_pv_number_events.pdf}
-	\caption{Annual number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and EirGrid (black outline) in 2023}
+	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and EirGrid (black outline) during 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
-This is due to the typically lower CF of PV compared to wind, especially in a region like Ireland where solar potential is limited.
+This is due to the typically lower CF of PV compared to wind, especially in a region such as Ireland where solar potential is limited.
 PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of RES droughts in PV compared to wind.
@@
 For wind (Fig.~\ref{fig:res_droughts_seasonality}a), RES drought percentages are higher in summer than in winter.
-In the Atlite dataset, for instance, an average of 24\% of hours in summer (JJA) are identified as RES droughts, compared to only 4\% in winter (DJF).
+In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as RES droughts, compared to only 4\% in winter (December-January-February).
 This seasonal variation is influenced by the wind power curve model used to estimate CF, where the shape of the curve in lower wind speed regions (3-10 m/s) leads to significant differences in CF under low wind conditions.
@@
 
-\newpage
-\section*{Appendix}
-
 \end{document}

### 2024-12-10 Article/main.tex
Update Article

@@
 
-RES droughts are identified using wind and PV CF time series.
+RES droughts are identified using onshore wind and PV CF time series.
 In this study, three different datasets are used, all of which are driven by ERA5 data~\cite{hersbach2020era5}.
@@
 This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
-The primary data sources include: EirGrid, the transmission system operator (TSO) for the Republic of Ireland; the ERA5 reanalysis dataset; and the C3S-E datasets.
+The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S-E datasets.
 
@@
 Atlite transforms weather data into energy data using the gridded ERA5 data and the locations of existing RES farms, as described in C3S-E G.
-ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 irradiance variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate PV generation.
+ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate PV generation.
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
@@
 These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
+\textcolor{cyan}{Add that wind offshore are not considered in this study, refer to SEAI 2024 to explain why}.
 New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
@@
 	\includegraphics{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid (shaded)}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid availability (shaded)}
 	\label{fig:power_curve}
@@
 	\includegraphics{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}
+	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}
 	\label{fig:wind_verification_contour}
@@
 	\includegraphics{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and EirGrid (black outline)}
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and EirGrid (black outline).
+The time series are obtained with a time series of the capacity from 2014 to 2023}
 	\label{fig:bar_number_events_verification_wind}
@@
 The Atlite model allows the user to select certain PV panel characteristics.
-In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kanena).
+In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kaneka).
 Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kanena Hybrid panel option.
+Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
 For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
@@
 	\includegraphics{verification_pv_contour.png}
-	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left panel), C3S-E G (central panel) and C3S-E N (right panel) models}	
+	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}	
 	\label{fig:solar_verification_contour}
@@
 All models show a slight positive bias, with Atlite exhibiting a slightly lower correlation and higher RMSE.
-These results reflect the broader challenges in accurately modelling PV, where variability is hard to capture.
+These results reflect the broader challenges in accurately modelling PV, where hourly variability is hard to capture.
 
@@
 	\includegraphics{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month identified as a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.

### 2024-12-10 Article/main.tex
Update plots in Article

@@
 
-\subsection{Wind and PV Generation \& Capacity}
+\subsection{Wind and PV Capacity \& Availability}
 \label{sec:eirgrid}
@@
 Generation and availability values are available from 2014 onward for wind and from 2018 onward for PV, although PV only became present in the Republic of Ireland data in 2023.
-Availability values are used to calculate the skill of the RES models.
+This study focuses on availability for all analyses.
 
@@
 	\centering
+	\caption{ERA5 variables used to calculate wind and PV generation.}
 	\begin{tabular}{|l|c|}
@@
 	\end{tabular}
-	\caption{ERA5 variables used to calculate wind and PV generation.}
 	\label{tab:var_name}
@@
 	\includegraphics{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and EirGrid availability (shaded)}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and Observed (shaded)}
 	\label{fig:power_curve}
@@
 	\includegraphics{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed EirGrid CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}
 	\label{fig:wind_verification_contour}
@@
 The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
-The two C3S-E datasets are shown to overestimate the EirGrid CF, whereas the Atlite model is in good agreement with the EirGrid data.
+The two C3S-E datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
 The skill scores presented in Table~\ref{tab:wind_skill_scores} show that Atlite performs better than the C3S-E datasets for all of the skill scores.
@@
 	\end{tabular}
-	\caption{Skill scores for wind for the three datasets compared to EirGrid data}
+	\caption{Skill scores for wind for the three datasets compared to observed data}
 	\label{tab:wind_skill_scores}
@@
 	\includegraphics{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and EirGrid (black outline).
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and observed (black outline).
 The time series are obtained with a time series of the capacity from 2014 to 2023}
@@
 	\includegraphics{verification_pv_contour.png}
-	\caption{PV 2D density plot of the observed EirGrid (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}	
+	\caption{PV 2D density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (a), C3S-E G (b) and C3S-E N (c) models}	
 	\label{fig:solar_verification_contour}
@@
 
-Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to EirGrid, especially for high CF values.
+Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
 The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E~G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
@@
 	\end{tabular}
-	\caption{Skill scores for wind CF for the three datasets compared to EirGrid data}
+	\caption{Skill scores for wind CF for the three datasets compared to observed data}
 	\label{tab:pv_skill_scores}
@@
 Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of PV drought events in 2023 across different duration ranges.
-The figure reveals partial agreement between the three datasets and the EirGrid data, with consistent results observed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
-However, discrepancies appear in the other ranges, where the models diverge from the EirGrid data.
+The figure reveals partial agreement between the three datasets and the observed data, with consistent results observed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
+However, discrepancies appear in the other ranges, where the models diverge from the observed data.
 The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, leading to uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
@@
 	\includegraphics{verification_pv_number_events.pdf}
-	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and EirGrid (black outline) during 2023}
+	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and observed (black outline) during 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 The balanced mix of wind and PV in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
-By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns.
+By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns in Ireland.
 
@@
 Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
-While this approach slightly improves PV models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with EirGrid data for wind generation estimates.
+While this approach slightly improves PV models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
 This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES model is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
 
-Atlite shows the best alignment with EirGrid data for wind generation.
+Atlite shows the best alignment with observed data for wind generation.
 Differences between the models are smaller for PV, with C3S-G performing marginally better than the other two.

### 2024-12-12 Article/main.tex
Changed Figure Size

@@
 
-In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV).
-Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW.
+In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind and 0.6 GW for PV).
+Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
 These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
-\textcolor{cyan}{Add that wind offshore are not considered in this study, refer to SEAI 2024 to explain why}.
+This study does not include offshore wind in the analysis.
+Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenario~\cite{seai2024future}.
+
 New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
-The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).
-
+The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (11.45 GW for wind and 8.6 GW for PV).
 In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three models for each source) in the existing scenario, and an additional six time series that include the combined CF for both existing and balanced scenarios across the different models.
 
-This approach allows for a comparative analysis between a wind-dominated system (the existing scenario) and a more evenly distributed system (the balanced scenario), assessing how the balance of RES capacity affects the occurrence of RES droughts.
+It is important to note that the specific capacity values used in this study are illustrative and are not intended to reflect precise future realities.
+Instead, they serve to explore the impact of transitioning from a wind-dominated system (existing scenario) to a more evenly distributed system (balanced scenario).
+This approach allows for a comparative analysis between the two scenarios, assessing how the balance of RES capacity affects the occurrence of RES droughts.
 
@@
 	\includegraphics{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and observed (black outline).
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline).
 The time series are obtained with a time series of the capacity from 2014 to 2023}
@@
 To enable actual PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
-This installed capacity was used for PV generation, with farm output capped to the MEC value.
+This scaling factor was estimated by analysing proprietary data from individual PV farms provided by EirGrid, which showed that the installed capacities of many farms exceed their MEC values by approximately 40\%.
 
@@
 All models show a slight positive bias, with Atlite exhibiting a slightly lower correlation and higher RMSE.
-These results reflect the broader challenges in accurately modelling PV, where hourly variability is hard to capture.
 
@@
 	\includegraphics{verification_pv_number_events.pdf}
-	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and observed (black outline) during 2023}
+	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline) during 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 89\% of wind capacity (6.7 GW) and 11\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 53\% (9 GW) and PV capacity increases to 47\% (8 GW).
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 91\% of wind capacity (5.8 GW) and 9\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 57\% (11.45 GW) and PV capacity increases to 43\% (8.6 GW).
 Both scenarios were driven by 45 years of ERA5 data.
@@
 
-The bottom panels show the combination of wind and PV under the two capacity scenarios.
+Fig.~\ref{fig:boxplot_number_events}a~and~b show the combination of wind and PV under the two capacity scenarios.
 In the existing RES installation scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
@@
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
-The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
+The horizontal dashed line marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
 	\label{fig:return_periods}
@@
 This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influence drought duration estimates.
-Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
+Additionally, in all four graphs, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
 
@@
 
-Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the wind panel (Fig.\ref{fig:res_droughts_seasonality}a).
+Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
 However, in this panel, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
@@
 Future work is planned to extend the current analysis.
-First, climate projection data will be integrated with energy scenarios to better understand how climate change might affect RES droughts.
+First, climate projection data will be integrated with different energy scenarios, including the apparition of offshore wind to better understand how climate change might affect RES droughts.
 Second, expanding the geographic domain of the study to include the EU would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.

### 2024-12-17 Article/main.tex
Update Article

@@
 %opening
-%\title{ \textcolor{blue}{Change Title} The Impact of Model Assumptions on RES Drought Analysis}
 \title{Reducing RES Droughts through the integration of wind and PV}
@@
 This study investigates the impact of targeting a balanced distribution of wind and PV capacity on reducing periods of low renewable generation, known as RES droughts.
-Three different RES models are used to estimate the capacity factors for different installed capacities of wind and PV.
+Three different RES models are used to estimate the capacity factors for different installed capacities of wind and PV energy.
 The skill of the RES models is quantified by comparing capacity factor time series to observed data.
@@
 
-This study focuses on Ireland, a region with a strong reliance on wind energy, which has ambitious targets for PV expansion.
+This study focuses on Ireland, a region with a strong reliance on wind power, which has ambitious targets for PV power expansion.
 This case study provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
@@
 
-\subsection{Wind and PV Capacity \& Availability}
+\subsection{Wind and PV Capacity Factor}
 \label{sec:eirgrid}
@@
 Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-Generation and availability values are available from 2014 onward for wind and from 2018 onward for PV, although PV only became present in the Republic of Ireland data in 2023.
+Generation and availability values are available from 2014 onward for wind power and from 2018 onward for PV power, although PV only became present in the Republic of Ireland data in 2023.
 This study focuses on availability for all analyses.
@@
 	\centering
-	\caption{ERA5 variables used to calculate wind and PV generation.}
+	\caption{ERA5 variables used to calculate wind and PV generation}
 	\begin{tabular}{|l|c|}
@@
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
-This study identifes the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\cite{staffell2016wake}.
+This study identifies the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\cite{staffell2016wake}.
 The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
@@
 
-In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind and 0.6 GW for PV).
-Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
+In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for PV power).
+Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV capacities might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
 These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
@@
 New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
-For wind, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
+For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
 The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (11.45 GW for wind and 8.6 GW for PV).
@@
 	\includegraphics{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the Atlite  (left), C3S-E G (centre) and C3S-E N (right) models}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N models}
 	\label{fig:wind_verification_contour}
@@
 	\end{tabular}
-	\caption{Skill scores for wind for the three datasets compared to observed data}
+	\caption{Skill scores for wind power for the three datasets compared to observed data}
 	\label{tab:wind_skill_scores}
@@
 	\includegraphics{verification_pv_contour.png}
-	\caption{PV 2D density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the Atlite  (a), C3S-E G (b) and C3S-E N (c) models}	
+	\caption{PV 2D density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
 	\label{fig:solar_verification_contour}
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 91\% of wind capacity (5.8 GW) and 9\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 57\% (11.45 GW) and PV capacity increases to 43\% (8.6 GW).
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 91\% of wind capacity (5.9 GW) and 9\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 57\% (11.45 GW) and PV capacity increases to 43\% (8.6 GW).
 Both scenarios were driven by 45 years of ERA5 data.
@@
 The first part of the analysis examines the annual number of RES drought events across the three datasets.
-For wind (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting more than seven days.
-For PV (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
+For Fig.~\ref{fig:boxplot_number_events}a, the number of events decreases as the duration range increases, with very few events lasting more than seven days.
+For Fig.~\ref{fig:boxplot_number_events}b, the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
 This increase is due to extended low-generation periods occurring from November to March, depending on the dataset.
-When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a \& Fig.~\ref{fig:boxplot_number_events}b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
-This is due to the typically lower CF of PV compared to wind, especially in a region such as Ireland where solar potential is limited.
+When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
+This is due to the typically lower CF of PV power compared to wind power, especially in a region such as Ireland where solar potential is limited.
 PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of RES droughts in PV compared to wind.
 
-Fig.~\ref{fig:boxplot_number_events}a~and~b show the combination of wind and PV under the two capacity scenarios.
+Fig.~\ref{fig:boxplot_number_events}a~\&~b show the combination of wind and PV under the two capacity scenarios.
 In the existing RES installation scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
 In contrast, the balanced scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
-This reduction is attributed to the anti-correlation between wind and PV production.
+This reduction is attributed to the anti-correlation between wind and PV generation.
 
@@
 
-For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), due to the low levels of installed PV.
+For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of Fig.\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
 In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
 For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the existing scenario, to four years for the balanced scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
-% The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one.
-This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics.
-Notably, there is a greater difference in how each model represents wind generation than PV, meaning the balanced scenario reduces these wind-related discrepancies, bringing the datasets closer together.
 
@@
 
-For wind (Fig.~\ref{fig:res_droughts_seasonality}a), RES drought percentages are higher in summer than in winter.
+For Fig.~\ref{fig:res_droughts_seasonality}a, RES drought percentages are higher in summer than in winter.
 In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as RES droughts, compared to only 4\% in winter (December-January-February).
 This seasonal variation is influenced by the wind power curve model used to estimate CF, where the shape of the curve in lower wind speed regions (3-10 m/s) leads to significant differences in CF under low wind conditions.
-In contrast, the results for PV (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
+In contrast, the results for Fig.~\ref{fig:res_droughts_seasonality}b show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
 The Atlite results show a higher percentage of RES drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.
@@
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-The balanced mix of wind and PV in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
+The balanced mix of wind and PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
 By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns in Ireland.
@@
 Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
-While this approach slightly improves PV models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
+While this approach slightly improves PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
 This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES model is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.

### 2025-01-08 Article/main.tex
Update Article

@@
 
-Fig.~\ref{fig:bar_number_events_verification_wind} presents results for the average annual number of wind drought events during the 2014 to 2023 validation period.
+Fig.~\ref{fig:bar_number_events_verification_wind} presents the average annual number of wind drought events during the 2014 to 2023 validation period.
 The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
@@
 	\includegraphics{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline).
-The time series are obtained with a time series of the capacity from 2014 to 2023}
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
 	\label{fig:bar_number_events_verification_wind}
@@
 
-Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of PV drought events in 2023 across different duration ranges.
-The figure reveals partial agreement between the three datasets and the observed data, with consistent results observed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
+Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of PV drought events during the 2023 validation period across different duration ranges.
+The figure reveals partial agreement between the three datasets and the observed data, with consistent results noticed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
 However, discrepancies appear in the other ranges, where the models diverge from the observed data.
@@
 
+Nevertheless, the ultimate goal of this analysis is to assess the combination of wind and PV generation, where the complementary nature of these energy sources mitigates the limitations seen in PV-only results.
+
 \begin{figure}[!ht]
@@
 	\includegraphics{verification_pv_number_events.pdf}
-	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline) during 2023}
+	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline). The PV droughts are identified for 2023}
 	\label{fig:bar_number_events_verification_pv}
@@
 	\includegraphics{droughts_number_events.pdf}
-	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 	\includegraphics{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 	\includegraphics{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
@@
 Ireland currently has a highly wind-dependent energy system, but with ambitious targets for PV installations in the coming years, the energy mix is expected to approach a balance between wind and PV capacity.
-This balanced approach offers a more stable and secure energy supply by mitigating RES drought risks.
+While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and PV capacities may not optimize other aspects, such as annual energy production or meeting nighttime loads.
 For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.

### 2025-01-09 Article/main.tex
Update Article

@@
 Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-Generation and availability values are available from 2014 onward for wind power and from 2018 onward for PV power, although PV only became present in the Republic of Ireland data in 2023.
+Generation and availability values are available from 2014 onward for wind power and from 2018 onward for PV power, although PV availability data only became present in the Republic of Ireland in 2023.
 This study focuses on availability for all analyses.
@@
 Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
-This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree, which is approximately every 27\,km.
+This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
 It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
@@
 This choice is based on expert advice and the trend in wind turbine installation.
-The PV generation model used by C3S-E uses two ERA5 variables: Surface solar radiation downwards ($ssrd$) and air temperature ($t2m$).
+The PV generation model used by C3S-E uses two ERA5 variables: surface solar radiation downwards ($ssrd$) and air temperature ($t2m$).
 PV generation is calculated multiple times, using the same model with different azimuth and tilt angles.
@@
 Fig.~\ref{fig:bar_number_events_verification_wind} presents the average annual number of wind drought events during the 2014 to 2023 validation period.
-The figure shows that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
+The figure presents that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
 This pattern is particularly evident for shorter-duration events, which are the most frequent.
@@
 	\end{tabular}
-	\caption{Skill scores for wind CF for the three datasets compared to observed data}
+	\caption{Skill scores for PV CF for the three datasets compared to observed data}
 	\label{tab:pv_skill_scores}
@@
 
-Nevertheless, the ultimate goal of this analysis is to assess the combination of wind and PV generation, where the complementary nature of these energy sources mitigates the limitations seen in PV-only results.
+Nevertheless, the goal of this analysis is to assess the combination of wind and PV generation, where the complementary nature of these energy sources mitigates the limitations seen in PV-only results.
 
@@
 This increase is due to extended low-generation periods occurring from November to March, depending on the dataset.
-When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher or equal to those for wind, across all duration ranges and datasets.
+When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher than or equal to those for wind, across all duration ranges and datasets.
 This is due to the typically lower CF of PV power compared to wind power, especially in a region such as Ireland where solar potential is limited.
@@
 
-For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of Fig.\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
+For the existing installed capacity scenario (Fig.~\ref{fig:return_periods}c), the return periods mirror those of Fig.~\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
 In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
@@
 
-Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
-For instance, in the existing scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
+Across Fig.~\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
+For instance, in the existing scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
 This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influence drought duration estimates.
@@
 
-Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
-However, in this panel, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
+Similar to previous results, the existing installed capacity wind-PV combination (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
+However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
 This reduction can be explained by the contribution of PV generation during the summer months in the existing scenario, even though it constitutes only 11\% of total capacity.
@@
 This distinction was also evident in the analysis, where Atlite reported higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
-Again, the results from RES drought modelling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
+Again, the results from RES drought modelling rely more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
 Atlite’s superior performance highlights the importance of selecting validated models for assessing RES drought risks.
@@
 Ireland currently has a highly wind-dependent energy system, but with ambitious targets for PV installations in the coming years, the energy mix is expected to approach a balance between wind and PV capacity.
-While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and PV capacities may not optimize other aspects, such as annual energy production or meeting nighttime loads.
+While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and PV capacities may not optimise other aspects, such as annual energy production or meeting nighttime loads.
 For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.

### 2025-01-14 Article/main.tex
Update Article

@@
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
+\\
 
@@
 In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for PV power).
-Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV capacities might impact RES droughts, this study also considered a balanced scenario, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
+This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% PV capacity.
+Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
 These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
 This study does not include offshore wind in the analysis.
-Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenario~\cite{seai2024future}.
+Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenarios~\cite{seai2024future}.
 
@@
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
-The PV and wind time series were then combined using updated weights that reflect the installed capacities in the balanced scenario (11.45 GW for wind and 8.6 GW for PV).
-In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three models for each source) in the existing scenario, and an additional six time series that include the combined CF for both existing and balanced scenarios across the different models.
+In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three models for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different models.
 
 It is important to note that the specific capacity values used in this study are illustrative and are not intended to reflect precise future realities.
-Instead, they serve to explore the impact of transitioning from a wind-dominated system (existing scenario) to a more evenly distributed system (balanced scenario).
+Instead, they serve to explore the impact of transitioning from a wind-dominated system (91W-9PV) to a more evenly distributed system (57W-43PV).
 This approach allows for a comparative analysis between the two scenarios, assessing how the balance of RES capacity affects the occurrence of RES droughts.
@@
 The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period (Fig.~\ref{fig:find_res_droughts}).
-This means that a single hour above the threshold does not``break" a drought event if it is surrounded by prolonged low-generation hours.
+This means that a single hour above the threshold does not "break" a drought event if it is surrounded by prolonged low-generation hours.
 As a result, fewer but longer-lasting drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the existing scenario, with 91\% of wind capacity (5.9 GW) and 9\% of PV capacity (0.6 GW), and a balanced scenario, where wind capacity comprises 57\% (11.45 GW) and PV capacity increases to 43\% (8.6 GW).
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of PV capacity, and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and PV capacity increases to 8.6 GW.
 Both scenarios were driven by 45 years of ERA5 data.
@@
 Fig.~\ref{fig:boxplot_number_events}a~\&~b show the combination of wind and PV under the two capacity scenarios.
-In the existing RES installation scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
-In contrast, the balanced scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
+In the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
+In contrast, the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
 This reduction is attributed to the anti-correlation between wind and PV generation.
@@
 	\includegraphics{droughts_number_events.pdf}
-	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 
-For the existing installed capacity scenario (Fig.~\ref{fig:return_periods}c), the return periods mirror those of Fig.~\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
-In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
-For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the existing scenario, to four years for the balanced scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
+For the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the return periods mirror those of Fig.~\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
+In the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
+For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
 
@@
 	\includegraphics{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 Across Fig.~\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
-For instance, in the existing scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
+For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
 This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influence drought duration estimates.
@@
 	\includegraphics{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
@@
 
-Similar to previous results, the existing installed capacity wind-PV combination (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
+Similar to previous results, the 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of PV generation during the summer months in the existing scenario, even though it constitutes only 11\% of total capacity.
+This reduction can be explained by the contribution of PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
 Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
-In the balanced scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
+In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
 The balanced mix of wind and PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
-By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns in Ireland.
-
-\begin{comment}
-\subsubsection{Change of Threshold}
-
-In this study, a threshold CF value of 0.1 was used to identify RES droughts.
-To assess how sensitive the results are to this threshold, the annual number of RES drought events was calculated for a range of CF threshold values (Fig.~\ref{fig:number_days_threshold}).
-The results reveal a general trend: as the threshold increases, so too does the number of identified events.
-This increase is not uniform across datasets and energy sources, but the overall pattern remains consistent.
-When the threshold increases, smaller events merge into larger, less frequent events.
-Depending on the dataset, the number of events reaches a peak around or after a threshold of 0.2, and then gradually decreases until only a single event encompasses the entire time series.
-
-Overall, the four panels indicate that regardless of the threshold applied, the Atlite dataset consistently identifies more events than the two C3S-E datasets, except for PV when the threshold is below 0.08.
-This finding suggests that Atlite, likely due to its more accurate representation of wind generation in Ireland, captures more RES drought events that might be missed with different modelling assumptions in the C3S-E datasets.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_varying_threshold}
-	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid) The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment.
-The y-axis indicates the annual number of events.
-Each line correspond to the median value and the first and third quartiles are shown in shaded area}
-	\label{fig:number_days_threshold}
-\end{figure}
-\end{comment}
 
@@
 
-Looking at the balanced scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
+Looking at the 57W-43PV scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
 Wind and PV together perform better in terms of reducing drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
@@
 Future work is planned to extend the current analysis.
-First, climate projection data will be integrated with different energy scenarios, including the apparition of offshore wind to better understand how climate change might affect RES droughts.
+First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change might affect RES droughts.
 Second, expanding the geographic domain of the study to include the EU would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.

### 2025-01-15 Article2/elsarticle-template-harv.tex
Create new Article template for Preprint

--- 2025-01-14 Article/main.tex
@@
-\documentclass[a4paper, 11p1t]{article}
-\usepackage[utf8]{inputenc}
+%% 
+%% Copyright 2007-2025 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.3 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.3 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%% 
+%% Template article for Elsevier's document class `elsarticle'
+%% with harvard style bibliographic references
 
-\usepackage[breaklinks, colorlinks=true, citecolor=black, linkcolor=blue, urlcolor=blue, filecolor=blue]{hyperref}
+\documentclass[preprint, 12pt, authoryear]{elsarticle}
 
-\usepackage{comment}
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times,authoryear]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
 
-\usepackage[center, labelfont=bf]{caption}
-\captionsetup{justification=justified, singlelinecheck=false} % Justify left
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
 
-\usepackage{amsmath} % Math equation
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsmath package provides various useful equation environments.
+\usepackage{amsmath}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
 
-\usepackage{url}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
 
-% For table
-\usepackage{multirow}
-\usepackage[table, xcdraw]{xcolor}
-
-% Page definitions
-\setlength{\topmargin}{1.5cm}
-\setlength{\headheight}{1.1\baselineskip}
-\setlength{\headsep}{20pt}
-\setlength{\topskip}{12pt}
-\setlength{\evensidemargin}{0pt}
-\setlength{\oddsidemargin}{0pt}
-\setlength{\textheight}{240mm}
-\setlength{\textwidth}{160mm}
-\setlength{\voffset}{-2cm}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-\usepackage{natbib} % Bibliography
-\bibliographystyle{unsrt}
-
-\usepackage{graphicx}
 \graphicspath{ {../Scripts/} }
 
-%opening
-\title{Reducing RES Droughts through the integration of wind and PV}
-\author{Boris Morin, Damian Flynn, Aina Maimó Far, Conor Sweeney}
+%\usepackage{natbib}
+
+\journal{Renewable Energy}
 
@@
 
-\maketitle
+\begin{frontmatter}
 
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \affiliation for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%            addressline={}, 
+%%            city={},
+%%            postcode={}, 
+%%            state={},
+%%            country={}}
+%% \fntext[label3]{}
+
+\title{Reducing RES Droughts through the integration of wind and PV} %% Article title
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1, label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+\author[Math]{Boris Morin} %% Author name
+\author[Math]{Aina Maimo Far} %% Author name
+\author[Eng]{Damian Flynn} %% Author name
+\author[Math]{Conor Sweeney} %% Author name
+
+%% Author affiliation
+\affiliation[Math]{organization={University College Dublin, School of Mathematics and Statistics}, %Department and Organization
+            addressline={}, 
+            city={Dublin},
+            postcode={}, 
+            country={Ireland}}
+
+\affiliation[Eng]{organization={University College Dublin, School of Electrical and Electronic Engineering}, %Department and Organization
+        	addressline={}, 
+        	city={Dublin},
+        	postcode={}, 
+        	country={Ireland}}
+
+%% Abstract
 \begin{abstract}
-
-\noindent The dependence of renewable energy sources (RES) such as wind and photovoltaic (PV) systems on the weather poses a critical challenge for energy systems.
+The dependence of renewable energy sources (RES) such as wind and photovoltaic (PV) systems on the weather poses a critical challenge for energy systems.
 This study investigates the impact of targeting a balanced distribution of wind and PV capacity on reducing periods of low renewable generation, known as RES droughts.
@@
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
-\\
+\end{abstract}
 
-\textbf{Keywords}: RES Drought, Wind Power, Solar PV Power, Renewable Energy Sources, Return Periods
-\end{abstract}
+%%Research highlights
+\begin{highlights}
+\item Energy droughts are defined as periods of low production
+
+\item Three models compared to evaluate renewable energy droughts in Ireland
+
+\item Analysis of 45 years of ERA5 weather data for energy drought trends
+
+\item Diverse energy mix reduces drought frequency and duration significantly
+
+\item Model choice impacts energy drought results
+\end{highlights}
+
+%% Keywords
+\begin{keyword}
+RES Drought\sep Wind Power \sep Solar PV Power \sep Renewable Energy Sources \sep Return Periods
+
+\end{keyword}
+
+\end{frontmatter}
+
+%% Add \usepackage{lineno} before \begin{document} and uncomment 
+%% following line to enable line numbers
+\linenumbers
 
@@
 	\includegraphics{verification_pv_contour.png}
-	\caption{PV 2D density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
+	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
 	\label{fig:solar_verification_contour}
@@
 	\includegraphics{verification_pv_number_events.pdf}
-	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline). The PV droughts are identified for 2023}
+	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline).
+The PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
 	\label{fig:bar_number_events_verification_pv}
@@
 
-\newpage
-\section{Acknowledgments}
-\begin{comment}
-Using the Climate Data Store. Thank EirGrid for the data and add a reference to the funding project or projects
-\end{comment}
+%% For citations use: 
+%%       \citet{<label>} ==> Lamport (1994)
+%%       \citep{<label>} ==> (Lamport, 1994)
 
-\section{Funding information}
+%% If you have bib database file and want bibtex to generate the
+%% bibitems, please use
 
-\section{Author contribution}
-
+% \bibliographystyle{unsrt}
 \bibliography{litrev}
 
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
+
+%% Refer following link for more details about bibliography and citations.
+%% https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management
 \end{document}
+
+\endinput
+%%
+%% End of file `elsarticle-template-harv.tex'.

### 2025-01-15 Article2/elsarticle-template-harv.tex
Fix references

@@
 
-The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \cite{eurostat2023share}.
+The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \citep{eurostat2023share}.
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
-This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\cite{bloomfield2021, bloomfield2016, vanderWiel2019drought}.
+This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2021, bloomfield2016, vanderWiel2019drought}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
@@
 
-For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in other research~\cite{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
+For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in other research~\citep{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
 Alternative methods exist for defining RES droughts.
-One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable energy generation~\cite{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
-Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\cite{bracken2024drought, allen2023drought}.
-Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\cite{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
+One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable energy generation~\citep{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
+Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\citep{bracken2024drought, allen2023drought}.
+Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\citep{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
 In this paper, the focus is exclusively on energy generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
@@
 RES droughts are identified using onshore wind and PV CF time series.
-In this study, three different datasets are used, all of which are driven by ERA5 data~\cite{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3S-E)~\cite{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\cite{dubus2023energy}.
+In this study, three different datasets are used, all of which are driven by ERA5 data~\citep{hersbach2020era5}.
+Two of the datasets are part of C3S Energy (C3S-E)~\citep{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
 One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset was generated using the Atlite model~\cite{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\cite{hofman2021atlite} and is widely used for estimating wind and PV generation~\cite{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
+The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and is widely used for estimating wind and PV generation~\citep{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
 
@@
 
-EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid2023spreadsheet}.
+EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\citep{eirgrid2023spreadsheet}.
 These datasets include information such as each farm’s installed capacity, name, and connection date.
@@
 
-Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\cite{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
-It is widely used for estimating PV and wind energy~\cite{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
+It is widely used for estimating PV and wind energy~\citep{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
 Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.
@@
 
-The EU Copernicus Climate Change Service developed the C3S-E renewable energy dataset for Europe~\cite{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
+The EU Copernicus Climate Change Service developed the C3S-E renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
 This dataset provides hourly CF for wind and PV energy from 1979 to the present.
@@
 PV generation is calculated multiple times, using the same model with different azimuth and tilt angles.
-The results are aggregated based on a statistical distribution of the module angles based on the geographical location~\cite{saintdrenan2018solar}.
+The results are aggregated based on a statistical distribution of the module angles based on the geographical location~\citep{saintdrenan2018solar}.
 
@@
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
-This study identifies the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\cite{staffell2016wake}.
+This study identifies the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\citep{staffell2016wake}.
 The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
@@
 Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
-These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\cite{cap2024future}.
+These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\citep{cap2024future}.
 This study does not include offshore wind in the analysis.
-Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenarios~\cite{seai2024future}.
+Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenarios~\citep{seai2024future}.
 
@@
 In this study, a RES drought event was defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \cite{kaspar2019drought, ohba2022drought, mayer2023drought}.
+The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \citep{kaspar2019drought, ohba2022drought, mayer2023drought}.
 By using a 24-hour moving average, fewer but longer-lasting events were captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
@@
 The Atlite model allows the user to specify the power curve.
-We considered the 121 power curves available for download from Renewables.ninja~\cite{staffell2016wake}.
+We considered the 121 power curves available for download from Renewables.ninja~\citep{staffell2016wake}.
 For each power curve, Renewables.ninja also provides four associated smoothed power curves.
@@
 Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and area under the curve).
-Based on the best-performing metrics, the Breyer PV panel model was selected \cite{beyer2004pv}, using the Kaneka Hybrid panel option.
+Based on the best-performing metrics, the Breyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
 For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
@@
 
-% \bibliographystyle{unsrt}
+\bibliographystyle{elsarticle-harv}
 \bibliography{litrev}

### 2025-01-15 Article2/elsarticle-template-harv.tex
Add info for title page

@@
 
-\author[Math]{Boris Morin} %% Author name
+\author[Math]{Boris Morin \corref{cor1}} %% Author name
+\ead{boris.morin@ucdconnect.ie}
+
 \author[Math]{Aina Maimo Far} %% Author name
+\ead{aina.maimofar@ucd.ie}
+
 \author[Eng]{Damian Flynn} %% Author name
+\ead{damian.flynn@ucd.ie}
+
 \author[Math]{Conor Sweeney} %% Author name
+\ead{conor.sweeney@ucd.ie}
 
 %% Author affiliation
-\affiliation[Math]{organization={University College Dublin, School of Mathematics and Statistics}, %Department and Organization
-            addressline={}, 
+\affiliation[Math]{organization={School of Mathematics and Statistics, University College Dublin}, %Department and Organization
+	addressline={Belfied, Dublin 4}, 
 	city={Dublin},
-            postcode={}, 
+	postcode={D04 V1W8}, 
 	country={Ireland}}
 
-\affiliation[Eng]{organization={University College Dublin, School of Electrical and Electronic Engineering}, %Department and Organization
-        	addressline={}, 
+\affiliation[Eng]{organization={School of Electrical and Electronic Engineering, University College Dublin}, %Department and Organization
+	addressline={Belfied, Dublin 4}, 
 	city={Dublin},
-        	postcode={}, 
+	postcode={D04 V1W8}, 
 	country={Ireland}}
+
+\cortext[cor1]{Correponding author}
 

### 2025-01-15 Article2/elsarticle-template-harv.tex
Fix highlights

@@
 \begin{highlights}
-\item Energy droughts are defined as periods of low production
+\item The C3S-Energy dataset provides renewable energy capacity factor time series
 
-\item Three models compared to evaluate renewable energy droughts in Ireland
+\item Two C3S-Energy and one custom models are compared to assess RES droughts in Ireland
 
-\item Analysis of 45 years of ERA5 weather data for energy drought trends
+\item RES drought trends are explored by analysing 45 years of ERA5 weather data
 
-\item Diverse energy mix reduces drought frequency and duration significantly
+\item Diverse energy mix reduces RES drought frequency, duration and extremes
 
-\item Model choice impacts energy drought results
+\item Model selection is crucial in the accurate identification of RES droughts
+
 \end{highlights}

### 2025-01-15 Article2/elsarticle-template-harv.tex
Update highlights

@@
 
-\author[Math]{Aina Maimo Far} %% Author name
+\author[Math]{Aina Maimó Far} %% Author name
 \ead{aina.maimofar@ucd.ie}
@@
 
-\item Two C3S-Energy and one custom models are compared to assess RES droughts in Ireland
+\item RES droughts are analysed using 45 years of hourly wind and PV generation data
 
-\item RES drought trends are explored by analysing 45 years of ERA5 weather data
+\item RES droughts from C3S-Energy and ERA5-Atlite datasets are compared
 
-\item Diverse energy mix reduces RES drought frequency, duration and extremes
+\item Adding PV to a wind-dominated system reduces RES drought frequency and duration
 
-\item Model selection is crucial in the accurate identification of RES droughts
+\item Validated RES datasets are crucial to accurately identify RES drought extremes
 

### 2025-01-15 Article/RES_Droughts_Morin.tex
Update Files Names

--- 2025-01-14 Article/main.tex
@@
 	\includegraphics{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline).
+The wind droughts are identified from 2014 to 2023, considering the actual capacity of the system at any given time}
 	\label{fig:bar_number_events_verification_wind}

### 2025-01-15 Article2/RES_droughts_Morin_RE.tex
Update Files Names

--- 2025-01-15 Article2/elsarticle-template-harv.tex
@@
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
+%% The lineno packages adds line numbers.
 \usepackage{lineno}
@@
 \graphicspath{ {../Scripts/} }
-
-%\usepackage{natbib}
 
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%            addressline={}, 
-%%            city={},
-%%            postcode={}, 
-%%            state={},
-%%            country={}}
-%% \fntext[label3]{}
-
 \title{Reducing RES Droughts through the integration of wind and PV} %% Article title
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1, label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
 
@@
 
-\cortext[cor1]{Correponding author}
+\cortext[cor1]{Corresponding author}
 
@@
 \begin{highlights}
-\item The C3S-Energy dataset provides renewable energy capacity factor time series
-
 \item RES droughts are analysed using 45 years of hourly wind and PV generation data
@@
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
-This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2021, bloomfield2016, vanderWiel2019drought}.
+This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderWiel2019drought}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
@@
 	\centering
-	\includegraphics{droughts_methodology.pdf}
+	\includegraphics[width=\textwidth]{droughts_methodology.pdf}
 	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021.
@@
 	\centering
-	\includegraphics{verification_power_curve.pdf}
+	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
 	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and Observed (shaded)}
@@
 	\centering
-	\includegraphics{verification_wind_contour.png}
+	\includegraphics[width=\textwidth]{verification_wind_contour.png}
 	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N models}
@@
 	\centering
-	\includegraphics{verification_wind_number_events.pdf}
+	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
 	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
@@
 	\centering
-	\includegraphics{verification_pv_contour.png}
+	\includegraphics[width=\textwidth]{verification_pv_contour.png}
 	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
@@
 	\centering
-	\includegraphics{verification_pv_number_events.pdf}
+	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
 	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline).
@@
 	\centering
-	\includegraphics{droughts_number_events.pdf}
+	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
 	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
@@
 	\centering
-	\includegraphics{droughts_return_periods.pdf}
+	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
 	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
@@
 	\centering
-	\includegraphics{droughts_seasonality.pdf}
+	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
 	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
@@
 
-%% For citations use: 
-%%       \citet{<label>} ==> Lamport (1994)
-%%       \citep{<label>} ==> (Lamport, 1994)
-
 %% If you have bib database file and want bibtex to generate the
@@
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
 %% Refer following link for more details about bibliography and citations.
@@
 \endinput
-%%
-%% End of file `elsarticle-template-harv.tex'.

### 2025-01-16 Article/RES_droughts_Morin_ReEn.tex
New Article Template for Renewable Energy

--- 2025-01-15 Article2/RES_droughts_Morin_RE.tex
@@
 \usepackage{lineno}
+
+%% Use this package to add line brakes in URL
+\usepackage{xurl}
 
@@
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
-This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderWiel2019drought}.
+This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderwiel2019drought}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
@@
 
-%% If you have bib database file and want bibtex to generate the
-%% bibitems, please use
+\section*{Data Availability}
+
+The meteorological data, including wind speed, irradiance, and temperature, can be accessed through the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
+The C3S-E dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
+Information on wind and PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).
+The Atlite scripts used in this study are open-source and can be found on GitHub (\url{https://github.com/pypsa/atlite}).
+The data required to reproduce the figures in this article will be made available upon acceptance of the manuscript.
 
 \bibliographystyle{elsarticle-harv}
-\bibliography{litrev}
+\bibliography{RES_droughts_Morin_ref}
 

### 2025-01-16 Article/RES_droughts_Morin_ReEn.tex
Update article

@@
 \begin{abstract}
-The dependence of renewable energy sources (RES) such as wind and photovoltaic (PV) systems on the weather poses a critical challenge for energy systems.
-This study investigates the impact of targeting a balanced distribution of wind and PV capacity on reducing periods of low renewable generation, known as RES droughts.
+The increasing share of renewable energy sources (RES) combined with their dependence on weather poses a critical challenge for energy systems.
+This study investigates the impact of targeting a balanced distribution of wind and photovoltaic (PV) capacity on reducing periods of low renewable generation, known as RES droughts.
 Three different RES models are used to estimate the capacity factors for different installed capacities of wind and PV energy.
-The skill of the RES models is quantified by comparing capacity factor time series to observed data.
-Their skill at representing RES droughts is also quantified.
-The RES models are used to generate a 45-year hourly time series of RES generation, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
+The skill of the RES models is quantified by comparing capacity factor time series to observed data and by assessing their representation of observed RES droughts.
+The RES models are used to generate a 45-year hourly time series of RES capacity factor, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
 Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
-The addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, as well as reducing seasonal drought patterns.
+The addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, as well as reducing extremes and seasonal drought patterns.
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
@@
 In this study, three different datasets are used, all of which are driven by ERA5 data~\citep{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3S-E)~\citep{cds2023energy}, an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
+Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy, cds2023energy}.
 One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
@@
 
-\subsection{Wind and PV Capacity Factor}
+\subsection{Wind and PV Capacity and Availability}
 \label{sec:eirgrid}
@@
 	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and Observed (shaded)}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (blue) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (blue) and Observed (shaded)}
 	\label{fig:power_curve}
@@
 
-Fig.~\ref{fig:bar_number_events_verification_wind} presents the average annual number of wind drought events during the 2014 to 2023 validation period.
-The figure presents that Atlite shows the best agreement overall with the observed frequency and duration of wind drought events.
+Fig.~\ref{fig:bar_number_events_verification_wind} shows the average annual number of wind drought events during the 2014 to 2023 validation period.
+The figure reveals that Atlite presents the best overall agreement with the observed frequency and duration of wind drought events.
 This pattern is particularly evident for shorter-duration events, which are the most frequent.
@@
 	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
+	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline).
+The wind droughts are identified from 2014 to 2023, considering the actual capacity of the system at any given time}
 	\label{fig:bar_number_events_verification_wind}
@@
 To enable actual PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
-This scaling factor was estimated by analysing proprietary data from individual PV farms provided by EirGrid, which showed that the installed capacities of many farms exceed their MEC values by approximately 40\%.
+This scaling factor was estimated by analysing proprietary data from individual PV farms provided by EirGrid, which showed that, on average, assuming that the installed capacities of farms exceed their MEC values by 40\% yields the best agreement with the observed availability.
 
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of PV capacity, and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and PV capacity increases to 8.6 GW.
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and PV capacity increases to 8.6 GW.
 Both scenarios were driven by 45 years of ERA5 data.
@@
 The first part of the analysis examines the annual number of RES drought events across the three datasets.
-For Fig.~\ref{fig:boxplot_number_events}a, the number of events decreases as the duration range increases, with very few events lasting more than seven days.
-For Fig.~\ref{fig:boxplot_number_events}b, the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
+When only wind energy is considered (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting more than seven days.
+In the case of only PV energy (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
 This increase is due to extended low-generation periods occurring from November to March, depending on the dataset.
@@
 This is due to the typically lower CF of PV power compared to wind power, especially in a region such as Ireland where solar potential is limited.
-PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of RES droughts in PV compared to wind.
+PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of drought events in PV compared to wind.
 
-Fig.~\ref{fig:boxplot_number_events}a~\&~b show the combination of wind and PV under the two capacity scenarios.
+Fig.~\ref{fig:boxplot_number_events}c~\&~d show the combination of wind and PV under the two capacity scenarios.
 In the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
@@
 	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenarios for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 The RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
-A return period refers to the estimated expected interval between occurrences of events with a specified duration or intensity.
+A return period is the estimated average time interval between events of a specified duration or intensity (not to be confused with the frequency of their occurrence within a fixed time frame).
 Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
@@
 
-For wind (Fig.~\ref{fig:return_periods}a), the duration of RES droughts increases in a log-linear fashion across the three datasets.
-The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer RES droughts becoming exponentially less likely as duration increases.
+The duration of wind droughts (Fig.~\ref{fig:return_periods}a) increases in a log-linear fashion across the three datasets.
+The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
 
-For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
+In the case of PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
 The Atlite results show a log-linear increase but reach higher values in general with the longest event lasting forty days.
-For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days.
+For C3S-E G and C3S-E N, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
 Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
@@
 In the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
-For example, the return period for a five-day drought event, shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}, extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
+For example, the return period for a five-day drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
 
@@
 
-For Fig.~\ref{fig:res_droughts_seasonality}a, RES drought percentages are higher in summer than in winter.
-In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as RES droughts, compared to only 4\% in winter (December-January-February).
-This seasonal variation is influenced by the wind power curve model used to estimate CF, where the shape of the curve in lower wind speed regions (3-10 m/s) leads to significant differences in CF under low wind conditions.
-In contrast, the results for Fig.~\ref{fig:res_droughts_seasonality}b show a higher percentage in winter, with RES droughts occurring over 60\% of the time regardless of the dataset.
-The Atlite results show a higher percentage of RES drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.
+The percentage of hours that are part of a wind drought (Fig.~\ref{fig:res_droughts_seasonality}a) are higher in summer than in winter.
+In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as wind droughts, compared to only 4\% in winter (December-January-February).
+This seasonal variation is less prominent for the two C3S-E datasets compared to the Atlite one.
+This difference can be linked to the shape of the two power curves (Fig.~\ref{fig:power_curve}).
+CFs near or under the 0.1 threshold are produced by higher wind speeds for the Atlite power curve than for the C3S-E one.
+In contrast, the results for PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with PV droughts occurring over 60\% of the time regardless of the dataset.
+The Atlite results show a higher percentage of PV drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.
 
@@
 
-Similar to previous results, the 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to \ref{fig:res_droughts_seasonality}a.
+Similar to previous results, the 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
@@
 First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change might affect RES droughts.
-Second, expanding the geographic domain of the study to include the EU would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
-This would require extensive verification across all EU countries, making it a more complex but highly relevant challenge.
+Second, expanding the geographic domain of the study to include the rest of Europe would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
+This would require extensive verification across other European countries, making it a more complex but highly relevant challenge.
 
@@
 
-The meteorological data, including wind speed, irradiance, and temperature, can be accessed through the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
+The ERA5 data can be obtained from the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
 The C3S-E dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
 Information on wind and PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).
-The Atlite scripts used in this study are open-source and can be found on GitHub (\url{https://github.com/pypsa/atlite}).
-The data required to reproduce the figures in this article will be made available upon acceptance of the manuscript.
+The Atlite model used in this study is open-source and can be found on GitHub (\url{https://github.com/pypsa/atlite}).
+The data and code required to reproduce the analysis in this article will be made available upon acceptance of the manuscript in a public GitHub repository.
 

### 2025-01-16 Article/RES_droughts_Morin_ReEn.tex
Update Article

@@
 
-\documentclass[preprint, 12pt, authoryear]{elsarticle}
+\documentclass[preprint, 12pt]{elsarticle}
 
@@
 
-\title{Reducing RES Droughts through the integration of wind and PV} %% Article title
+\newpageafter{author}
 
-\author[Math]{Boris Morin \corref{cor1}} %% Author name
+\title{Reducing RES Droughts through the integration of wind and PV}
+
+\author[Math]{Boris Morin \corref{cor1}}
 \ead{boris.morin@ucdconnect.ie}
 
-\author[Math]{Aina Maimó Far} %% Author name
+\author[Math]{Aina Maimó Far}
 \ead{aina.maimofar@ucd.ie}
 
-\author[Eng]{Damian Flynn} %% Author name
+\author[Eng]{Damian Flynn}
 \ead{damian.flynn@ucd.ie}
 
-\author[Math]{Conor Sweeney} %% Author name
+\author[Math]{Conor Sweeney}
 \ead{conor.sweeney@ucd.ie}
@@
 \begin{keyword}
-RES Drought\sep Wind Power \sep Solar PV Power \sep Renewable Energy Sources \sep Return Periods
+RES Drought \sep Wind Power \sep Solar PV Power \sep Renewable Energy Sources \sep Return Periods
 
@@
 Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy, cds2023energy}.
-One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
+One of the n~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
+
+\section{Data}
+\label{sec:Data}C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
 The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
@@
 In section~\ref{sec:Results}, the models are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
-Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
-
-\section{Data}
-\label{sec:Data}
+Finally, sectio
 
@@
 
-\bibliographystyle{elsarticle-harv}
+\section*{Acknowledgments}
+
+The research conducted in this publication was funded by Science Foundation Ireland and co-funding partners under grant number 21/SPP/3756 through the NexSys Strategic Partnership Programme.
+
+\bibliographystyle{elsarticle-num-names}
 \bibliography{RES_droughts_Morin_ref}
 
-%% Refer following link for more details about bibliography and citations.
-%% https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management
 \end{document}

### 2025-01-17 Article/RES_droughts_Morin_ReEn.tex
Update Article Renewable Energy

@@
 One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable energy generation~\citep{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
-Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\citep{bracken2024drought, allen2023drought}.
+Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\citep{allen2023drought, bracken2024drought}.
 Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\citep{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.

### 2025-01-17 Article/RES_droughts_Morin_ReEn.tex
Updated Abstract

@@
 \begin{abstract}
-The increasing share of renewable energy sources (RES) combined with their dependence on weather poses a critical challenge for energy systems.
-This study investigates the impact of targeting a balanced distribution of wind and photovoltaic (PV) capacity on reducing periods of low renewable generation, known as RES droughts.
-Three different RES models are used to estimate the capacity factors for different installed capacities of wind and PV energy.
-The skill of the RES models is quantified by comparing capacity factor time series to observed data and by assessing their representation of observed RES droughts.
+Increasing the share of electricity produced from renewable energy sources (RES), combined with RES dependence on weather, poses a critical challenge for energy systems.
+This study investigates the importance of the balance between wind and photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
+Three different RES models are used to estimate the capacity factors for different scenarios of installed capacities for wind and PV power.
+The skill of the RES models is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
 The RES models are used to generate a 45-year hourly time series of RES capacity factor, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
 Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
-The addition of PV capacity to a wind-dominated system results in a large reduction in the frequency and duration of RES droughts, as well as reducing extremes and seasonal drought patterns.
+The addition of PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal drought patterns.
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.

### 2025-01-17 Article/RES_droughts_Morin_ReEn.tex
Update Final Draft Article

@@
 This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderwiel2019drought}.
-Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.
+Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasising the need for a resilient energy system to meet both growing electricity demand and decarbonisation targets.
 
@@
 Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy, cds2023energy}.
-One of the n~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
+One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
+The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and is widely used for estimating wind and PV generation~\citep{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
 
 \section{Data}
-\label{sec:Data}C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and is widely used for estimating wind and PV generation~\citep{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
+\label{sec:Data}
 
@@
 In section~\ref{sec:Results}, the models are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
-Finally, sectio
+Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
 

### 2025-01-17 Article/RES_droughts_Morin_ReEn.tex
Update Final Draft

@@
 
-\section{Data}
-\label{sec:Data}
-
 The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their characteristics and relevance for evaluating RES droughts.
@@
 Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
+
+\section{Data}
+\label{sec:Data}
 
@@
 A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
-The performance of each CF time series was then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve.
+The performance of each CF time series was then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve (AUC).
 The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
@@
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
+This is further supported by the AUC, a negatively oriented skill score, which is lower for Atlite than for C3S-E, indicating better alignment with observed data.
 
@@
 
-\newpage
 \subsubsection{PV Energy}
@@
 In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kaneka).
-Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and area under the curve).
+Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and AUC).
 Based on the best-performing metrics, the Breyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
@@
 However, discrepancies appear in the other ranges, where the models diverge from the observed data.
-The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, leading to uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
-With over 65\% of the total PV capacity installed in 2023, these data uncertainties significantly impact the ability to perform rigorous validation for PV drought events.
+The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, with over 65\% of the total PV capacity installed in 2023.
+This results in uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
 
-Nevertheless, the goal of this analysis is to assess the combination of wind and PV generation, where the complementary nature of these energy sources mitigates the limitations seen in PV-only results.
+As, the goal of this analysis is to assess the combination of wind and PV generation, the complementary nature of these energy sources mitigates the limitations in PV-only results.
 
@@
 
-\newpage
 \subsection{Analysis}
@@
 In the case of only PV energy (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
-This increase is due to extended low-generation periods occurring from November to March, depending on the dataset.
+This increase occurs because Ireland, being located above the 50\textdegree parallel, experiences significantly reduced sunlight during the winter months.
+From November to March, PV output often remains consistently low, leading to extended periods where generation stays below the CF threshold.
 When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher than or equal to those for wind, across all duration ranges and datasets.
@@
 	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenarios for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 
-\newpage
 \subsubsection{Return Periods of RES Drought Duration}
@@
 In the case of PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
-The Atlite results show a log-linear increase but reach higher values in general with the longest event lasting forty days.
+The Atlite results show a generally log-linear increase.
 For C3S-E G and C3S-E N, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
 Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
-This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months.
+This sudden increase again reflects the impact of winter on PV generation in Ireland, as PV output usually remains below the CF threshold for extended periods during winter months.
+
 The difference between Atlite and the C3S-E results arises from differences in the datasets near the threshold of 0.1 CF.
@@
 In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
-This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions.
 
@@
 	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 
-\newpage
 \subsubsection{Seasonal Distribution of RES Droughts}
@@
 
-The percentage of hours that are part of a wind drought (Fig.~\ref{fig:res_droughts_seasonality}a) are higher in summer than in winter.
+The percentage of hours that are part of a wind drought (Fig.~\ref{fig:res_droughts_seasonality}a~\&~c) are higher in summer than in winter.
 In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as wind droughts, compared to only 4\% in winter (December-January-February).
@@
 This difference can be linked to the shape of the two power curves (Fig.~\ref{fig:power_curve}).
-CFs near or under the 0.1 threshold are produced by higher wind speeds for the Atlite power curve than for the C3S-E one.
+CFs near or under the 0.1 threshold occur at higher wind speeds for the Atlite power curve than for the C3S-E one.
 In contrast, the results for PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with PV droughts occurring over 60\% of the time regardless of the dataset.
@@
 	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
-Lines correspond to the median values and the area between the first and third quartiles is shaded.}
+Lines correspond to the median values and the area between the first and third quartiles is shaded.
+The scale in b) is different than the other three }
 	\label{fig:res_droughts_seasonality}
@@
 
-Similar to previous results, the 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
+The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
@@
 
-\newpage
 \section{Discussion and Conclusions}
@@
 One of the most evident differences is how each dataset incorporates the specific locations of RES farms.
-Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation.
+Both Atlite and C3S-E G consider the locations of wind and PV farms, which one would expect to result in a more accurate representation of RES generation.
 While this approach slightly improves PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
@@
 The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
-This distinction was also evident in the analysis, where Atlite reported higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
+In this case, that distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
 Again, the results from RES drought modelling rely more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.

### 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
Add minor changes

@@
 The performance of each CF time series was then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve (AUC).
-The area under the curve was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
+The AUC was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
 Based on these metrics, the most representative power curve for Ireland was the Enercon E112.4500 power curve with the $0.3w$  smoothing filter.
@@
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
-This is further supported by the AUC, a negatively oriented skill score, which is lower for Atlite than for C3S-E, indicating better alignment with observed data.
+This is further supported by the AUC - a negatively oriented skill score - which is lower for Atlite than for C3S-E, indicating better alignment with observed data.
 
@@
 
-As, the goal of this analysis is to assess the combination of wind and PV generation, the complementary nature of these energy sources mitigates the limitations in PV-only results.
+As the goal of this analysis is to assess the combination of wind and PV generation, the complementary nature of these energy sources mitigates the limitations in PV-only results.
 
@@
 In the case of only PV energy (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
-This increase occurs because Ireland, being located above the 50\textdegree parallel, experiences significantly reduced sunlight during the winter months.
+This increase occurs because Ireland, being located above the 50\textdegree~parallel, experiences reduced sunlight during the winter months.
 From November to March, PV output often remains consistently low, leading to extended periods where generation stays below the CF threshold.
@@
 Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
-This sudden increase again reflects the impact of winter on PV generation in Ireland, as PV output usually remains below the CF threshold for extended periods during winter months.
+This sudden increase again reflects the impact of winter on PV generation in Ireland, as PV output usually remains below the CF threshold for extended periods during winter.
 
@@
 
-The percentage of hours that are part of a wind drought (Fig.~\ref{fig:res_droughts_seasonality}a~\&~c) are higher in summer than in winter.
+For wind-dominated scenarios (Fig.~\ref{fig:res_droughts_seasonality}a~\&~c), the percentage of hours that are part of a drought is higher in summer than in winter.
 In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as wind droughts, compared to only 4\% in winter (December-January-February).
@@
 Lines correspond to the median values and the area between the first and third quartiles is shaded.
-The scale in b) is different than the other three }
+Note the different y-axis scale for b).}
 	\label{fig:res_droughts_seasonality}

### 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
Update Article
Update Final Draft

@@
 The smoothing of the wind turbine power curve represents losses associated with each turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
-The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.
+The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the observed wind availability data.
 This is further supported by the AUC - a negatively oriented skill score - which is lower for Atlite than for C3S-E, indicating better alignment with observed data.
@@
 From November to March, PV output often remains consistently low, leading to extended periods where generation stays below the CF threshold.
+
 When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher than or equal to those for wind, across all duration ranges and datasets.
@@
 Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
-This sudden increase again reflects the impact of winter on PV generation in Ireland, as PV output usually remains below the CF threshold for extended periods during winter.
+This sudden increase again reflects the impact of extended periods of low PV generation during winter in Ireland.
 

### 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
Update Final Draft

@@
 A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
-The performance of each CF time series was then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and area under the curve (AUC).
-The AUC was calculated from histograms of the hourly CF values for the most recent decade, 2014-2023.
-Based on these metrics, the most representative power curve for Ireland was the Enercon E112.4500 power curve with the $0.3w$  smoothing filter.
+
+The performance of each CF time series is then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and the percentage of overlap.
+The percentage of overlap quantifies the similarity between the observed and modelled distributions.
+It is a positively oriented skill score, where 100\% shows full agreement between the two distributions, and 0\% indicates no overlap.
+The histograms of hourly CF values for the most recent decade (2014-2023) are used to calculate this skill score.
 
@@
 
+Based on these metrics, the most representative power curve for Ireland is the Enercon E112.4500 power curve with the $0.3w$ smoothing filter.
 The smoothing of the wind turbine power curve represents losses associated with each turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the observed wind availability data.
-This is further supported by the AUC - a negatively oriented skill score - which is lower for Atlite than for C3S-E, indicating better alignment with observed data.
+This is further supported by the percentage of overlap which is higher for Atlite (97.2\%) than for C3S-En (83.2\%), indicating better agreement with observed data.
 
@@
 In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kaneka).
-Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MB, and AUC).
+Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MBE, and the percentage of overlap).
 Based on the best-performing metrics, the Breyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
@@
 Differences between the models are smaller for PV, with C3S-G performing marginally better than the other two.
-The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact.
-In this case, that distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
+The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact in this case.
+This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
 Again, the results from RES drought modelling rely more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.

### 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
Update Final Draft

@@
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the observed wind availability data.
-This is further supported by the percentage of overlap which is higher for Atlite (97.2\%) than for C3S-En (83.2\%), indicating better agreement with observed data.
+This is further supported by the percentage of overlap which is higher for Atlite (97.2\%) than for C3S-E (83.2\%), indicating better agreement with observed data.
 

### 2025-02-06 Article/RES_droughts_Morin_Earth.tex
Update preprint version of the Article

--- 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
@@
 \end{abstract}
-
-%%Research highlights
-\begin{highlights}
-\item RES droughts are analysed using 45 years of hourly wind and PV generation data
-
-\item RES droughts from C3S-Energy and ERA5-Atlite datasets are compared
-
-\item Adding PV to a wind-dominated system reduces RES drought frequency and duration
-
-\item Validated RES datasets are crucial to accurately identify RES drought extremes
-
-\end{highlights}
 

### 2025-03-11 Article/RES_droughts_Morin_ReEn_R2.tex
Created new tex file
It contains the modifications that I am making to answer the reviewers

--- 2025-01-20 Article/RES_droughts_Morin_ReEn.tex
@@
 This case study provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
-The performance of different RES models are compared, and a 45-year time series of RES generation is produced.
+The performance of different RES datasets are compared, and a 45-year time series of RES generation is produced.
 The results highlight the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
@@
 Alternative methods exist for defining RES droughts.
-One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable energy generation~\citep{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
+One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable electricity generation~\citep{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
 Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\citep{allen2023drought, bracken2024drought}.
 Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\citep{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
-In this paper, the focus is exclusively on energy generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
+In this paper, the focus is exclusively on renewable electricity generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
 
@@
 The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their characteristics and relevance for evaluating RES droughts.
-Section~\ref{sec:Methods} outlines the RES models used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:Results}, the models are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
+Section~\ref{sec:Methods} outlines the RES datasets used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
+In section~\ref{sec:Results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
 Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
@@
 
-This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy.
+This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and PV energy.
 The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S-E datasets.
@@
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
-It is widely used for estimating PV and wind energy~\citep{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
+It is widely used for estimating solar and wind energy~\citep{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
 Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.
@@
 
-In addition to analysing wind and PV generation separately, a combined CF was computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for PV power).
+In addition to analysing wind and PV generation separately, a combined CF was computed for each dataset by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for PV power).
 This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% PV capacity.
@@
 
-New time series were generated for both the Atlite and C3S-E G PV models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the Atlite and C3S-E G PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
-In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three models for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different models.
+In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three datasets for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different datasets.
 
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N models}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}
 	\label{fig:wind_verification_contour}
@@
 Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MBE, and the percentage of overlap).
-Based on the best-performing metrics, the Breyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
+Based on the best-performing metrics, the Beyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
 For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
@@
 	\includegraphics[width=\textwidth]{verification_pv_contour.png}
-	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
+	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}	
 	\label{fig:solar_verification_contour}
@@
 
-This study has investigated the ability of three RES models to represent RES droughts: Atlite, C3S-E G, and C3S-E N.
+This study has investigated the ability of three RES datasets to represent RES droughts: Atlite, C3S-E G, and C3S-E N.
 One of the most evident differences is how each dataset incorporates the specific locations of RES farms.
@@
 While this approach slightly improves PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
-This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES model is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
+This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES dataset is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
 
 Atlite shows the best alignment with observed data for wind generation.
-Differences between the models are smaller for PV, with C3S-G performing marginally better than the other two.
+Differences between the datasets are smaller for PV, with C3S-G performing marginally better than the other two.
 The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact in this case.

### 2025-03-12 Article/RES_droughts_Morin_ReEn_R2.tex
Update reviewed version of the article

@@
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
-This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderwiel2019drought}.
+This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasising the need for a resilient energy system to meet both growing electricity demand and decarbonisation targets.
@@
 
-For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in other research~\citep{kaspar2019drought, ohba2022drought, mockert2023drought, mayer2023drought}.
+For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in previous research.
+Kaspar et al.~\citep{kaspar2019drought} analysed the shortfall risks of low wind and PV in Europe, with a focus on Germany.
+Mockert et al.~\citep{mockert2023drought} expanded on this by examining the link between weather regimes and RES droughts in Germany.
+Similar analyses were conducted using machine learning for Japan \citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
 Alternative methods exist for defining RES droughts.
-One approach uses relative CF thresholds that change over the year to account for seasonal variations in renewable electricity generation~\citep{raynaud2018drought, rinaldi2021drought, gangopadhyay2022drought, allen2023drought, kapica2024drought}.
-Another common method relies on percentile-based thresholds, where drought events are defined by identifying periods of unusually low generation relative to historical production levels, typically based on the lowest production percentiles~\citep{allen2023drought, bracken2024drought}.
-Additionally, some studies combine these definitions with metrics that incorporate the demand side of energy consumption, analysing the balance between supply and demand during drought periods~\citep{raynaud2018drought, rinaldi2021drought, allen2023drought, bracken2024drought}.
+One approach uses relative CF thresholds that adjust throughout the year to account for seasonal variations in renewable electricity generation.
+Raynaud et al.~\citep{raynaud2018drought} defined a drought as a sequence of days with energy production below a threshold, applying this method over a number of European regions.
+This methodology was later used to study wind and PV droughts in India using machine learning methods~\citep{gangopadhyay2022drought}.
+Kapica et al.~\citep{kapica2024drought} built upon this approach to compare the likelihood of RES droughts increasing in Europe under different climate models.
+Rinaldi et al.~\citep{rinaldi2021drought} instead defined RES droughts as periods when wind or PV CF falls below a percentage of the daily mean for that time of year and assessed RES drought risks in the U.S.
+Western Interconnection.
+Also focusing on the U.S., Brown et al.~\citep{brown2021drought} used a weekly timescale rather than hourly or daily data and examined RES droughts from a meteorological perspective.
+Another method defines energy drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
+This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
+It has been applied in other studies, including analyses of RES droughts in the U.S.~\citep{bracken2024drought} and China~\citep{lei2024drought}.
+In addition to examining periods of low renewable electricity generation, Raynaud et al.~\citep{raynaud2018drought} analysed the imbalance between electricity demand and renewable generation, known as residual load.
+These events were studied alongside low-generation periods to assess their correlation.
+Similar analyses have been conducted in Europe~\citep{allen2023drought} and the U.S.~\citep{bracken2024drought}, revealing differing results across regions.
+
 In this paper, the focus is exclusively on renewable electricity generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
@@
 In this study, three different datasets are used, all of which are driven by ERA5 data~\citep{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy, cds2023energy}.
+Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
 One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
 The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and is widely used for estimating wind and PV generation~\citep{mockert2023drought, li2023atlite, parzen2023atlite, ali2023comparative}.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and PV generation in order to study RES droughts~\citep{mockert2023drought}.
 

### 2025-03-13 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed version of the Article

@@
 \section{Introduction}
-\label{sec:Intro}
+\label{sec:intro}
 
@@
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
-This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
+This challenge is amplified by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasising the need for a resilient energy system to meet both growing electricity demand and decarbonisation targets.
@@
 
-The datasets used in this study are detailed in section~\ref{sec:Data}, which describes their characteristics and relevance for evaluating RES droughts.
-Section~\ref{sec:Methods} outlines the RES datasets used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:Results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
-Finally, section~\ref{sec:Conclusion} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
+The datasets used in this study are detailed in section~\ref{sec:data}, which describes their characteristics and relevance for evaluating RES droughts.
+Section~\ref{sec:methods} outlines the RES datasets used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
+In section~\ref{sec:results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
+Finally, section~\ref{sec:conclusions} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
 
 \section{Data}
-\label{sec:Data}
+\label{sec:data}
 
@@
 
-The C3S-E dataset estimates wind energy using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
-This choice is based on expert advice and the trend in wind turbine installation.
-The PV generation model used by C3S-E uses two ERA5 variables: surface solar radiation downwards ($ssrd$) and air temperature ($t2m$).
-PV generation is calculated multiple times, using the same model with different azimuth and tilt angles.
-The results are aggregated based on a statistical distribution of the module angles based on the geographical location~\citep{saintdrenan2018solar}.
+The wind CF in the C3S-E model is calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
+This choice reflects trends in wind turbine installations and was guided by expert recommendations.
+Since real-time data on the exact wind turbine fleet across Europe is difficult to obtain, C3S-E assumes a homogeneous distribution of turbines across the ERA5 grid.
+While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of climate variability on wind power generation.
+The turbine power curves used in the model are sourced from publicly available databases, ensuring consistency with industry standards.
+The PV CF in the C3S-E model is calculated at the grid level and represents the aggregated output of all PV systems within each pixel, rather than a single installation.
+It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference PV plant model.
+This model incorporates empirical calculations for key system components such as optical losses, module efficiency, and inverters.
+The final capacity factor accounts for a mix of module orientations typical for each location~\citep{saintdrenan2018solar}.
 
 \section{Methods}
-\label{sec:Methods}
+\label{sec:methods}
 
@@
 \section{Results}
-\label{sec:Results}
+\label{sec:results}
 
@@
 \subsection{Analysis}
-\label{sec:Analysis}
+\label{sec:analysis}
 
@@
 
-The first part of the analysis examines the annual number of RES drought events across the three datasets.
-When only wind energy is considered (Fig.~\ref{fig:boxplot_number_events}a), the number of events decreases as the duration range increases, with very few events lasting more than seven days.
-In the case of only PV energy (Fig.~\ref{fig:boxplot_number_events}b), the number of events also declines as the duration range extends from one to eight days, followed by a slight increase for longer durations.
-This increase occurs because Ireland, being located above the 50\textdegree~parallel, experiences reduced sunlight during the winter months.
-From November to March, PV output often remains consistently low, leading to extended periods where generation stays below the CF threshold.
+The analysis of annual RES drought events reveals trends that are largely consistent with earlier studies.
+When only wind energy is considered (Fig.\ref{fig:boxplot_number_events}a), the number of drought events decreases as the duration range increases, with very few events lasting more than seven days.
+This pattern aligns with previous research showing that wind droughts tend to be short and frequent.
+In contrast, for PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
+This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low PV output.
 
-When comparing wind and PV results (Fig.~\ref{fig:boxplot_number_events}a~\&~b), the median, first, and third quartiles for PV are consistently higher than or equal to those for wind, across all duration ranges and datasets.
-This is due to the typically lower CF of PV power compared to wind power, especially in a region such as Ireland where solar potential is limited.
-PV generation is also zero at night and constrained by the daily solar cycle, leading to a naturally higher frequency of drought events in PV compared to wind.
+Moreover, the comparison between wind and PV results indicates that the median, first, and third quartiles for PV are consistently higher than or equal to those for wind.
+This is expected, given that PV generation is inherently lower, zero at night, and limited by the solar cycle, as observed in other studies.
+When wind and PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, reaffirming wind’s dominance in the current energy mix.
+However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N, demonstrating the beneficial effects of a more balanced energy mix.
+These findings are in line with earlier studies that highlight how increasing PV capacity can mitigate drought frequency through the anti-correlated seasonal patterns of wind and solar generation.
 
-Fig.~\ref{fig:boxplot_number_events}c~\&~d show the combination of wind and PV under the two capacity scenarios.
-In the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the identified RES droughts closely match those for wind alone, which is expected due to the dominance of installed wind capacity.
-In contrast, the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d) shows a clear reduction in the number of drought events across all datasets and durations, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N.
-This reduction is attributed to the anti-correlation between wind and PV generation.
-
-The median, first, and third quartiles for the Atlite dataset are consistently greater than or equal to those of the other two datasets, regardless of the duration range or type of renewable energy considered.
-This difference arises from the wind turbine power curve model used in the C3S-E datasets, which tends to overestimate the wind CF (Fig.~\ref{fig:wind_verification_contour}).
-As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite.
+Additionally, the consistently higher drought counts reported by the Atlite dataset, compared to the C3S-E datasets, underscore the impact of model selection, particularly the influence of wind turbine power curve representation, on quantifying RES droughts.
+This observation is consistent with previous research, which has also noted that assumptions regarding turbine characteristics can significantly affect drought duration estimates.
 
@@
 Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
-This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the potential impact of RES droughts on energy reliability and security of supply.
+This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the impact of extreme RES droughts on system resilience.
+Understanding these return periods is essential, as even infrequent can challenge energy security by placing significant strain on conventional backup sources necessary to maintain supply in high-RES scenarios.
 
-The duration of wind droughts (Fig.~\ref{fig:return_periods}a) increases in a log-linear fashion across the three datasets.
-The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
+For wind (Fig.~\ref{fig:return_periods}a), the log‐linear increase in return periods observed in this study confirms that longer droughts occur exponentially less frequently, a trend consistent with earlier research on wind variability.
+In the case of PV droughts (Fig.~\ref{fig:return_periods}b), the Atlite dataset shows a generally log‐linear trend, while the C3S-E datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
+This abrupt rise reflects differences in how PV output is handled near the CF threshold during low irradiance conditions.
+Interestingly, when the share of wind power is decreased in the balanced scenario, the overall return periods for drought events increase.
+Although wind energy is generally more stable, reducing its share makes the combined system more vulnerable to PV’s extended low-generation periods in winter.
+This highlights that the seasonal limitations of PV can dominate the overall drought characteristics when its share increases, emphasizing the importance of carefully balancing the energy mix for RES drought risk assessment.
 
-In the case of PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S-E datasets.
-The Atlite results show a generally log-linear increase.
-For C3S-E G and C3S-E N, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
-Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
-This sudden increase again reflects the impact of extended periods of low PV generation during winter in Ireland.
-
-The difference between Atlite and the C3S-E results arises from differences in the datasets near the threshold of 0.1 CF.
-Atlite remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events.
-In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
-
-For the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the return periods mirror those of Fig.~\ref{fig:return_periods}a, due to the low levels of installed PV capacity.
-In the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations.
-For example, the return period for a five-day drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets.
+Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods largely mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
+In contrast, the balanced 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
 
@@
 For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
-This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influence drought duration estimates.
-Additionally, in all four graphs, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
+This difference underscores the importance of model selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence drought duration estimates.
+Additionally, in all four graphs, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
 
@@
 
-The seasonality of RES droughts was analysed by comparing the percentage of hours in each month classified as part of a RES drought.
+The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events.
+For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the Atlite dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February).
+This strong seasonal signal is less evident in the C3S-E datasets, which suggests that the differences in the underlying wind power curves play a significant role.
+In Atlite, CF near or below the 0.1 threshold occur at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
 
-For wind-dominated scenarios (Fig.~\ref{fig:res_droughts_seasonality}a~\&~c), the percentage of hours that are part of a drought is higher in summer than in winter.
-In the Atlite dataset, for instance, an average of 24\% of hours in summer (June-July-August) are identified as wind droughts, compared to only 4\% in winter (December-January-February).
-This seasonal variation is less prominent for the two C3S-E datasets compared to the Atlite one.
-This difference can be linked to the shape of the two power curves (Fig.~\ref{fig:power_curve}).
-CFs near or under the 0.1 threshold occur at higher wind speeds for the Atlite power curve than for the C3S-E one.
-In contrast, the results for PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) show a higher percentage in winter, with PV droughts occurring over 60\% of the time regardless of the dataset.
-The Atlite results show a higher percentage of PV drought hours for wind, and a slightly lower percentage for PV, compared to the two C3S-E datasets.
+In contrast, PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
+Across all datasets, over 60\% of winter hours are classified as PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
+Moreover, Atlite tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for PV relative to the C3S-E datasets.
+These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
+
+The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
+However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
+This reduction can be explained by the contribution of PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
+Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
+In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
+The balanced mix of wind and PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
 
@@
 
-The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
-However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
-Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
-In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-The balanced mix of wind and PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
+The seasonal variations observed in this study have important implications for energy planning.
+Given that energy demand peaks in winter for Northern European countries, understanding these seasonal patterns is critical for assessing the need for conventional backup or storage solutions during periods of prolonged low renewable output.
+The findings underscore that even small differences in model assumptions leads to significant variations in drought estimates, thereby affecting the reliability of the energy system during critical periods.
+Such insights are essential for policymakers to develop targeted strategies that enhance grid resilience and ensure a stable energy supply throughout the year.
 
-\section{Discussion and Conclusions}
-\label{sec:Conclusion}
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 Additionally, the choice of model for RES drought assessment becomes increasingly critical as more renewable capacity is integrated into the system.
+
+This study has several limitations.
+Although ERA5 is among the best reanalysis datasets for renewable energy analysis \citep{doi:10.3390/atmos12050624} ?, its resolution may not capture local-scale phenomena, making it less reliable at the individual farm level.
+In addition, our methodology employs a fixed threshold to define RES drought events, which is necessary to compare the three models but does not account for demand variations.
+Consequently, while this approach enables a consistent inter-comparison, it may overlook events that are most critical for power system operations.
 

### 2025-03-14 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed version of the Article

@@
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
-It is widely used for estimating solar and wind energy~\citep{mockert2023drought, dubus2023energy, brown2021drought, otero2022drought}.
+It has proven to be the best choice for studying renewable energy in Ireland \citep{doddy2021era5}.
 Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.
@@
 
+It is important to note that this analysis is not focused on the absolute numbers characterizing RES droughts, but rather on the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.
+
 \subsubsection{Annual Number of RES Droughts}
@@
 The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events.
+Wind droughts tend to be more frequent during summer, whereas PV droughts are more common in winter due to reduced sunlight.
+By comparing these seasonal patterns across different datasets and energy scenarios, the study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of drought events.
+
 For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the Atlite dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February).

### 2025-03-20 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 
-\title{Reducing RES Droughts through the integration of wind and PV}
+\title{Reducing RES Droughts through the integration of wind and Solar PV}
 
@@
 This study investigates the importance of the balance between wind and photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
-Three different RES models are used to estimate the capacity factors for different scenarios of installed capacities for wind and PV power.
+Three different RES models are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
 The skill of the RES models is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
@@
 Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
-The addition of PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal drought patterns.
+The addition of solar PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal drought patterns.
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
@@
 \begin{highlights}
-\item RES droughts are analysed using 45 years of hourly wind and PV generation data
+\item RES droughts are analysed using 45 years of hourly wind and solar PV generation data
 
@@
 
-\item Adding PV to a wind-dominated system reduces RES drought frequency and duration
+\item Adding solar PV to a wind-dominated system reduces RES drought frequency and duration
 
@@
 
-This study focuses on Ireland, a region with a strong reliance on wind power, which has ambitious targets for PV power expansion.
+This study focuses on Ireland, a region with a strong reliance on wind power, which has ambitious targets for solar PV power expansion.
 This case study provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
 The performance of different RES datasets are compared, and a 45-year time series of RES generation is produced.
-The results highlight the role of increased PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
+The results highlight the role of increased solar PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
 
 For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in previous research.
-Kaspar et al.~\citep{kaspar2019drought} analysed the shortfall risks of low wind and PV in Europe, with a focus on Germany.
+Kaspar et al.~\citep{kaspar2019drought} analysed the shortfall risks of low wind and solar PV in Europe, with a focus on Germany.
 Mockert et al.~\citep{mockert2023drought} expanded on this by examining the link between weather regimes and RES droughts in Germany.
@@
 Raynaud et al.~\citep{raynaud2018drought} defined a drought as a sequence of days with energy production below a threshold, applying this method over a number of European regions.
-This methodology was later used to study wind and PV droughts in India using machine learning methods~\citep{gangopadhyay2022drought}.
+This methodology was later used to study wind and solar PV droughts in India using machine learning methods~\citep{gangopadhyay2022drought}.
 Kapica et al.~\citep{kapica2024drought} built upon this approach to compare the likelihood of RES droughts increasing in Europe under different climate models.
-Rinaldi et al.~\citep{rinaldi2021drought} instead defined RES droughts as periods when wind or PV CF falls below a percentage of the daily mean for that time of year and assessed RES drought risks in the U.S.
+Rinaldi et al.~\citep{rinaldi2021drought} instead defined RES droughts as periods when wind or solar PV CF falls below a percentage of the daily mean for that time of year and assessed RES drought risks in the U.S.
 Western Interconnection.
@@
 
-In this paper, the focus is exclusively on renewable electricity generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and PV capacities.
+In this paper, the focus is exclusively on renewable electricity generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
 
-RES droughts are identified using onshore wind and PV CF time series.
+RES droughts are identified using onshore wind and solar PV CF time series.
 In this study, three different datasets are used, all of which are driven by ERA5 data~\citep{hersbach2020era5}.
@@
 The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and PV generation in order to study RES droughts~\citep{mockert2023drought}.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts~\citep{mockert2023drought}.
+
+The aim of this real case study is to answer two questions which could help on the decision making for the planning of reserve capacity in wind-dominated renewable energy system 
+\begin{itemize}
+	\item What is the impact of selecting different modelling assumptions on the analysis of RES droughts?
+	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts in a real-case setting?
+\end{itemize}
 
 The datasets used in this study are detailed in section~\ref{sec:data}, which describes their characteristics and relevance for evaluating RES droughts.
-Section~\ref{sec:methods} outlines the RES datasets used to simulate wind and PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to PV capacities.
+Section~\ref{sec:methods} outlines the RES datasets used to simulate wind and solar PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
+In section~\ref{sec:results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to solar PV capacities.
 Finally, section~\ref{sec:conclusions} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
@@
 
-This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and PV energy.
+This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV energy.
 The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S-E datasets.
 
-\subsection{Wind and PV Capacity and Availability}
+\subsection{Wind and solar PV Capacity and Availability}
 \label{sec:eirgrid}
 
-EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\citep{eirgrid2023spreadsheet}.
+EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and solar PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\citep{eirgrid2023spreadsheet}.
 These datasets include information such as each farm’s installed capacity, name, and connection date.
@@
 Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-Generation and availability values are available from 2014 onward for wind power and from 2018 onward for PV power, although PV availability data only became present in the Republic of Ireland in 2023.
+Generation and availability values are available from 2014 onward for wind power and from 2018 onward for solar PV power, although solar PV availability data only became present in the Republic of Ireland in 2023.
 This study focuses on availability for all analyses.
@@
 	\centering
-	\caption{ERA5 variables used to calculate wind and PV generation}
+	\caption{ERA5 variables used to calculate wind and solar PV generation}
 	\begin{tabular}{|l|c|}
@@
 The EU Copernicus Climate Change Service developed the C3S-E renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
-This dataset provides hourly CF for wind and PV energy from 1979 to the present.
+This dataset provides hourly CF for wind and solar PV energy from 1979 to the present.
 The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°.
@@
 The turbine power curves used in the model are sourced from publicly available databases, ensuring consistency with industry standards.
-The PV CF in the C3S-E model is calculated at the grid level and represents the aggregated output of all PV systems within each pixel, rather than a single installation.
-It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference PV plant model.
+The solar PV CF in the C3S-E model is calculated at the grid level and represents the aggregated output of all solar PV systems within each pixel, rather than a single installation.
+It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference solar PV plant model.
 This model incorporates empirical calculations for key system components such as optical losses, module efficiency, and inverters.
@@
 Atlite transforms weather data into energy data using the gridded ERA5 data and the locations of existing RES farms, as described in C3S-E G.
-ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate PV generation.
+ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate solar PV generation.
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
@@
 
-In addition to analysing wind and PV generation separately, a combined CF was computed for each dataset by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for PV power).
+In addition to analysing wind and solar PV generation separately, a combined CF was computed for each dataset by averaging wind and solar PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
 This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% PV capacity.
-Given that PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
+Given that solar PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and solar PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed solar PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
 These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\citep{cap2024future}.
@@
 
-New time series were generated for both the Atlite and C3S-E G PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the Atlite and C3S-E G solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
-In total, twelve CF time series were analysed in this study, six for individual wind and PV CF (three datasets for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different datasets.
+In total, twelve CF time series were analysed in this study, six for individual wind and solar PV CF (three datasets for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different datasets.
 
@@
 
-\subsubsection{PV Energy}
+\subsubsection{Solar PV Energy}
 \label{sec:pv_verification}
@@
 Based on the best-performing metrics, the Beyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
-For all PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
+For all solar PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
 
-The PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed PV capacity.
-To enable actual PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
-This scaling factor was estimated by analysing proprietary data from individual PV farms provided by EirGrid, which showed that, on average, assuming that the installed capacities of farms exceed their MEC values by 40\% yields the best agreement with the observed availability.
+The solar PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed solar PV capacity.
+To enable actual solar PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
+This scaling factor was estimated by analysing proprietary data from individual solar PV farms provided by EirGrid, which showed that, on average, assuming that the installed capacities of farms exceed their MEC values by 40\% yields the best agreement with the observed availability.
 
@@
 	\includegraphics[width=\textwidth]{verification_pv_contour.png}
-	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}	
+	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}	
 	\label{fig:solar_verification_contour}
@@
 	\end{tabular}
-	\caption{Skill scores for PV CF for the three datasets compared to observed data}
+	\caption{Skill scores for Solar PV CF for the three datasets compared to observed data}
 	\label{tab:pv_skill_scores}
@@
 
-Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of PV drought events during the 2023 validation period across different duration ranges.
+Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of solar PV drought events during the 2023 validation period across different duration ranges.
 The figure reveals partial agreement between the three datasets and the observed data, with consistent results noticed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
 However, discrepancies appear in the other ranges, where the models diverge from the observed data.
-The main challenge in validating PV data stems from the recent installation of a large share of Ireland’s PV capacity, with over 65\% of the total PV capacity installed in 2023.
+The main challenge in validating solar PV data stems from the recent installation of a large share of Ireland’s PV capacity, with over 65\% of the total PV capacity installed in 2023.
 This results in uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
@@
 This strong seasonal signal is less evident in the C3S-E datasets, which suggests that the differences in the underlying wind power curves play a significant role.
-In Atlite, CF near or below the 0.1 threshold occur at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
+In Atlite, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
 

### 2025-03-21 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 Increasing the share of electricity produced from renewable energy sources (RES), combined with RES dependence on weather, poses a critical challenge for energy systems.
-This study investigates the importance of the balance between wind and photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
+This study investigates the importance of the balance between wind and solar photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
 Three different RES models are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
@@
 The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \citep{eurostat2023share}.
-While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power.
+While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and solar photovoltaic (PV) power.
 This challenge is amplified by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
@@
 
-For this study, a RES drought event is defined as occurring when the average capacity factor (CF) remains below a fixed threshold for a given duration, following the methodology used in previous research.
-Kaspar et al.~\citep{kaspar2019drought} analysed the shortfall risks of low wind and solar PV in Europe, with a focus on Germany.
-Mockert et al.~\citep{mockert2023drought} expanded on this by examining the link between weather regimes and RES droughts in Germany.
-Similar analyses were conducted using machine learning for Japan \citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
-Alternative methods exist for defining RES droughts.
-One approach uses relative CF thresholds that adjust throughout the year to account for seasonal variations in renewable electricity generation.
-Raynaud et al.~\citep{raynaud2018drought} defined a drought as a sequence of days with energy production below a threshold, applying this method over a number of European regions.
-This methodology was later used to study wind and solar PV droughts in India using machine learning methods~\citep{gangopadhyay2022drought}.
-Kapica et al.~\citep{kapica2024drought} built upon this approach to compare the likelihood of RES droughts increasing in Europe under different climate models.
-Rinaldi et al.~\citep{rinaldi2021drought} instead defined RES droughts as periods when wind or solar PV CF falls below a percentage of the daily mean for that time of year and assessed RES drought risks in the U.S.
-Western Interconnection.
-Also focusing on the U.S., Brown et al.~\citep{brown2021drought} used a weekly timescale rather than hourly or daily data and examined RES droughts from a meteorological perspective.
+Various approaches have been proposed in the literature to define and quantify RES drought events.
+One common method defines a RES drought as a period during which the average capacity factor (CF) remains below a fixed threshold for a specified duration.
+For example, Kaspar et al.~\citep{kaspar2019drought} investigated the shortfall risks of low wind and solar PV generation in Europe, focusing on Germany, while Mockert et al.~\citep{mockert2023drought} extended this work to examine the link between weather regimes and RES droughts.
+Similar fixed-threshold approaches have also been applied using machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
+
+Alternative methods adjust the CF threshold dynamically over the year to account for seasonal variations in renewable production.
+Raynaud et al.~\citep{raynaud2018drought} defined droughts as sequences of days with energy production below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}.
+Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models.
+Other studies have defined droughts based on deviations from daily mean production, as done by Rinaldi et al.~\citep{rinaldi2021drought} in the U.S.
+Western Interconnection, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences.
 Another method defines energy drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
@@
 It has been applied in other studies, including analyses of RES droughts in the U.S.~\citep{bracken2024drought} and China~\citep{lei2024drought}.
+
 In addition to examining periods of low renewable electricity generation, Raynaud et al.~\citep{raynaud2018drought} analysed the imbalance between electricity demand and renewable generation, known as residual load.
@@
 
-The aim of this real case study is to answer two questions which could help on the decision making for the planning of reserve capacity in wind-dominated renewable energy system 
+The aim of this study is to answer two questions which could help on the decision making for the planning of reserve capacity in real case wind-dominated renewable energy system:
 \begin{itemize}
@@
 
-This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV energy.
+This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV power.
 The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S-E datasets.
@@
 This model incorporates empirical calculations for key system components such as optical losses, module efficiency, and inverters.
-The final capacity factor accounts for a mix of module orientations typical for each location~\citep{saintdrenan2018solar}.
+The final CF accounts for a mix of module orientations typical for each location~\citep{saintdrenan2018solar}.
 
@@
 
-For national-level analyses, the aggregated CF time series provided by C3S-E were used at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
-These are based on the assumption by C3S-E that RES generation occurs at every ERA5 grid point in Ireland.
-We computed a weighted average of these, based on the installed capacity of each one, to represent the total CF for Ireland.
+For national-level analyses, the inputs for this dataset are the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+These values are based on the assumption that RES generation occurs at every ERA5 grid point in Ireland.
+A weighted average is computed to represent the total CF for the country.
 
@@
 
-The gridded dataset from C3S-E was used to create CF datasets which account for the location of RES farms in Ireland.
-A list of the RES farms in Ireland was compiled, including each farm’s latitude, longitude and installed capacity.
-Using these coordinates, the nearest grid point on the C3S-E grid was identified for each farm.
-The CF values from the C3S-E dataset corresponding to these grid points were retrieved.
-A weighted average of the CF values was calculated, with the installed capacity of each farm serving as the weight, to construct the CF time series for Ireland.
-This process resulted in a time series of RES generation for each energy source (wind and PV) for Ireland, which takes the location of the RES farms into account.
+For the gridded dataset, the inputs consist of the CF time series from C3S-E, along with the location of individual RES farms across Ireland.
+Using these inputs, the nearest grid point on the C3S-E dataset was identified for each farm, and the corresponding CF values were retrieved.
+A weighted average, based on installed capacity, was then computed to construct a CF time series that accounts for the actual spatial distribution of wind and solar PV farms in Ireland.
 
@@
 
-Atlite transforms weather data into energy data using the gridded ERA5 data and the locations of existing RES farms, as described in C3S-E G.
-ERA5 data for wind speed at 100 metres ($u_{100}$, $v_{100}$) are used to calculate wind generation, while the ERA5 radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) and air temperature ($t2m$) are used to calculate solar PV generation.
+For the Atlite dataset, the inputs include the locations of RES farms and ERA5 weather variables, such as wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
+The meteorological inputs are processed using the Atlite model to estimate CF time series for wind and solar PV, incorporating specific characteristics such as wind turbines power curve and PV panel model.
 A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
@@
 
+The output of those three datasets are one CF time series for both wind and Solar PV.
 In addition to analysing wind and solar PV generation separately, a combined CF was computed for each dataset by averaging wind and solar PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
@@
 This approach allows for a comparative analysis between the two scenarios, assessing how the balance of RES capacity affects the occurrence of RES droughts.
+
+For each dataset (Atlite, C3S-E G, and C3S-E N), four distinct scenarios are examined, as summarised below:
+
+\begin{itemize}
+	\item Wind Power / 91W-9PV
+	\item Solar PV Power / 91W-9PV
+	\item Combined RES / 91W-9PV
+	\item Combined RES / 57W-43PV
+\end{itemize}
 
@@
 However, discrepancies appear in the other ranges, where the models diverge from the observed data.
-The main challenge in validating solar PV data stems from the recent installation of a large share of Ireland’s PV capacity, with over 65\% of the total PV capacity installed in 2023.
-This results in uncertainties in PV generation data and the actual generating capacity in the first few months after each farm is connected.
+The main challenge in validating solar PV data stems from the recent installation of a large share of Ireland’s solar PV capacity, with over 65\% of the total solar PV capacity installed in 2023.
+This results in uncertainties in solar PV generation data and the actual generating capacity in the first few months after each farm is connected.
 
-As the goal of this analysis is to assess the combination of wind and PV generation, the complementary nature of these energy sources mitigates the limitations in PV-only results.
+As the goal of this analysis is to assess the combination of wind and solar PV generation, the complementary nature of these energy sources mitigates the limitations in solar PV-only results.
 
@@
 	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
-	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline).
-The PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
+	\caption{Number of solar PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline).
+The solar PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
 	\label{fig:bar_number_events_verification_pv}
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and PV capacity increases to 8.6 GW.
+In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of solar PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and solar PV capacity increases to 8.6 GW.
 Both scenarios were driven by 45 years of ERA5 data.
-Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and PV droughts are first analysed separately before presenting the results for combined (wind + PV) RES droughts under both scenarios.
+Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and solar PV droughts are first analysed separately before presenting the results for combined (wind + solar PV) RES droughts under both scenarios.
 
-It is important to note that this analysis is not focused on the absolute numbers characterizing RES droughts, but rather on the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.
+It is important to highlight that this analysis considers two key aspects: the absolute values that characterise RES droughts, which are crucial for power system planning, and the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.
 
@@
 This pattern aligns with previous research showing that wind droughts tend to be short and frequent.
-In contrast, for PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
-This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low PV output.
+In contrast, for solar PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
+This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low solar PV output.
 
-Moreover, the comparison between wind and PV results indicates that the median, first, and third quartiles for PV are consistently higher than or equal to those for wind.
-This is expected, given that PV generation is inherently lower, zero at night, and limited by the solar cycle, as observed in other studies.
-When wind and PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, reaffirming wind’s dominance in the current energy mix.
+Moreover, the comparison between wind and solar PV results indicates that the median, first, and third quartiles for solar PV are consistently higher than or equal to those for wind.
+This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle, as observed in other studies.
+When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, reaffirming wind’s dominance in the current energy mix.
 However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N, demonstrating the beneficial effects of a more balanced energy mix.
-These findings are in line with earlier studies that highlight how increasing PV capacity can mitigate drought frequency through the anti-correlated seasonal patterns of wind and solar generation.
+These findings are in line with earlier studies that highlight how increasing solar PV capacity can mitigate drought frequency through the anti-correlated seasonal patterns of wind and solar generation.
 
@@
 	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
-This analysis provides insight into the frequency and likelihood of prolonged low-generation periods, which is crucial for evaluating the impact of extreme RES droughts on system resilience.
-Understanding these return periods is essential, as even infrequent can challenge energy security by placing significant strain on conventional backup sources necessary to maintain supply in high-RES scenarios.
+This analysis not only quantifies the likelihood of prolonged low-generation periods but also provides insight into how extreme events are distributed across different timescales, helping to assess the variability of rare but impactful events.
+Understanding these return periods is crucial, as even infrequent droughts can challenge energy security by placing significant strain on conventional backup sources necessary to maintain supply in high-RES scenarios.
 
 For wind (Fig.~\ref{fig:return_periods}a), the log‐linear increase in return periods observed in this study confirms that longer droughts occur exponentially less frequently, a trend consistent with earlier research on wind variability.
-In the case of PV droughts (Fig.~\ref{fig:return_periods}b), the Atlite dataset shows a generally log‐linear trend, while the C3S-E datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
-This abrupt rise reflects differences in how PV output is handled near the CF threshold during low irradiance conditions.
-Interestingly, when the share of wind power is decreased in the balanced scenario, the overall return periods for drought events increase.
-Although wind energy is generally more stable, reducing its share makes the combined system more vulnerable to PV’s extended low-generation periods in winter.
-This highlights that the seasonal limitations of PV can dominate the overall drought characteristics when its share increases, emphasizing the importance of carefully balancing the energy mix for RES drought risk assessment.
+In the case of solar PV droughts (Fig.\ref{fig:return_periods}b), the Atlite dataset shows a general log‐linear trend, whereas the C3S-E datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
+This abrupt rise reflects differences in how solar PV output is handled near the CF threshold during low irradiance conditions.
+In the balanced scenario, the reduced share of wind and increased share of solar PV leverages their complementary seasonal patterns, resulting in higher return periods for combined drought events.
+This outcome highlights the benefit of a diversified energy mix in enhancing system resilience.
 
@@
 	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events.
-Wind droughts tend to be more frequent during summer, whereas PV droughts are more common in winter due to reduced sunlight.
+Wind droughts tend to be more frequent during summer, whereas solar PV droughts are more common in winter due to reduced sunlight.
 By comparing these seasonal patterns across different datasets and energy scenarios, the study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of drought events.
@@
 In Atlite, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
-
-In contrast, PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
-Across all datasets, over 60\% of winter hours are classified as PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
-Moreover, Atlite tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for PV relative to the C3S-E datasets.
+In contrast, solar PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
+Across all datasets, over 60\% of winter hours are classified as solar PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
+Moreover, Atlite tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for solar PV relative to the C3S-E datasets.
 These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
@@
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
-Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
+This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
+Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
 In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
 Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
-The balanced mix of wind and PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
+The balanced mix of wind and solar PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
 
@@
 	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, c)~91W-9PV and d)~57W-43PV for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
@@
 One of the most evident differences is how each dataset incorporates the specific locations of RES farms.
-Both Atlite and C3S-E G consider the locations of wind and PV farms, which one would expect to result in a more accurate representation of RES generation.
-While this approach slightly improves PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
+Both Atlite and C3S-E G consider the locations of wind and solar PV farms, which one would expect to result in a more accurate representation of RES generation.
+While this approach slightly improves solar PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
 This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES dataset is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
@@
 Atlite shows the best alignment with observed data for wind generation.
-Differences between the datasets are smaller for PV, with C3S-G performing marginally better than the other two.
+Differences between the datasets are smaller for solar PV, with C3S-G performing marginally better than the other two.
 The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact in this case.
@@
 
-Looking at the 57W-43PV scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation.
-Wind and PV together perform better in terms of reducing drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
-This diversification reduces the seasonal impact on RES droughts, as PV generation peaks in the summer and wind generation is more consistent in winter.
-Ireland currently has a highly wind-dependent energy system, but with ambitious targets for PV installations in the coming years, the energy mix is expected to approach a balance between wind and PV capacity.
-While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and PV capacities may not optimise other aspects, such as annual energy production or meeting nighttime loads.
+Looking at the 57W-43PV scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and solar PV generation.
+Wind and solar PV together perform better in terms of reducing drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
+This diversification reduces the seasonal impact on RES droughts, as solar PV generation peaks in the summer and wind generation is more consistent in winter.
+Ireland currently has a highly wind-dependent energy system, but with ambitious targets for solar PV installations in the coming years, the energy mix is expected to approach a balance between wind and solar PV capacity.
+While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and solar PV capacities may not optimise other aspects, such as annual energy production or meeting nighttime loads.
 For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.
@@
 This study has several limitations.
-Although ERA5 is among the best reanalysis datasets for renewable energy analysis \citep{doi:10.3390/atmos12050624} ?, its resolution may not capture local-scale phenomena, making it less reliable at the individual farm level.
-In addition, our methodology employs a fixed threshold to define RES drought events, which is necessary to compare the three models but does not account for demand variations.
+Although ERA5 is among the best reanalysis datasets for renewable energy analysis, its resolution may not capture local-scale phenomena, making it less reliable at the individual farm level.
+In addition, previous studies have indicated biases in ERA5 variables especially wind speed.
+Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three models but does not account for demand variations.
 Consequently, while this approach enables a consistent inter-comparison, it may overlook events that are most critical for power system operations.
@@
 The C3S-E dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
-Information on wind and PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).
+Information on wind and solar PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).
 The Atlite model used in this study is open-source and can be found on GitHub (\url{https://github.com/pypsa/atlite}).

### 2025-03-21 Article/RES_droughts_Morin_ReEn_R2.tex
Make some changes for the review

@@
 
-This study focuses on Ireland, a region with a strong reliance on wind power, which has ambitious targets for solar PV power expansion.
-This case study provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts.
-The performance of different RES datasets are compared, and a 45-year time series of RES generation is produced.
-The results highlight the role of increased solar PV capacity in reducing RES drought risks, offering insights for policymakers and energy planners.
-
-Various approaches have been proposed in the literature to define and quantify RES drought events.
+RES drought events do not have a fixed definition, with various approaches present in the literature.
 One common method defines a RES drought as a period during which the average capacity factor (CF) remains below a fixed threshold for a specified duration.
-For example, Kaspar et al.~\citep{kaspar2019drought} investigated the shortfall risks of low wind and solar PV generation in Europe, focusing on Germany, while Mockert et al.~\citep{mockert2023drought} extended this work to examine the link between weather regimes and RES droughts.
-Similar fixed-threshold approaches have also been applied using machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
+For example, Kaspar et al.~\citep{kaspar2019drought} used this method to investigate the shortfall risks of low wind and solar PV generation in Europe, with a focus on Germany, testing multiple CF thresholds and durations.
+Similarly, Mockert et al.~\citep{mockert2023drought} examined the link between weather regimes and RES droughts in Germany using a 48-hour rolling window under a threshold to define RES droughts.
+Similar fixed-threshold approaches have also been applied using CF series reconstructed through machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
 
@@
 Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models.
-Other studies have defined droughts based on deviations from daily mean production, as done by Rinaldi et al.~\citep{rinaldi2021drought} in the U.S.
-Western Interconnection, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences.
+Other studies have defined droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S.
+Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe drought events.
 Another method defines energy drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
 This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
-It has been applied in other studies, including analyses of RES droughts in the U.S.~\citep{bracken2024drought} and China~\citep{lei2024drought}.
+Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.
+~\citep{lei2024drought} used it to quantify droughts in wind-PV-hydro systems in China.
 
-In addition to examining periods of low renewable electricity generation, Raynaud et al.~\citep{raynaud2018drought} analysed the imbalance between electricity demand and renewable generation, known as residual load.
-These events were studied alongside low-generation periods to assess their correlation.
-Similar analyses have been conducted in Europe~\citep{allen2023drought} and the U.S.~\citep{bracken2024drought}, revealing differing results across regions.
+In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual load) is high.
+Raynaud et al.~\citep{raynaud2018drought} defined both energy production and energy supply droughts, and showed the difference in their patterns in a hypothetical fully renewable system composed of wind, PV and run-of-the-river hydropower.
+Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the energy production index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
+This index was also applied to the U.S.
+by Bracken et al~\citep{bracken2024drought}, revealing a consistent increase in the drought magnitude when load is considered, despite showing differing results across regions.
 
-In this paper, the focus is exclusively on renewable electricity generation, and a fixed threshold approach to define RES droughts is used, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
+In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand, whose role will be addressed in the discussion.
+A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
+The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion.
+This provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts in the context of realistic scenarios.
 
 RES droughts are identified using onshore wind and solar PV CF time series.
-In this study, three different datasets are used, all of which are driven by ERA5 data~\citep{hersbach2020era5}.
+In this study, three different datasets are used and compared, all of which are driven by the ERA5 reanalysis~\citep{hersbach2020era5}.
 Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
@@
 
-The aim of this study is to answer two questions which could help on the decision making for the planning of reserve capacity in real case wind-dominated renewable energy system:
+Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
+Despite most of them undergoing a validation process, they are often not fully representative of all geographical locations, and even show strong differences between each other \citep{kies2021drought}.
+In this work, we quantify the skill of a generic model for Ireland (a region not commonly used for generic model validation), and explore the effect that using it has on RES droughts when compared to a specifically-designed model.
+This comparison is propagated through the whole analysis to fully see the effect of these differences in the context of a transition from a wind-only system to a more balanced mix that includes solar PV.
+In terms of the drought definition, the literature often uses daily averaged values, which limits the start and end times of events.
+We opt for a 24-hour rolling average, which avoids potential masking of day-long events due to their start time.
+
+Therefore, the aim of this study is to answer three questions which could help on the decision making for the planning of reserve capacity in real case wind-dominated renewable energy system:
 \begin{itemize}
-	\item What is the impact of selecting different modelling assumptions on the analysis of RES droughts?
+	\item Can generic datasets be used to quantify extreme events like RES droughts?
+	\item What is the impact of modelling assumptions on the analysis of RES droughts?
 	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts in a real-case setting?
@@
 
-This study uses three datasets to analyse RES droughts across the island of Ireland.
+This study uses onshore wind and solar PV CF time series from three datasets to analyse RES droughts across the island of Ireland.
 Data downloaded from C3S-E were used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
@@
 
-\subsection{C3S-Energy National}
+\subsection{C3S-Energy National: C3S NAT}
 \label{sec:c3se_n}
 
-For national-level analyses, the inputs for this dataset are the aggregated CF time series provided by C3S-E at two levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+The C3S NAT dataset is created by combining different national and regional data sources.
+The inputs used are the total installed capacity in the Republic of Ireland and Northern Ireland, and the aggregated CF time series provided by C3S-E at the two corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
 These values are based on the assumption that RES generation occurs at every ERA5 grid point in Ireland.
-A weighted average is computed to represent the total CF for the country.
+To get the C3S NAT CF for all of Ireland, a weighted average of the two obtained CF time series is performed, using the actual installed capacity as weights.
 
-\subsection{C3S-E Gridded}
+\subsection{C3S-Energy Gridded: C3S GRID}
 \label{sec:c3se_g}
 
-For the gridded dataset, the inputs consist of the CF time series from C3S-E, along with the location of individual RES farms across Ireland.
-Using these inputs, the nearest grid point on the C3S-E dataset was identified for each farm, and the corresponding CF values were retrieved.
-A weighted average, based on installed capacity, was then computed to construct a CF time series that accounts for the actual spatial distribution of wind and solar PV farms in Ireland.
+The C3S GRID dataset combines information that contains the spatial variability over Ireland.
+The inputs used to create this dataset consist of the CF time series from C3S-E over the ERA5 grid, along with the location and characteristics of individual RES farms across Ireland, as explained in section~\ref{sec:eirgrid}.
+For each farm, the nearest grid point on the C3S-E dataset was identified, the generation from that farm was calculated using the retrieved CF from C3S-Energy, and it was added to a total generation.
+This total generation was divided by the total installed capacity to convert it back to CF.
+This is equivalent to performing a weighted average of the CF associated with each farm using the farm's installed capacities as weights.
+The C3S GRID dataset contains the resulting CF time series, which accounts for the actual spatial distribution of wind and solar PV farms in Ireland.
 
-\subsection{Atlite} 
+\subsection{Atlite: ATLITE} 
 \label{sec:atlite}
 
-For the Atlite dataset, the inputs include the locations of RES farms and ERA5 weather variables, such as wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
-The meteorological inputs are processed using the Atlite model to estimate CF time series for wind and solar PV, incorporating specific characteristics such as wind turbines power curve and PV panel model.
-A key distinction between C3S-E and Atlite lies in their representation of wind turbines and PV panels.
-This study identifies the most appropriate wind turbine power curve to use from the 121 power curves made available by Renewables.ninja~\citep{staffell2016wake}.
+The ATLITE dataset is constructed by using the Atlite model to process weather variables into energy variables.
+This model allows for a tuning process where the models used to transform wind and solar PV can be adjusted to best represent the observed data up until this point for the relevant region (in this case, Ireland).
+The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
+These meteorological inputs are processed using the Atlite model to estimate CF time series for wind and solar PV, incorporating specific characteristics such as the wind turbine power curve and PV panel model selected for this specific case.
+The output of the Atlite model is a generation time series, which divided by the total capacity to transform it back into CF, which is what we call the ATLITE dataset.
+The flexibility in the usage of different power curves and PV panel models represents the key distinction between this dataset and the two C3S-derived ones.
+This study identifies the most appropriate wind turbine power curve to use from the 121 power curves at five different levels of smoothing made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within the Atlite model.
 The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
@@
 \begin{itemize}
-	\item Wind Power / 91W-9PV
-	\item Solar PV Power / 91W-9PV
-	\item Combined RES / 91W-9PV
-	\item Combined RES / 57W-43PV
+	\item Wind Power - based on the actual capacity at the end of 2023
+	\item Solar PV Power - based on the actual capacity at the end of 2023
+	\item Combined RES / 91W-9PV - based on the actual capacity at the end of 2023
+	\item Combined RES / 57W-43PV - based on the capacity projected for 2030
 \end{itemize}
@@
 It is important to highlight that this analysis considers two key aspects: the absolute values that characterise RES droughts, which are crucial for power system planning, and the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.
+This complete analysis puts forward the differences between the datasets, showing the impact of possible misrepresentations of the system on the analysis of RES drought events.
 

### 2025-03-21 Article/RES_droughts_Morin_ReEn_R2.tex
Added some more comments in response to reviewers

@@
 
+The verification of wind generation has put forward two main facts.
+On the one hand, the ATLITE dataset is skilled at reproducing onshore wind CF and droughts over Ireland .On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
+This puts forward one of the problems that come with using generalised datasets for the characterisation of RES droughts: biases severely affect the ability of any given model to reproducing them.
+The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in the ability to reproduce the changes in CF (visible in the very similar CC).
+However, their ability to reproduce the exact value is much lower than that of ATLITE (RMSE almost 4 times bigger for the two C3S with respect to ATLITE), and particularly biased towards an overestimation of CF (clear in the MBE values), which leads to the underestimation of droughts.
+This puts forward the need to use fully verified models to assess RES droughts, whose impact on the analysis of RES droughts will be explored in section~\ref{sec:analysis}.
+
 \subsubsection{Solar PV Energy}
@@
 
-As the goal of this analysis is to assess the combination of wind and solar PV generation, the complementary nature of these energy sources mitigates the limitations in solar PV-only results.
-
 \begin{figure}[!ht]
@@
 \end{figure}
+
+As one of the goals of this analysis is to assess the combination of wind and solar PV generation, the complementary nature of these energy sources will mitigate the limitations in solar PV-only results.
+Still, a few differences between wind and solar PV stand out.
+First, no single model has proven best for the characterisation of solar PV, although the statistical approach taken by the two C3S datasets has proven superior to the single panel model in ATLITE (Tab.~\ref{tab:pv_skill_scores}).
+The short time considered in the verification due to the limited data availability for solar PV in Ireland is part of the cause for our limited ability to select a specific overall best solar PV dataset.
+However, our results seem to indicate that it is possible that the generic datasets perform better for solar PV than for wind, as the variability between the different panel models is smaller in comparison to the existing differences between wind turbine power curves.
+Therefore, all through section~\ref{sec:analysis} the differences between the three models will be highlighted, considering that neither stands out positively.
 
@@
 This observation is consistent with previous research, which has also noted that assumptions regarding turbine characteristics can significantly affect drought duration estimates.
+Despite the differences in the number of RES droughts detected by each dataset, the overall effect of balancing the shares of wind and solar PV is consistent, independently of the complexity of the dataset used.
 
@@
 Additionally, in all four graphs, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
+
+Contrary to the number of events per year and per duration, the return periods show large differences in the behaviours of the three datasets.
+The reproduction of average behaviours is still manageable using the C3S datasets, but the magnitude of extremes is not well reproduced when the share of wind is large.
+System planning based on the wrong datasets could yield an underestimation of the potential for extreme RES droughts, eventually leading to shortages linked to undersized reserve capacity.
+Even though this effect will be mitigated, affecting only the most extreme events with the introduction of widespread solar PV, it is important to keep in mind this effect for all wind-dominated electricity systems.
 
@@
 Ireland currently has a highly wind-dependent energy system, but with ambitious targets for solar PV installations in the coming years, the energy mix is expected to approach a balance between wind and solar PV capacity.
-While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and solar PV capacities may not optimise other aspects, such as annual energy production or meeting nighttime loads.
+While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and solar PV capacities may not optimise other aspects, such as annual energy production or meeting night-time loads.
 For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.

### 2025-03-22 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 This study investigates the importance of the balance between wind and solar photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
-Three different RES models are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
+Three different RES datasets are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
 The skill of the RES models is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
@@
 In this study, three different datasets are used and compared, all of which are driven by the ERA5 reanalysis~\citep{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3S-E), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
+Two of the datasets are part of C3S Energy (C3SE), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
 One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
@@
 This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV power.
-The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S-E datasets.
+The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S datasets.
 
@@
 
-Atlite and C3S-E datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+ATL and the C3S datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
 It has proven to be the best choice for studying renewable energy in Ireland \citep{doddy2021era5}.
-Table~\ref{tab:var_name} lists the ERA5 variables used by Atlite and C3S-Energy.
+Table~\ref{tab:var_name} lists the ERA5 variables used to produce the ATL and the C3S datasets.
 
@@
 
-The EU Copernicus Climate Change Service developed the C3S-E renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
+The EU Copernicus Climate Change Service developed the C3S-Energy renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
 This dataset provides hourly CF for wind and solar PV energy from 1979 to the present.
@@
 
-The wind CF in the C3S-E model is calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
+The wind CF in the C3SE model is calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
 This choice reflects trends in wind turbine installations and was guided by expert recommendations.
-Since real-time data on the exact wind turbine fleet across Europe is difficult to obtain, C3S-E assumes a homogeneous distribution of turbines across the ERA5 grid.
+Since real-time data on the exact wind turbine fleet across Europe is difficult to obtain, C3SE assumes a homogeneous distribution of turbines across the ERA5 grid.
 While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of climate variability on wind power generation.
 The turbine power curves used in the model are sourced from publicly available databases, ensuring consistency with industry standards.
-The solar PV CF in the C3S-E model is calculated at the grid level and represents the aggregated output of all solar PV systems within each pixel, rather than a single installation.
+The solar PV CF in the C3SE model is calculated at the grid level and represents the aggregated output of all solar PV systems within each pixel, rather than a single installation.
 It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference solar PV plant model.
@@
 This study uses onshore wind and solar PV CF time series from three datasets to analyse RES droughts across the island of Ireland.
-Data downloaded from C3S-E were used to obtain two datasets: one based on national-level data (C3S-E N), and another on grid-level data (C3S-E G).
-The third dataset was computed using the Atlite model (Atlite).
+Data downloaded from C3SE were used to obtain two datasets: one based on national-level data (C3S NAT), and another on grid-level data (C3S GRD).
+The third dataset was computed using the Atlite model (ATL).
 
-\subsection{C3S-Energy National: C3S NAT}
+\subsection{C3S Energy National: C3S NAT}
 \label{sec:c3se_n}
@@
 The C3S NAT dataset is created by combining different national and regional data sources.
-The inputs used are the total installed capacity in the Republic of Ireland and Northern Ireland, and the aggregated CF time series provided by C3S-E at the two corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+The inputs used are the total installed capacity in the Republic of Ireland and Northern Ireland, and the aggregated CF time series provided by C3SE at the two corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
 These values are based on the assumption that RES generation occurs at every ERA5 grid point in Ireland.
@@
 
-\subsection{C3S-Energy Gridded: C3S GRID}
+\subsection{C3S Energy Gridded: C3S GRID}
 \label{sec:c3se_g}
@@
 The C3S GRID dataset combines information that contains the spatial variability over Ireland.
-The inputs used to create this dataset consist of the CF time series from C3S-E over the ERA5 grid, along with the location and characteristics of individual RES farms across Ireland, as explained in section~\ref{sec:eirgrid}.
-For each farm, the nearest grid point on the C3S-E dataset was identified, the generation from that farm was calculated using the retrieved CF from C3S-Energy, and it was added to a total generation.
+The inputs used to create this dataset consist of the CF time series from C3SE over the ERA5 grid, along with the location and characteristics of individual RES farms across Ireland, as explained in section~\ref{sec:eirgrid}.
+For each farm, the nearest grid point on the C3SE dataset was identified, the generation from that farm was calculated using the retrieved CF from C3SE, and it was added to a total generation.
 This total generation was divided by the total installed capacity to convert it back to CF.
@@
 
-\subsection{Atlite: ATLITE} 
+\subsection{Atlite: ATL} 
 \label{sec:atlite}
 
-The ATLITE dataset is constructed by using the Atlite model to process weather variables into energy variables.
-This model allows for a tuning process where the models used to transform wind and solar PV can be adjusted to best represent the observed data up until this point for the relevant region (in this case, Ireland).
+The ATL dataset is constructed by using the Atlite model to process weather variables into energy variables.
+This model allows for a tuning process where the model used to transform wind and solar PV can be adjusted to best represent the observed data up until this point for the relevant region (in this case, Ireland).
 The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
 These meteorological inputs are processed using the Atlite model to estimate CF time series for wind and solar PV, incorporating specific characteristics such as the wind turbine power curve and PV panel model selected for this specific case.
-The output of the Atlite model is a generation time series, which divided by the total capacity to transform it back into CF, which is what we call the ATLITE dataset.
-The flexibility in the usage of different power curves and PV panel models represents the key distinction between this dataset and the two C3S-derived ones.
+The output of the Atlite model is a generation time series, which divided by the total capacity to transform it back into CF, which is what we call the ATL dataset.
+The flexibility in the usage of different power curves and PV panel models represents the key distinction between this dataset and the two C3SE-derived ones.
 This study identifies the most appropriate wind turbine power curve to use from the 121 power curves at five different levels of smoothing made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within the Atlite model.
@@
 
-New time series were generated for both the Atlite and C3S-E G solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the ATL and C3S GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
@@
 
-For each dataset (Atlite, C3S-E G, and C3S-E N), four distinct scenarios are examined, as summarised below:
+For each dataset (ATL, C3S GRID, and C3S NAT), four distinct scenarios are examined, as summarised below:
 
@@
 
-The C3S-E datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve}a).
+The C3S datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve}a).
 The Atlite model allows the user to specify the power curve.
@@
 	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (blue) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (blue) and Observed (shaded)}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by ATL (orange) and the Vestas V136/3450 used by C3SE (blue) b) Histograms of wind CF for Ireland from ATL (orange), C3SE (blue) and Observed (shaded)}
 	\label{fig:power_curve}
@@
 The smoothing of the wind turbine power curve represents losses associated with each turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
-The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the observed wind availability data.
-This is further supported by the percentage of overlap which is higher for Atlite (97.2\%) than for C3S-E (83.2\%), indicating better agreement with observed data.
+The histogram in Fig.~\ref{fig:power_curve}b shows that the C3SE power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed ATL power curve more closely follows the observed wind availability data.
+This is further supported by the percentage of overlap which is higher for ATL (97.2\%) than for C3S (83.2\%), indicating better agreement with observed data.
 
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3SE GRID and c) C3S NAT datasets}
 	\label{fig:wind_verification_contour}
@@
 The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
-The two C3S-E datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
-The skill scores presented in Table~\ref{tab:wind_skill_scores} show that Atlite performs better than the C3S-E datasets for all of the skill scores.
+The two C3SE datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
+The skill scores presented in Table~\ref{tab:wind_skill_scores} show that ATL performs better than the two C3S datasets for all of the skill scores.
 
@@
 		\cline{2-4}
-		& \textbf{Atlite} & \textbf{C3S-E G} & \textbf{C3S-E N} \\ \hline
+		& \textbf{ATL} & \textbf{C3S GRID} & \textbf{C3S NAT} \\ \hline
 		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.981           & 0.972            & 0.970            \\ \hline
@@
 Fig.~\ref{fig:bar_number_events_verification_wind} shows the average annual number of wind drought events during the 2014 to 2023 validation period.
-The figure reveals that Atlite presents the best overall agreement with the observed frequency and duration of wind drought events.
+The figure reveals that ATL presents the best overall agreement with the observed frequency and duration of wind drought events.
 This pattern is particularly evident for shorter-duration events, which are the most frequent.
@@
 	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline).
+	\caption{Average annual number of wind drought events for ATL (red), C3S GRD (blue), C3S NAT (purple), and the observed data (black outline).
 The wind droughts are identified from 2014 to 2023, considering the actual capacity of the system at any given time}
@@
 The verification of wind generation has put forward two main facts.
-On the one hand, the ATLITE dataset is skilled at reproducing onshore wind CF and droughts over Ireland .On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
+On the one hand, the ATL dataset is skilled at reproducing onshore wind CF and droughts over Ireland.
+On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
 This puts forward one of the problems that come with using generalised datasets for the characterisation of RES droughts: biases severely affect the ability of any given model to reproducing them.
 The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in the ability to reproduce the changes in CF (visible in the very similar CC).
-However, their ability to reproduce the exact value is much lower than that of ATLITE (RMSE almost 4 times bigger for the two C3S with respect to ATLITE), and particularly biased towards an overestimation of CF (clear in the MBE values), which leads to the underestimation of droughts.
+However, their ability to reproduce the exact value is much lower than that of ATL (RMSE almost 4 times bigger for the two C3S with respect to ATL), and particularly biased towards an overestimation of CF (clear in the MBE values), which leads to the underestimation of droughts.
 This puts forward the need to use fully verified models to assess RES droughts, whose impact on the analysis of RES droughts will be explored in section~\ref{sec:analysis}.
@@
 	\includegraphics[width=\textwidth]{verification_pv_contour.png}
-	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N datasets}	
+	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) ATL, b) C3S GRD and c) C3S NAT datasets}	
 	\label{fig:solar_verification_contour}
@@
 Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
-The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S-E~G performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
-All models show a slight positive bias, with Atlite exhibiting a slightly lower correlation and higher RMSE.
+The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
+All models show a slight positive bias, with ATL exhibiting a slightly lower correlation and higher RMSE.
 
@@
 		\cline{2-4}
-		& \textbf{Atlite} & \textbf{C3S-E G} & \textbf{C3S-E N} \\ \hline
+		& \textbf{ATL} & \textbf{C3S GRD} & \textbf{C3S NAT} \\ \hline
 		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.931            \\ \hline
@@
 	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
-	\caption{Number of solar PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline).
+	\caption{Number of solar PV drought events for ATL (red), C3S GRD (blue), and C3S NAT (purple) and the observed data (black outline).
 The solar PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
@@
 Still, a few differences between wind and solar PV stand out.
-First, no single model has proven best for the characterisation of solar PV, although the statistical approach taken by the two C3S datasets has proven superior to the single panel model in ATLITE (Tab.~\ref{tab:pv_skill_scores}).
+First, no single model has proven best for the characterisation of solar PV, although the statistical approach taken by the two C3S datasets has proven superior to the single panel model in ATL (Tab.~\ref{tab:pv_skill_scores}).
 The short time considered in the verification due to the limited data availability for solar PV in Ireland is part of the cause for our limited ability to select a specific overall best solar PV dataset.
@@
 When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, reaffirming wind’s dominance in the current energy mix.
-However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for Atlite, 52\% for C3S-E G, and 50\% for C3S-E N, demonstrating the beneficial effects of a more balanced energy mix.
+However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
 These findings are in line with earlier studies that highlight how increasing solar PV capacity can mitigate drought frequency through the anti-correlated seasonal patterns of wind and solar generation.
 
-Additionally, the consistently higher drought counts reported by the Atlite dataset, compared to the C3S-E datasets, underscore the impact of model selection, particularly the influence of wind turbine power curve representation, on quantifying RES droughts.
+Additionally, the consistently higher drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the impact of model selection, particularly the influence of wind turbine power curve representation, on quantifying RES droughts.
 This observation is consistent with previous research, which has also noted that assumptions regarding turbine characteristics can significantly affect drought duration estimates.
@@
 	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red), C3S-E G (blue), and C3S-E N (purple).
+	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red), C3S GRD (blue), and C3S NAT (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 For wind (Fig.~\ref{fig:return_periods}a), the log‐linear increase in return periods observed in this study confirms that longer droughts occur exponentially less frequently, a trend consistent with earlier research on wind variability.
-In the case of solar PV droughts (Fig.\ref{fig:return_periods}b), the Atlite dataset shows a general log‐linear trend, whereas the C3S-E datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
+In the case of solar PV droughts (Fig.\ref{fig:return_periods}b), the ATL dataset shows a general log‐linear trend, whereas the C3S datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
 This abrupt rise reflects differences in how solar PV output is handled near the CF threshold during low irradiance conditions.
+
 In the balanced scenario, the reduced share of wind and increased share of solar PV leverages their complementary seasonal patterns, resulting in higher return periods for combined drought events.
@@
 
-Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods largely mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
-In contrast, the balanced 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
+Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
+In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
 
@@
 	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red triangle), C3S GRD (blue circle), and C3S NAT (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 
-Across Fig.~\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets.
-For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets.
+Across Fig.~\ref{fig:return_periods}a, c, and, d, the return periods in the ATL dataset are consistently higher than those in the two C3S datasets.
+For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the ATL dataset, compared to only five days in the C3S datasets.
 This difference underscores the importance of model selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence drought duration estimates.
-Additionally, in all four graphs, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
+Additionally, in all four graphs, the similarity between results from the two C3S datasets suggests that assumptions in the ATL dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
 
@@
 
-For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the Atlite dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February).
-This strong seasonal signal is less evident in the C3S-E datasets, which suggests that the differences in the underlying wind power curves play a significant role.
-In Atlite, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
+For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February).
+This strong seasonal signal is less evident in the C3S datasets, which suggests that the differences in the underlying wind power curves play a significant role.
+In ATL, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
 In contrast, solar PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
 Across all datasets, over 60\% of winter hours are classified as solar PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
-Moreover, Atlite tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for solar PV relative to the C3S-E datasets.
+Moreover, ATL tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
 These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
@@
 In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S GRD, and from 9\% to 4\% for C3S-E NAT.
 The balanced mix of wind and solar PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
@@
 	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red dotted), C3S GRD (blue dashed), and C3S NAT (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
@@
 
-This study has investigated the ability of three RES datasets to represent RES droughts: Atlite, C3S-E G, and C3S-E N.
-One of the most evident differences is how each dataset incorporates the specific locations of RES farms.
-Both Atlite and C3S-E G consider the locations of wind and solar PV farms, which one would expect to result in a more accurate representation of RES generation.
-While this approach slightly improves solar PV models, our analysis indicates that for wind energy, the Atlite dataset performs better overall, especially in its close alignment with observed data for wind generation estimates.
-This finding suggests that, although the inclusion of RES farm locations is beneficial, the accuracy of the RES dataset is more strongly influenced by underlying model assumptions, such as selecting an appropriate wind power curve.
+The first question this study aimed to answer was whether generic datasets can be used to quantify extreme events like RES droughts.
+The analysis compared three datasets, ATL, C3S GRD, and C3S NAT, over a 45‐year period to determine their ability to represent RES drought events.
+The results show that while all three datasets capture overall trends in drought occurrence, significant differences emerge when using non-tailored data for the study of RES droughts.
+This finding highlights that the choice of dataset and its underlying assumptions can lead to consequential differences in estimated RES drought characteristics, emphasising the need for datasets that are specifically designed for extreme event analysis.
 
-Atlite shows the best alignment with observed data for wind generation.
-Differences between the datasets are smaller for solar PV, with C3S-G performing marginally better than the other two.
-The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact in this case.
-This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with a balanced share of RES.
-Again, the results from RES drought modelling rely more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms.
-Atlite’s superior performance highlights the importance of selecting validated models for assessing RES drought risks.
-This careful model selection can better quantify risks, support effective planning, and avoid the potential underestimation of capacity needs, which is essential for ensuring energy security.
+The second key question addressed the impact of modeling assumptions on RES drought analysis.
+The study revealed that differences in model parametrisation, particularly in the representation of wind turbine power curves and solar PV panel characteristics, have a stronger influence on the drought estimates than the inclusion of RES farm locations.
+For instance, the ATL dataset, which uses a carefully selected wind turbine power curve from Renewables.ninja, consistently produced higher return periods and a greater number of drought events for wind energy than the C3S datasets.
+This suggests that fine-tuning model parameters to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
 
-Looking at the 57W-43PV scenario, the analysis showed a significant improvement in the management of RES droughts due to the complementary nature of wind and solar PV generation.
-Wind and solar PV together perform better in terms of reducing drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources.
-This diversification reduces the seasonal impact on RES droughts, as solar PV generation peaks in the summer and wind generation is more consistent in winter.
-Ireland currently has a highly wind-dependent energy system, but with ambitious targets for solar PV installations in the coming years, the energy mix is expected to approach a balance between wind and solar PV capacity.
-While this balanced approach offers a more stable and secure energy supply by mitigating RES drought risks, it is important to note that having similar wind and solar PV capacities may not optimise other aspects, such as annual energy production or meeting night-time loads.
-For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification.
-Additionally, the choice of model for RES drought assessment becomes increasingly critical as more renewable capacity is integrated into the system.
+The third question explored how the integration of solar PV into a predominantly wind-based system alters the characteristics of RES droughts in a real-case setting.
+In Ireland, where the existing energy mix (91W-9PV) is highly wind-dominated, the analysis demonstrated that transitioning to a more balanced system (57W-43PV) markedly reduces the frequency, duration, and seasonal variability of drought events.
+This improvement is attributed to the complementary nature of wind and solar PV generation—solar PV typically peaks in summer while wind generation is more consistent in winter.
+Thus, a more diversified renewable energy mix not only mitigates extreme drought conditions but also enhances overall system resilience, providing valuable insights for policymakers tasked with ensuring energy security.
 
@@
 The ERA5 data can be obtained from the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
-The C3S-E dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
+The C3S datasets are also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
 Information on wind and solar PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).

### 2025-03-24 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 
-In the balanced scenario, the reduced share of wind and increased share of solar PV leverages their complementary seasonal patterns, resulting in higher return periods for combined drought events.
-This outcome highlights the benefit of a diversified energy mix in enhancing system resilience.
-
 Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
 In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
+Despite the lower wind share in the 57W-43PV scenario, typically known for its relative stability, the balanced share with solar PV leads to extended return periods for RES droughts.
+This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of drought events in a diversified energy portfolio.
 

### 2025-03-25 Article/RES_droughts_Morin_ReEn_R2.tex
Address everything before the conclusions

@@
 	\item Combined RES / 91W-9PV - based on the actual capacity at the end of 2023
-	\item Combined RES / 57W-43PV - based on the capacity projected for 2030
+	\item Combined RES / 57W-43PV - based on the projected capacity for 2030
 \end{itemize}
@@
 
-The C3S datasets use the Vestas V136/3450 wind turbine power curve, (Fig.~\ref{fig:power_curve}a).
+The C3S datasets use the Vestas V136/3450 wind turbine power curve (Fig.~\ref{fig:power_curve}a).
 The Atlite model allows the user to specify the power curve.
@@
 The histogram in Fig.~\ref{fig:power_curve}b shows that the C3SE power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed ATL power curve more closely follows the observed wind availability data.
-This is further supported by the percentage of overlap which is higher for ATL (97.2\%) than for C3S (83.2\%), indicating better agreement with observed data.
+This is further supported by the percentage of overlap which is higher for ATL (97.2\%) than for C3SE (83.2\%), indicating better agreement with observed data.
 
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3SE GRID and c) C3S NAT datasets}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S GRID and c) C3S NAT datasets}
 	\label{fig:wind_verification_contour}
@@
 The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
-The two C3SE datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
+The two C3S datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
 The skill scores presented in Table~\ref{tab:wind_skill_scores} show that ATL performs better than the two C3S datasets for all of the skill scores.
@@
 
-For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June–July–August) identified as droughts compared to only 4\% in winter (December–January–February).
+For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June, July, August) identified as droughts compared to only 4\% in winter (December, January, February).
 This strong seasonal signal is less evident in the C3S datasets, which suggests that the differences in the underlying wind power curves play a significant role.
@@
 The findings underscore that even small differences in model assumptions leads to significant variations in drought estimates, thereby affecting the reliability of the energy system during critical periods.
+Additionally, the integration of large shares of solar PV in the system leads to a general reduction of RES droughts year-round.
+However, it can not counter the presence of RES droughts in the winter months due to its natural limitations, inevitably leading to higher reserve capacity needs during winter months as reliance on RES increases.
 Such insights are essential for policymakers to develop targeted strategies that enhance grid resilience and ensure a stable energy supply throughout the year.
@@
 Although ERA5 is among the best reanalysis datasets for renewable energy analysis, its resolution may not capture local-scale phenomena, making it less reliable at the individual farm level.
-In addition, previous studies have indicated biases in ERA5 variables especially wind speed.
+In addition, previous studies have indicated biases in ERA5 variables, especially wind speed.
 Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three models but does not account for demand variations.

### 2025-03-26 Article/RES_droughts_Morin_ReEn_R2.tex
Update RES_droughts_Morin_ReEn_R2.tex
Add modifications to answer reviewer comments

@@
 
-The first question this study aimed to answer was whether generic datasets can be used to quantify extreme events like RES droughts.
-The analysis compared three datasets, ATL, C3S GRD, and C3S NAT, over a 45‐year period to determine their ability to represent RES drought events.
-The results show that while all three datasets capture overall trends in drought occurrence, significant differences emerge when using non-tailored data for the study of RES droughts.
+This study has explored the characterisation of RES droughts in the real-life transition from a wind-dominated system to a more balanced system with integrated solar PV.
+This has been done through the comparison over a 45-year period of three different datasets: one based on a hand-crafted validated model and two based on a generic model, C3S-Energy.
+The generic model has two versions, one using large-scale aggregated information, and one which includes the locations of farms as well (also considered in the validated model).
+
+Our results show that generic datasets present limitations in the quantification of RES droughts.
+Although the three datasets used capture overall trends in drought occurrence, significant differences emerge when using non-tailored data for the study of RES droughts.
 This finding highlights that the choice of dataset and its underlying assumptions can lead to consequential differences in estimated RES drought characteristics, emphasising the need for datasets that are specifically designed for extreme event analysis.
 
-The second key question addressed the impact of modeling assumptions on RES drought analysis.
-The study revealed that differences in model parametrisation, particularly in the representation of wind turbine power curves and solar PV panel characteristics, have a stronger influence on the drought estimates than the inclusion of RES farm locations.
-For instance, the ATL dataset, which uses a carefully selected wind turbine power curve from Renewables.ninja, consistently produced higher return periods and a greater number of drought events for wind energy than the C3S datasets.
+This study reveals that differences in model parametrisation, particularly in the representation of wind turbine power curves and solar PV panel characteristics, have a stronger influence on the drought estimates than the inclusion of RES farm locations.
+The use of a validated dataset with a carefully selected wind turbine power curve consistently produced higher return periods and a greater number of drought events for wind energy than the datasets derived from C3S-Energy.
 This suggests that fine-tuning model parameters to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
 
-The third question explored how the integration of solar PV into a predominantly wind-based system alters the characteristics of RES droughts in a real-case setting.
-In Ireland, where the existing energy mix (91W-9PV) is highly wind-dominated, the analysis demonstrated that transitioning to a more balanced system (57W-43PV) markedly reduces the frequency, duration, and seasonal variability of drought events.
-This improvement is attributed to the complementary nature of wind and solar PV generation—solar PV typically peaks in summer while wind generation is more consistent in winter.
+Finally, the effect on RES droughts of the integration of solar PV in a wind-dominated system has been explored in a real-case setting.
+In the presented example of Ireland, the analysis has demonstrated that transitioning to a more balanced system with similar amounts of wind and solar PV markedly reduces the frequency, duration, and seasonal variability of drought events.
+This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV typically peaks in summer while wind generation is more consistent in winter.
 Thus, a more diversified renewable energy mix not only mitigates extreme drought conditions but also enhances overall system resilience, providing valuable insights for policymakers tasked with ensuring energy security.
 
-This study has several limitations.
-Although ERA5 is among the best reanalysis datasets for renewable energy analysis, its resolution may not capture local-scale phenomena, making it less reliable at the individual farm level.
-In addition, previous studies have indicated biases in ERA5 variables, especially wind speed.
-Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three models but does not account for demand variations.
-Consequently, while this approach enables a consistent inter-comparison, it may overlook events that are most critical for power system operations.
+The results presented in this study have several limitations.
+Although ERA5 is among the best reanalysis datasets for renewable energy analysis, it still presents some biases, and its resolution may not capture local-scale phenomena.
+This is especially limiting if individual farms are considered, but its still shows good skill at statistically representing nation-wide behaviours considering the distribution of farms.
+Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three datasets considering only weather-derived drivers, but does not account for demand variations.
+Consequently, while this approach enables a consistent inter-comparison, it will not signal events that are most critical for power system operations for their mismatch of RES generation and demand.
 

### 2025-03-27 Article/RES_droughts_Morin_ReEn_R2.tex
Edits with Conor

@@
 Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
-Despite most of them undergoing a validation process, they are often not fully representative of all geographical locations, and even show strong differences between each other \citep{kies2021drought}.
-In this work, we quantify the skill of a generic model for Ireland (a region not commonly used for generic model validation), and explore the effect that using it has on RES droughts when compared to a specifically-designed model.
-This comparison is propagated through the whole analysis to fully see the effect of these differences in the context of a transition from a wind-only system to a more balanced mix that includes solar PV.
-In terms of the drought definition, the literature often uses daily averaged values, which limits the start and end times of events.
-We opt for a 24-hour rolling average, which avoids potential masking of day-long events due to their start time.
+Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events \citep{kies2021drought}.
+In this work, we quantify the skill of a dataset developed for the European region (C3SE), when used for a specific country (Ireland).
+In particular, we investigate the impact of a generic versus a tailored dataset on the analysis of RES droughts, in the context of a transition from a wind-dominated system to one with a larger share of solar PV.
 
-Therefore, the aim of this study is to answer three questions which could help on the decision making for the planning of reserve capacity in real case wind-dominated renewable energy system:
+The aim of this study is to answer three questions which are relevant for systems with a large share of RES generation:
 \begin{itemize}
-	\item Can generic datasets be used to quantify extreme events like RES droughts?
-	\item What is the impact of modelling assumptions on the analysis of RES droughts?
-	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts in a real-case setting?
+	\item Do generic datasets have sufficient skill to reliably quantify extreme RES drought events?
+	\item What is the importance of using accurate RES farm locations, and regionally-validated wind and solar PV models, when analysing of RES droughts?
+	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES drought events?
 \end{itemize}
@@
 
-ATL and the C3S datasets are driven by the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+All of the datasets used in this study are driven by data from the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
-It has proven to be the best choice for studying renewable energy in Ireland \citep{doddy2021era5}.
-Table~\ref{tab:var_name} lists the ERA5 variables used to produce the ATL and the C3S datasets.
+Table~\ref{tab:var_name} lists the relevant ERA5 variables.
 
@@
 
-The EU Copernicus Climate Change Service developed the C3S-Energy renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
+The EU Copernicus Climate Change Service developed the C3S-Energy (C3SE) renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
 This dataset provides hourly CF for wind and solar PV energy from 1979 to the present.
@@
 
-The wind CF in the C3SE model is calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
-This choice reflects trends in wind turbine installations and was guided by expert recommendations.
-Since real-time data on the exact wind turbine fleet across Europe is difficult to obtain, C3SE assumes a homogeneous distribution of turbines across the ERA5 grid.
+The wind CF in C3SE was calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
+As data on wind turbine fleet locations and specifications are difficult to obtain across Europe, C3SE assumes a homogeneous distribution of wind turbines across the ERA5 grid.
 While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of climate variability on wind power generation.
-The turbine power curves used in the model are sourced from publicly available databases, ensuring consistency with industry standards.
-The solar PV CF in the C3SE model is calculated at the grid level and represents the aggregated output of all solar PV systems within each pixel, rather than a single installation.
+The C3SE solar PV CF was also calculated for the ERA5 grid.
 It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference solar PV plant model.
@@
 
-The C3S NAT dataset is created by combining different national and regional data sources.
-The inputs used are the total installed capacity in the Republic of Ireland and Northern Ireland, and the aggregated CF time series provided by C3SE at the two corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
-These values are based on the assumption that RES generation occurs at every ERA5 grid point in Ireland.
-To get the C3S NAT CF for all of Ireland, a weighted average of the two obtained CF time series is performed, using the actual installed capacity as weights.
+The C3S NAT dataset is created by combining two inputs provided by C3SE at the corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+The two inputs are combined, using the actual installed capacity as weights.
+This dataset assumes that RES generation occurs at every ERA5 grid point in Ireland.
 
@@
 
-The C3S GRID dataset combines information that contains the spatial variability over Ireland.
-The inputs used to create this dataset consist of the CF time series from C3SE over the ERA5 grid, along with the location and characteristics of individual RES farms across Ireland, as explained in section~\ref{sec:eirgrid}.
-For each farm, the nearest grid point on the C3SE dataset was identified, the generation from that farm was calculated using the retrieved CF from C3SE, and it was added to a total generation.
-This total generation was divided by the total installed capacity to convert it back to CF.
-This is equivalent to performing a weighted average of the CF associated with each farm using the farm's installed capacities as weights.
-The C3S GRID dataset contains the resulting CF time series, which accounts for the actual spatial distribution of wind and solar PV farms in Ireland.
+The C3S GRID dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
+For each farm, the CF from the nearest grid point on the C3SE dataset was selected.
+A weighted average of the CF associated with each farm, using the farm's installed capacities, was used to produce the total CF time series.
 
@@
 
-The ATL dataset is constructed by using the Atlite model to process weather variables into energy variables.
-This model allows for a tuning process where the model used to transform wind and solar PV can be adjusted to best represent the observed data up until this point for the relevant region (in this case, Ireland).
+The ATL dataset is produced using the Atlite model.
+Atlite allows the user to define the wind turbine power curve and PV panel model to use when converting weather variables to wind and solar PV gerenation.
 The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
-These meteorological inputs are processed using the Atlite model to estimate CF time series for wind and solar PV, incorporating specific characteristics such as the wind turbine power curve and PV panel model selected for this specific case.
-The output of the Atlite model is a generation time series, which divided by the total capacity to transform it back into CF, which is what we call the ATL dataset.
-The flexibility in the usage of different power curves and PV panel models represents the key distinction between this dataset and the two C3SE-derived ones.
-This study identifies the most appropriate wind turbine power curve to use from the 121 power curves at five different levels of smoothing made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within the Atlite model.
-The selection of a specific wind turbine and PV panel characteristics is further discussed and explained in section \ref{sec:verification}.
+The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into CF.
+The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S GRID.
+This study identifies the most appropriate wind turbine power curve to use from the 121 power curves, each at five different levels of smoothing, made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within Atlite.
 
@@
 
-The output of those three datasets are one CF time series for both wind and Solar PV.
-In addition to analysing wind and solar PV generation separately, a combined CF was computed for each dataset by averaging wind and solar PV generation, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
+The datasets provide CF time series for both wind and solar PV.
+In addition to analysing the CF of wind and solar PV separately, a combined CF was computed for each dataset by averaging wind and solar PV CF, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
 This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% PV capacity.
@@
 In this study, a RES drought event was defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-The choice of this threshold is somewhat arbitrary, but aligns with similar studies on low renewable energy production \citep{kaspar2019drought, ohba2022drought, mayer2023drought}.
 By using a 24-hour moving average, fewer but longer-lasting events were captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
+The 24-hour rolling average also avoids potential masking of day-long events due to their start time.
 A fixed threshold approach was chosen in this study to enable consistent inter-comparison between datasets.
@@
 
-The verification of wind generation has put forward two main facts.
-On the one hand, the ATL dataset is skilled at reproducing onshore wind CF and droughts over Ireland.
+This verification for wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
+The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and droughts over Ireland.
 On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
-This puts forward one of the problems that come with using generalised datasets for the characterisation of RES droughts: biases severely affect the ability of any given model to reproducing them.
-The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in the ability to reproduce the changes in CF (visible in the very similar CC).
-However, their ability to reproduce the exact value is much lower than that of ATL (RMSE almost 4 times bigger for the two C3S with respect to ATL), and particularly biased towards an overestimation of CF (clear in the MBE values), which leads to the underestimation of droughts.
-This puts forward the need to use fully verified models to assess RES droughts, whose impact on the analysis of RES droughts will be explored in section~\ref{sec:analysis}.
+This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
+The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in their ability to reproduce the changes in CF, as seen by their similar CC scores.
+However, their ability to reproduce the actual CF values is much lower than that of ATL, with RMSE scores almost four times bigger for the two C3S datasets.
+There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of droughts.
+This highlights the need to use regionally verified models to assess RES droughts.
 
@@
 Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
-The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S performs best overall, with the lowest RMSE and a high correlation coefficient, suggesting a closer match to observed data.
-All models show a slight positive bias, with ATL exhibiting a slightly lower correlation and higher RMSE.
+The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S GRD and C3S NAT perform better than ATL for solar PV CF, with lower RMSE and MBE, and higher CC scores.
+This may be due to the statistical approach taken by C3SE for the orientation of the PV panels.
 
@@
 This results in uncertainties in solar PV generation data and the actual generating capacity in the first few months after each farm is connected.
+Overall, C3S GRD performs slightly better than the other datasets in reproducing observed solar PV drought events.
 
@@
 
-As one of the goals of this analysis is to assess the combination of wind and solar PV generation, the complementary nature of these energy sources will mitigate the limitations in solar PV-only results.
-Still, a few differences between wind and solar PV stand out.
-First, no single model has proven best for the characterisation of solar PV, although the statistical approach taken by the two C3S datasets has proven superior to the single panel model in ATL (Tab.~\ref{tab:pv_skill_scores}).
-The short time considered in the verification due to the limited data availability for solar PV in Ireland is part of the cause for our limited ability to select a specific overall best solar PV dataset.
-However, our results seem to indicate that it is possible that the generic datasets perform better for solar PV than for wind, as the variability between the different panel models is smaller in comparison to the existing differences between wind turbine power curves.
-Therefore, all through section~\ref{sec:analysis} the differences between the three models will be highlighted, considering that neither stands out positively.
-
 \subsection{Analysis}
@@
 
-In this section, RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of solar PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and solar PV capacity increases to 8.6 GW.
+In this section, RES droughts are analysed by calculating the frequency and duration of RES drought events, the return periods for different RES drought durations, and the seasonality of RES drought events.
+Understanding the characteristics and timing of RES drought events enables system operators to optimally plan for reserve capacity requirements, ensuring grid stability and security of supply.
+Results are presented for the three datasets, allowing their differences on the characterisation of RES droughts to be clearly identified.
+
+RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of solar PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and solar PV capacity increases to 8.6 GW.
 Both scenarios were driven by 45 years of ERA5 data.
@@
 
-It is important to highlight that this analysis considers two key aspects: the absolute values that characterise RES droughts, which are crucial for power system planning, and the relative differences observed when comparing the various datasets and energy scenarios described in Section~\ref{sec:scenarios}.
-This complete analysis puts forward the differences between the datasets, showing the impact of possible misrepresentations of the system on the analysis of RES drought events.
-
 \subsubsection{Annual Number of RES Droughts}
 
-The analysis of annual RES drought events reveals trends that are largely consistent with earlier studies.
+The first part of the analysis examines the annual number of RES drought events.
 When only wind energy is considered (Fig.\ref{fig:boxplot_number_events}a), the number of drought events decreases as the duration range increases, with very few events lasting more than seven days.
-This pattern aligns with previous research showing that wind droughts tend to be short and frequent.
 In contrast, for solar PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
@@
 Moreover, the comparison between wind and solar PV results indicates that the median, first, and third quartiles for solar PV are consistently higher than or equal to those for wind.
-This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle, as observed in other studies.
-When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, reaffirming wind’s dominance in the current energy mix.
+This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle.
+When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
 However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
-These findings are in line with earlier studies that highlight how increasing solar PV capacity can mitigate drought frequency through the anti-correlated seasonal patterns of wind and solar generation.
 
-Additionally, the consistently higher drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the impact of model selection, particularly the influence of wind turbine power curve representation, on quantifying RES droughts.
-This observation is consistent with previous research, which has also noted that assumptions regarding turbine characteristics can significantly affect drought duration estimates.
-Despite the differences in the number of RES droughts detected by each dataset, the overall effect of balancing the shares of wind and solar PV is consistent, independently of the complexity of the dataset used.
+The consistently higher drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the importance of wind turbine power curve representation when quantifying RES droughts.
+Whereas the three datasets agree on the overall effect of balancing the share of wind and solar PV generation, they differ at a quantitative level, which has crucial implications for energy planning.
 
@@
 The RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
-A return period is the estimated average time interval between events of a specified duration or intensity (not to be confused with the frequency of their occurrence within a fixed time frame).
-Fig.~\ref{fig:return_periods} illustrates the return periods for varying RES drought durations, highlighting how often different drought lengths are likely to occur across the datasets.
-This analysis not only quantifies the likelihood of prolonged low-generation periods but also provides insight into how extreme events are distributed across different timescales, helping to assess the variability of rare but impactful events.
-Understanding these return periods is crucial, as even infrequent droughts can challenge energy security by placing significant strain on conventional backup sources necessary to maintain supply in high-RES scenarios.
+A return period is the estimated average time interval between events of a specified duration (not to be confused with the frequency of their occurrence within a fixed time frame).
+Fig.~\ref{fig:return_periods} shows the return periods for different RES drought durations, which can be used to capture the most extreme events affecting the system.
+Understanding their return periods is crucial, as extreme yet rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
 
@@
 
-Contrary to the number of events per year and per duration, the return periods show large differences in the behaviours of the three datasets.
-The reproduction of average behaviours is still manageable using the C3S datasets, but the magnitude of extremes is not well reproduced when the share of wind is large.
-System planning based on the wrong datasets could yield an underestimation of the potential for extreme RES droughts, eventually leading to shortages linked to undersized reserve capacity.
-Even though this effect will be mitigated, affecting only the most extreme events with the introduction of widespread solar PV, it is important to keep in mind this effect for all wind-dominated electricity systems.
+The return periods calculated from the three datasets show large differences, in particular for the more extreme events with longer return periods.
+The C3S datasets produce shorter RES drought durations for these events, which would have the largest impact on the power system.
+This shows that system planning based on the wrong datasets could yield an underestimation of the duration of extreme RES droughts, potentially leading to shortages linked to undersized reserve capacity.
 
@@
 
-The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as drought events.
+The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as part of a RES drought event.
 Wind droughts tend to be more frequent during summer, whereas solar PV droughts are more common in winter due to reduced sunlight.
-By comparing these seasonal patterns across different datasets and energy scenarios, the study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of drought events.
+By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of RES drought events.
 

### 2025-03-28 Article/RES_droughts_Morin_ReEn_R2.tex
Modify seasonality and conclusions

@@
 
-The seasonal variations observed in this study have important implications for energy planning.
-Given that energy demand peaks in winter for Northern European countries, understanding these seasonal patterns is critical for assessing the need for conventional backup or storage solutions during periods of prolonged low renewable output.
-The findings underscore that even small differences in model assumptions leads to significant variations in drought estimates, thereby affecting the reliability of the energy system during critical periods.
-Additionally, the integration of large shares of solar PV in the system leads to a general reduction of RES droughts year-round.
-However, it can not counter the presence of RES droughts in the winter months due to its natural limitations, inevitably leading to higher reserve capacity needs during winter months as reliance on RES increases.
-Such insights are essential for policymakers to develop targeted strategies that enhance grid resilience and ensure a stable energy supply throughout the year.
+The seasonal variations of RES droughts observed in this study have important implications for energy planning.
+Energy demand peaks in winter for Northern European countries, making the seasonality of RES droughts critical for the sizing of reserve capacity.
+Our results show that selecting the wrong dataset could severely underestimate RES droughts during winter months, thereby affecting the reliability of the energy system during critical periods.
+Additionally, the integration of large shares of solar PV in the system leads to a generalised reduction of RES droughts, yet winter months present a slight increase.
+The natural limitations of solar PV lead to inevitably higher reserve capacity needs during winter months as reliance on RES increases.
+These types of insights are essential to develop targeted strategies that enhance grid resilience and ensure a stable energy supply throughout the year.
 
@@
 
-This study has explored the characterisation of RES droughts in the real-life transition from a wind-dominated system to a more balanced system with integrated solar PV.
-This has been done through the comparison over a 45-year period of three different datasets: one based on a hand-crafted validated model and two based on a generic model, C3S-Energy.
-The generic model has two versions, one using large-scale aggregated information, and one which includes the locations of farms as well (also considered in the validated model).
+This study has explored the characterisation of RES droughts in the transition from a wind-dominated system to a more balanced system with integrated solar PV, based on the real case of Ireland.
+Three different datasets were compared over a 45-year period: one created using a regionally validated model and two derived from a generic dataset for Europe, C3S-Energy.
+The two datasets derived from C3S-Energy present different approaches, with one using large-scale aggregated information only, and the other one including the locations of farms as well.
+The regionally validated model considered the locations of farms as well as tailored wind and PV models selected to best represent the actual generation in Ireland.
 
-Our results show that generic datasets present limitations in the quantification of RES droughts.
-Although the three datasets used capture overall trends in drought occurrence, significant differences emerge when using non-tailored data for the study of RES droughts.
-This finding highlights that the choice of dataset and its underlying assumptions can lead to consequential differences in estimated RES drought characteristics, emphasising the need for datasets that are specifically designed for extreme event analysis.
+Our results show the limitations in the quantification of RES droughts present in datasets that have not undergone regional validation.
+The three datasets used in this study are able to capture overall trends in drought occurrence such as the seasonal cycle or the effect of increasing the share of solar PV.
+However, significant differences in the quantitative values, particularly the extremes, emerge when using non-validated datasets for the study of RES droughts.
+This finding highlights that using a non-validated dataset can lead to undersized reserve capacity, with the associated negative consequences for grid stability and security of supply.
 
-This study reveals that differences in model parametrisation, particularly in the representation of wind turbine power curves and solar PV panel characteristics, have a stronger influence on the drought estimates than the inclusion of RES farm locations.
-The use of a validated dataset with a carefully selected wind turbine power curve consistently produced higher return periods and a greater number of drought events for wind energy than the datasets derived from C3S-Energy.
-This suggests that fine-tuning model parameters to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
+This study has also revealed that differences in the wind turbine power curves and solar PV panel models have a stronger influence on the estimation of RES droughts than the consideration of RES farm locations.
+The two datasets derived from C3S-Energy consistently underestimated the number of wind drought events and the frequency of extremes when compared to the regionally validated dataset with a specifically selected wind turbine power curve.
+This suggests that a meticulous selection of the wind turbine power curve to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
 
-Finally, the effect on RES droughts of the integration of solar PV in a wind-dominated system has been explored in a real-case setting.
-In the presented example of Ireland, the analysis has demonstrated that transitioning to a more balanced system with similar amounts of wind and solar PV markedly reduces the frequency, duration, and seasonal variability of drought events.
+Finally, the effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored in a real-case setting based on Ireland.
+Our analysis has demonstrated that transitioning to a system with similar amounts of wind and solar PV reduces the frequency, duration, and seasonal variability of drought events.
 This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV typically peaks in summer while wind generation is more consistent in winter.
-Thus, a more diversified renewable energy mix not only mitigates extreme drought conditions but also enhances overall system resilience, providing valuable insights for policymakers tasked with ensuring energy security.
+However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries like Ireland.
+Still, a more diversified renewable energy mix mitigates extreme drought conditions and enhances overall system resilience.
 
-The results presented in this study have several limitations.
-Although ERA5 is among the best reanalysis datasets for renewable energy analysis, it still presents some biases, and its resolution may not capture local-scale phenomena.
-This is especially limiting if individual farms are considered, but its still shows good skill at statistically representing nation-wide behaviours considering the distribution of farms.
-Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three datasets considering only weather-derived drivers, but does not account for demand variations.
-Consequently, while this approach enables a consistent inter-comparison, it will not signal events that are most critical for power system operations for their mismatch of RES generation and demand.
+The results presented in this study have three main limitations.
+First, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still presents some biases and limited spatial resolution.
+Second, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though we would expect the main takeaways to be the same.
+Third, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.
 

### 2025-03-28 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 
-\title{Reducing RES Droughts through the integration of wind and Solar PV}
+\title{Reducing RES Droughts through the integration of wind and solar PV}
 
@@
 Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
-The addition of solar PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal drought patterns.
+The addition of solar PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal RES drought patterns.
 These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
@@
 Alternative methods adjust the CF threshold dynamically over the year to account for seasonal variations in renewable production.
-Raynaud et al.~\citep{raynaud2018drought} defined droughts as sequences of days with energy production below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}.
+Raynaud et al.~\citep{raynaud2018drought} defined RES droughts as sequences of days with renewable electricity generation below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}.
 Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models.
-Other studies have defined droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S.
-Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe drought events.
-Another method defines energy drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
+Other studies have defined RES droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S.
+Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe RES drought events.
+Another method defines RES drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
 This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
 Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.
-~\citep{lei2024drought} used it to quantify droughts in wind-PV-hydro systems in China.
+~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.
 
 In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual load) is high.
-Raynaud et al.~\citep{raynaud2018drought} defined both energy production and energy supply droughts, and showed the difference in their patterns in a hypothetical fully renewable system composed of wind, PV and run-of-the-river hydropower.
-Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the energy production index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
+Raynaud et al.~\citep{raynaud2018drought} showed the difference between RES droughts and high residual load events in a hypothetical fully renewable system composed of wind, solar PV and run-of-the-river hydropower.
+Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the electricity generation index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
 This index was also applied to the U.S.
-by Bracken et al~\citep{bracken2024drought}, revealing a consistent increase in the drought magnitude when load is considered, despite showing differing results across regions.
+by Bracken et al~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
 
@@
 Two of the datasets are part of C3S Energy (C3SE), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
-One of the C3S-E datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
+One of the C3SE datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
 The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
@@
 This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV power.
-The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3S datasets.
+The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3SE dataset.
 
@@
 The EU Copernicus Climate Change Service developed the C3S-Energy (C3SE) renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
-This dataset provides hourly CF for wind and solar PV energy from 1979 to the present.
+This dataset provides hourly CF for wind and solar PV power from 1979 to the present.
 The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°.
@@
 
-This study uses onshore wind and solar PV CF time series from three datasets to analyse RES droughts across the island of Ireland.
-Data downloaded from C3SE were used to obtain two datasets: one based on national-level data (C3S NAT), and another on grid-level data (C3S GRD).
-The third dataset was computed using the Atlite model (ATL).
+This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S NAT) and grid-level data (C3S GRD), and one derived from the Atlite model (ATL).
+Fig.~\ref{fig:violin_plots} presents the statistics of the three datasets in violin plots for wind and solar PV.
+These plots illustrate the density of CF values over time, highlighting differences between the datasets and their alignment with observed data.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
+	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
+Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S GRD (blue), and C3S NAT (purple).
+The mean value for each dataset is marked with a green horizontal line.
+The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
+	\label{fig:violin_plots}
+\end{figure}
 
@@
 
-\subsection{C3S Energy Gridded: C3S GRID}
+\subsection{C3S Energy Gridded: C3S GRD}
 \label{sec:c3se_g}
 
-The C3S GRID dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
+The C3S GRD dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
 For each farm, the CF from the nearest grid point on the C3SE dataset was selected.
@@
 The ATL dataset is produced using the Atlite model.
-Atlite allows the user to define the wind turbine power curve and PV panel model to use when converting weather variables to wind and solar PV gerenation.
+Atlite allows the user to define the wind turbine power curve and PV panel model to use when converting weather variables to wind and solar PV generation.
 The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
 The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into CF.
-The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S GRID.
+The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S GRD.
 This study identifies the most appropriate wind turbine power curve to use from the 121 power curves, each at five different levels of smoothing, made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within Atlite.
@@
 
-The datasets provide CF time series for both wind and solar PV.
+The three datasets provide CF time series for both wind and solar PV.
 In addition to analysing the CF of wind and solar PV separately, a combined CF was computed for each dataset by averaging wind and solar PV CF, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
-This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% PV capacity.
+This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% solar PV capacity.
 Given that solar PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and solar PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed solar PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
@@
 
-For each dataset (ATL, C3S GRID, and C3S NAT), four distinct scenarios are examined, as summarised below:
+For each dataset (ATL, C3S GRD, and C3S NAT), four distinct scenarios are examined, as summarised below:
 
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S GRID and c) C3S NAT datasets}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S GRD and c) C3S NAT datasets}
 	\label{fig:wind_verification_contour}
@@
 The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
-The two C3S datasets are shown to overestimate the observed CF, whereas the Atlite model is in good agreement with the observed data.
+The two C3S datasets are shown to overestimate the observed CF, whereas the ATL dataset is in good agreement with the observed data.
 The skill scores presented in Table~\ref{tab:wind_skill_scores} show that ATL performs better than the two C3S datasets for all of the skill scores.
@@
 		\cline{2-4}
-		& \textbf{ATL} & \textbf{C3S GRID} & \textbf{C3S NAT} \\ \hline
+		& \textbf{ATL} & \textbf{C3S GRD} & \textbf{C3S NAT} \\ \hline
 		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.981           & 0.972            & 0.970            \\ \hline
@@
 This verification for wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
-The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and droughts over Ireland.
-On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
+The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts over Ireland.
+On the other hand, the power curve used for both C3S GRD and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
 This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
@@
 However, their ability to reproduce the actual CF values is much lower than that of ATL, with RMSE scores almost four times bigger for the two C3S datasets.
-There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of droughts.
+There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of RES droughts.
 This highlights the need to use regionally verified models to assess RES droughts.
@@
 
-Figure \ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
+Fig.~\ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
 The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S GRD and C3S NAT perform better than ATL for solar PV CF, with lower RMSE and MBE, and higher CC scores.
@@
 	\end{tabular}
-	\caption{Skill scores for Solar PV CF for the three datasets compared to observed data}
+	\caption{Skill scores for solar PV CF for the three datasets compared to observed data}
 	\label{tab:pv_skill_scores}
@@
 The first part of the analysis examines the annual number of RES drought events.
-When only wind energy is considered (Fig.\ref{fig:boxplot_number_events}a), the number of drought events decreases as the duration range increases, with very few events lasting more than seven days.
-In contrast, for solar PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
+When only wind energy is considered (Fig.~\ref{fig:boxplot_number_events}a), the number of RES drought events decreases as the duration range increases, with very few events lasting more than seven days.
+In contrast, for solar PV energy (Fig.~\ref{fig:boxplot_number_events}b), RES drought frequency declines from one to eight days and then slightly increases for longer durations.
 This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low solar PV output.
@@
 This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle.
-When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
-However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
+When wind and solar PV are combined under the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
+However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
 
-The consistently higher drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the importance of wind turbine power curve representation when quantifying RES droughts.
+The consistently higher RES drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the importance of wind turbine power curve representation when quantifying RES droughts.
 Whereas the three datasets agree on the overall effect of balancing the share of wind and solar PV generation, they differ at a quantitative level, which has crucial implications for energy planning.
@@
 
-For wind (Fig.~\ref{fig:return_periods}a), the log‐linear increase in return periods observed in this study confirms that longer droughts occur exponentially less frequently, a trend consistent with earlier research on wind variability.
-In the case of solar PV droughts (Fig.\ref{fig:return_periods}b), the ATL dataset shows a general log‐linear trend, whereas the C3S datasets exhibit a sudden increase in drought duration for events exceeding sixteen days.
-This abrupt rise reflects differences in how solar PV output is handled near the CF threshold during low irradiance conditions.
+The duration of wind droughts (Fig.~\ref{fig:return_periods}a) increases in a log-linear fashion across the three datasets.
+The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
+In the case of solar PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S datasets.
+The ATL dataset show a generally log-linear increase.
+For C3S GRD and C3S NAT, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
+Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
+This sudden increase again reflects the impact of extended periods of low PV generation during winter in Ireland.
+The difference between the ATL and the C3S results arises from differences in the datasets near the threshold of 0.1 CF.
+ATL remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented RES drought events.
+In contrast, C3S GRD and C3S NAT tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
 
@@
 In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
+For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
 Despite the lower wind share in the 57W-43PV scenario, typically known for its relative stability, the balanced share with solar PV leads to extended return periods for RES droughts.
-This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of drought events in a diversified energy portfolio.
+This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of RES drought events in a diversified energy portfolio.
 
@@
 For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the ATL dataset, compared to only five days in the C3S datasets.
-This difference underscores the importance of model selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence drought duration estimates.
+This difference underscores the importance of model selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES droughts duration estimates.
 Additionally, in all four graphs, the similarity between results from the two C3S datasets suggests that assumptions in the ATL dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
@@
 
-For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June, July, August) identified as droughts compared to only 4\% in winter (December, January, February).
+For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June, July, August) identified as RES droughts compared to only 4\% in winter (December, January, February).
 This strong seasonal signal is less evident in the C3S datasets, which suggests that the differences in the underlying wind power curves play a significant role.
-In ATL, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of drought hours during the summer months.
+In ATL, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of RES drought hours during the summer months.
 In contrast, solar PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
 Across all datasets, over 60\% of winter hours are classified as solar PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
-Moreover, ATL tends to record a slightly higher percentage of drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
+Moreover, ATL tends to record a slightly higher percentage of RES drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
 These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
@@
 This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
-Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours.
+Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall RES drought hours.
 In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S GRD, and from 9\% to 4\% for C3S-E NAT.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S GRD, and from 9\% to 4\% for C3S NAT.
 The balanced mix of wind and solar PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
@@
 The two datasets derived from C3S-Energy present different approaches, with one using large-scale aggregated information only, and the other one including the locations of farms as well.
-The regionally validated model considered the locations of farms as well as tailored wind and PV models selected to best represent the actual generation in Ireland.
+The regionally validated model considered the locations of farms as well as tailored wind and solar PV models selected to best represent the actual generation in Ireland.
 
 Our results show the limitations in the quantification of RES droughts present in datasets that have not undergone regional validation.
-The three datasets used in this study are able to capture overall trends in drought occurrence such as the seasonal cycle or the effect of increasing the share of solar PV.
+The three datasets used in this study are able to capture overall trends in RES drought occurrence such as the seasonal cycle or the effect of increasing the share of solar PV.
 However, significant differences in the quantitative values, particularly the extremes, emerge when using non-validated datasets for the study of RES droughts.
@@
 Finally, the effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored in a real-case setting based on Ireland.
-Our analysis has demonstrated that transitioning to a system with similar amounts of wind and solar PV reduces the frequency, duration, and seasonal variability of drought events.
+Our analysis has demonstrated that transitioning to a system with similar amounts of wind and solar PV reduces the frequency, duration, and seasonal variability of RES drought events.
 This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV typically peaks in summer while wind generation is more consistent in winter.
 However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries like Ireland.
-Still, a more diversified renewable energy mix mitigates extreme drought conditions and enhances overall system resilience.
+Still, a more diversified renewable energy mix mitigates extreme RES drought conditions and enhances overall system resilience.
 
-The results presented in this study have three main limitations.
-First, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still presents some biases and limited spatial resolution.
-Second, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though we would expect the main takeaways to be the same.
-Third, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.
+The results presented in this study have four main limitations.
+First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though we would expect the main takeaways to be the same.
+Second, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.
+Third, the availability of solar PV data is limited to a relatively short time period, as recent expansions in installed capacity have significantly changed the generation landscape.
+Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution.
 

### 2025-03-31 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 
-The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \citep{eurostat2023share}.
+The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022~\citep{eurostat2023share}.
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and solar photovoltaic (PV) power.
@@
 Similarly, Mockert et al.~\citep{mockert2023drought} examined the link between weather regimes and RES droughts in Germany using a 48-hour rolling window under a threshold to define RES droughts.
-Similar fixed-threshold approaches have also been applied using CF series reconstructed through machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary \citep{mayer2023drought}.
+Similar fixed-threshold approaches have also been applied using CF series reconstructed through machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary~\citep{mayer2023drought}.
 
@@
 This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
-Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.
-~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.
+Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.
 
@@
 This index was also applied to the U.S.
-by Bracken et al~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
+by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
 
-In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand, whose role will be addressed in the discussion.
+In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand.
 A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
 The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion.
-This provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts in the context of realistic scenarios.
+This provides valuable insights into the potential benefits of diversifying the share of wind and solar PV on RES droughts.
 
@@
 One of the C3SE datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts~\citep{mockert2023drought}.
+The third dataset produced by the authors was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts in Germany~\citep{mockert2023drought}.
 
 Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
-Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events \citep{kies2021drought}.
-In this work, we quantify the skill of a dataset developed for the European region (C3SE), when used for a specific country (Ireland).
-In particular, we investigate the impact of a generic versus a tailored dataset on the analysis of RES droughts, in the context of a transition from a wind-dominated system to one with a larger share of solar PV.
+Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events~\citep{kies2021drought}.
+This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country (Ireland).
+In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV.
 
@@
 Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MBE, and the percentage of overlap).
-Based on the best-performing metrics, the Beyer PV panel model was selected \citep{beyer2004pv}, using the Kaneka Hybrid panel option.
+Based on the best-performing metrics, the Beyer PV panel model was selected~\citep{beyer2004pv}, using the Kaneka Hybrid panel option.
 For all solar PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
@@
 
-The RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
+RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
 A return period is the estimated average time interval between events of a specified duration (not to be confused with the frequency of their occurrence within a fixed time frame).
@@
 Across all datasets, over 60\% of winter hours are classified as solar PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
-Moreover, ATL tends to record a slightly higher percentage of RES drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
+
+ATL tends to record a slightly higher percentage of RES drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
 These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
@@
 The results presented in this study have four main limitations.
-First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though we would expect the main takeaways to be the same.
+First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though the main takeaways would be expected to be the same.
 Second, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.

### 2025-03-31 Article/RES_droughts_Morin_ReEn_R2.tex
Add some more changes

@@
 The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period (Fig.~\ref{fig:find_res_droughts}).
-This means that a single hour above the threshold does not "break" a drought event if it is surrounded by prolonged low-generation hours.
-As a result, fewer but longer-lasting drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
+This means that a single hour above the threshold does not "break" a RES drought event if it is surrounded by prolonged low-generation hours.
+As a result, fewer but longer-lasting RES drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
 
@@
 The duration of wind droughts (Fig.~\ref{fig:return_periods}a) increases in a log-linear fashion across the three datasets.
-The log-linear trend indicates a predictable relationship between drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
+The log-linear trend indicates a predictable relationship between wind drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
 In the case of solar PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S datasets.
@@
 For C3S GRD and C3S NAT, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
-Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period.
+Beyond this duration, there is a sharp rise in solar PV drought duration for events up to a one-year return period.
 This sudden increase again reflects the impact of extended periods of low PV generation during winter in Ireland.
@@
 ATL remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented RES drought events.
-In contrast, C3S GRD and C3S NAT tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter.
+In contrast, C3S GRD and C3S NAT tend to fall below the threshold in similar conditions, resulting in longer continuous RES drought periods, especially during winter.
 
 Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
-In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged drought events.
+In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged RES drought events.
 For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
@@
 First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change might affect RES droughts.
-Second, expanding the geographic domain of the study to include the rest of Europe would provide a more comprehensive understanding of RES droughts in an interconnected energy grid.
+Second, expanding the geographic domain of the study to include the rest of Europe would provide a more comprehensive understanding of RES droughts in an interconnected electricity grid.
 This would require extensive verification across other European countries, making it a more complex but highly relevant challenge.
@@
 The ERA5 data can be obtained from the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
-The C3S datasets are also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
+The C3SE dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
 Information on wind and solar PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).

### 2025-04-01 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

@@
 Another method defines RES drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
-This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
-Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.
+This index was also applied to the U.S.
+by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
+The consideration of demand introduces a completely different factor to generation: weather is not the only driver, as the societal component plays a large role.
 
@@
 A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
-The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion.
-This provides valuable insights into the potential benefits of diversifying the share of wind and solar PV on RES droughts.
+The case study used in this paper is Ireland, a region where most RES generation comes from wind power and with ambitious targets for solar PV power expansion.
+This provides valuable insights into the potential benefits of adding solar PV installations in wind-dominated countries.
 

### 2025-04-01 Article/RES_droughts_Morin_ReEn_R3.tex
Create RES_droughts_Morin_ReEn_R3.tex

--- 2025-03-31 Article/RES_droughts_Morin_ReEn_R2.tex
@@
 
-In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual load) is high.
-Raynaud et al.~\citep{raynaud2018drought} showed the difference between RES droughts and high residual load events in a hypothetical fully renewable system composed of wind, solar PV and run-of-the-river hydropower.
-Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the electricity generation index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
+In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual demand) is high.
+Raynaud et al.~\citep{raynaud2018drought} showed the difference between RES droughts and high residual demand events in a hypothetical fully renewable system composed of wind, solar PV and run-of-the-river hydropower.
+Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual demand, whose correlation to the electricity generation index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
 This index was also applied to the U.S.
-by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
+by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when demand is considered, despite showing differing results across regions.
 
-In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand.
+In this paper, the focus is exclusively on renewable electricity generation, to keep the focus on RES droughts driven by the weather.
 A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
-The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion.
-This provides valuable insights into the potential benefits of diversifying the share of wind and solar PV on RES droughts.
+The case study used in this paper is Ireland, a region where most RES generation comes from wind power and with ambitious targets for solar PV power expansion.
+This provides valuable insights into the potential benefits of adding solar PV installations in wind-dominated countries.
 
@@
 
-The aim of this study is to answer three questions which are relevant for systems with a large share of RES generation:
+The aim of this study is to answer two questions which are relevant for systems with a large share of RES generation:
 \begin{itemize}
-	\item Do generic datasets have sufficient skill to reliably quantify extreme RES drought events?
-	\item What is the importance of using accurate RES farm locations, and regionally-validated wind and solar PV models, when analysing of RES droughts?
+	\item Do generic datasets have sufficient skill to reliably quantify RES drought events?
 	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES drought events?
@@
 This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S NAT) and grid-level data (C3S GRD), and one derived from the Atlite model (ATL).
-Fig.~\ref{fig:violin_plots} presents the statistics of the three datasets in violin plots for wind and solar PV.
-These plots illustrate the density of CF values over time, highlighting differences between the datasets and their alignment with observed data.
-
-\begin{figure}[ht!]
-	\centering
-	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
-	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
-Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S GRD (blue), and C3S NAT (purple).
-The mean value for each dataset is marked with a green horizontal line.
-The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
-	\label{fig:violin_plots}
-\end{figure}
 
@@
 This step allowed us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
+The general statistics of the observed data and the three datasets are presented in the violin plots in Fig.~\ref{fig:violin_plots}, considering the 2014-2023 validation period for wind and 2023 for solar PV.
+These plots illustrate the density of CF values for each dataset, highlighting differences between the datasets and their alignment with observed data.
+The general CF values show ATL aligns better with OBS for wind, and all datasets are similar for PV.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
+	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
+Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S GRD (blue), and C3S NAT (purple).
+The mean value for each dataset is marked with a green horizontal line.
+The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
+	\label{fig:violin_plots}
+\end{figure}
 
@@
 
-This study has explored the characterisation of RES droughts in the transition from a wind-dominated system to a more balanced system with integrated solar PV, based on the real case of Ireland.
+The two questions presented and answered in this study are: XXX?; YYY?
 Three different datasets were compared over a 45-year period: one created using a regionally validated model and two derived from a generic dataset for Europe, C3S-Energy.
@@
 
+% Answer XXX - generic models are sufficient to quantify RES droughts
 Our results show the limitations in the quantification of RES droughts present in datasets that have not undergone regional validation.
@@
 
+% Answer YYY - role of intergrating PV
 Finally, the effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored in a real-case setting based on Ireland.
@@
 
+% Limitations
 The results presented in this study have four main limitations.
 First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though the main takeaways would be expected to be the same.
-Second, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.
+Second, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
 Third, the availability of solar PV data is limited to a relatively short time period, as recent expansions in installed capacity have significantly changed the generation landscape.
-Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution.
+Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
 
+% Future work
 Future work is planned to extend the current analysis.

### 2025-04-01 Article/RES_droughts_Morin_ReEn_R2.tex
Changed R2 version to be the same as last Thrursday Evening

@@
 In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual load) is high.
-Raynaud et al.~\citep{raynaud2018drought} showed the difference between RES droughts and high residual load events in a hypothetical fully renewable system composed of wind, solar PV and run-of-the-river hydropower.
-Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the electricity generation index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
+Raynaud et al.~\citep{raynaud2018drought} defined both energy production and energy supply droughts, and showed the difference in their patterns in a hypothetical fully renewable system composed of wind, PV and run-of-the-river hydropower.
+Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual load, whose correlation to the energy production index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
 This index was also applied to the U.S.
-by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions.
+by Bracken et al~\citep{bracken2024drought}, revealing a consistent increase in the drought magnitude when load is considered, despite showing differing results across regions.
 
-In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand.
+In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand, whose role will be addressed in the discussion.
 A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
-The case study used in this paper is Ireland, a region where most RES generation comes from wind power and with ambitious targets for solar PV power expansion.
-This provides valuable insights into the potential benefits of adding solar PV installations in wind-dominated countries.
+The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion.
+This provides valuable insights into the potential benefits of diversifying the renewable energy mix on RES droughts in the context of realistic scenarios.
 
@@
 One of the C3SE datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset produced by the authors was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts in Germany~\citep{mockert2023drought}.
+The third dataset was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
+Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts~\citep{mockert2023drought}.
 
 Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
-Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events~\citep{kies2021drought}.
-This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country (Ireland).
-In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV.
+Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events \citep{kies2021drought}.
+In this work, we quantify the skill of a dataset developed for the European region (C3SE), when used for a specific country (Ireland).
+In particular, we investigate the impact of a generic versus a tailored dataset on the analysis of RES droughts, in the context of a transition from a wind-dominated system to one with a larger share of solar PV.
 
@@
 The EU Copernicus Climate Change Service developed the C3S-Energy (C3SE) renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
-This dataset provides hourly CF for wind and solar PV power from 1979 to the present.
+This dataset provides hourly CF for wind and solar PV energy from 1979 to the present.
 The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°.
@@
 
-This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S NAT) and grid-level data (C3S GRD), and one derived from the Atlite model (ATL).
-Fig.~\ref{fig:violin_plots} presents the statistics of the three datasets in violin plots for wind and solar PV.
-These plots illustrate the density of CF values over time, highlighting differences between the datasets and their alignment with observed data.
-
-\begin{figure}[ht!]
-	\centering
-	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
-	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
-Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S GRD (blue), and C3S NAT (purple).
-The mean value for each dataset is marked with a green horizontal line.
-The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
-	\label{fig:violin_plots}
-\end{figure}
+This study uses onshore wind and solar PV CF time series from three datasets to analyse RES droughts across the island of Ireland.
+Data downloaded from C3SE were used to obtain two datasets: one based on national-level data (C3S~NAT), and another on grid-level data (C3S~GRD).
+The third dataset was computed using the Atlite model (ATL).
 
@@
 This verification for wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
-The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts over Ireland.
-On the other hand, the power curve used for both C3S GRD and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
+The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and droughts over Ireland.
+On the other hand, the power curve used for both C3S GRID and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
 This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
@@
 However, their ability to reproduce the actual CF values is much lower than that of ATL, with RMSE scores almost four times bigger for the two C3S datasets.
-There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of RES droughts.
+There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of droughts.
 This highlights the need to use regionally verified models to assess RES droughts.
@@
 The first part of the analysis examines the annual number of RES drought events.
-When only wind energy is considered (Fig.~\ref{fig:boxplot_number_events}a), the number of RES drought events decreases as the duration range increases, with very few events lasting more than seven days.
-In contrast, for solar PV energy (Fig.~\ref{fig:boxplot_number_events}b), RES drought frequency declines from one to eight days and then slightly increases for longer durations.
+When only wind energy is considered (Fig.\ref{fig:boxplot_number_events}a), the number of drought events decreases as the duration range increases, with very few events lasting more than seven days.
+In contrast, for solar PV energy (Fig.\ref{fig:boxplot_number_events}b), drought frequency declines from one to eight days and then slightly increases for longer durations.
 This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low solar PV output.
@@
 This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle.
-When wind and solar PV are combined under the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
-However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
+When wind and solar PV are combined under the 91W-9PV scenario (Fig.\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
+However, in the 57W-43PV scenario (Fig.\ref{fig:boxplot_number_events}d), a marked reduction in drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
 
@@
 
-This study has explored the characterisation of RES droughts in the transition from a wind-dominated system to a more balanced system with integrated solar PV, based on the real case of Ireland.
-Three different datasets were compared over a 45-year period: one created using a regionally validated model and two derived from a generic dataset for Europe, C3S-Energy.
-The two datasets derived from C3S-Energy present different approaches, with one using large-scale aggregated information only, and the other one including the locations of farms as well.
-The regionally validated model considered the locations of farms as well as tailored wind and solar PV models selected to best represent the actual generation in Ireland.
+This study has explored the characterisation of RES droughts in the real-life transition from a wind-dominated system to a more balanced system with integrated solar PV.
+This has been done through the comparison over a 45-year period of three different datasets: one based on a hand-crafted validated model and two based on a generic model, C3S-Energy.
+The generic model has two versions, one using large-scale aggregated information, and one which includes the locations of farms as well (also considered in the validated model).
 
-Our results show the limitations in the quantification of RES droughts present in datasets that have not undergone regional validation.
-The three datasets used in this study are able to capture overall trends in RES drought occurrence such as the seasonal cycle or the effect of increasing the share of solar PV.
-However, significant differences in the quantitative values, particularly the extremes, emerge when using non-validated datasets for the study of RES droughts.
-This finding highlights that using a non-validated dataset can lead to undersized reserve capacity, with the associated negative consequences for grid stability and security of supply.
+Our results show that generic datasets present limitations in the quantification of RES droughts.
+Although the three datasets used capture overall trends in drought occurrence, significant differences emerge when using non-tailored data for the study of RES droughts.
+This finding highlights that the choice of dataset and its underlying assumptions can lead to consequential differences in estimated RES drought characteristics, emphasising the need for datasets that are specifically designed for extreme event analysis.
 
-This study has also revealed that differences in the wind turbine power curves and solar PV panel models have a stronger influence on the estimation of RES droughts than the consideration of RES farm locations.
-The two datasets derived from C3S-Energy consistently underestimated the number of wind drought events and the frequency of extremes when compared to the regionally validated dataset with a specifically selected wind turbine power curve.
-This suggests that a meticulous selection of the wind turbine power curve to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
+This study reveals that differences in model parametrisation, particularly in the representation of wind turbine power curves and solar PV panel characteristics, have a stronger influence on the drought estimates than the inclusion of RES farm locations.
+The use of a validated dataset with a carefully selected wind turbine power curve consistently produced higher return periods and a greater number of drought events for wind energy than the datasets derived from C3S-Energy.
+This suggests that fine-tuning model parameters to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
 
-Finally, the effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored in a real-case setting based on Ireland.
-Our analysis has demonstrated that transitioning to a system with similar amounts of wind and solar PV reduces the frequency, duration, and seasonal variability of RES drought events.
+Finally, the effect on RES droughts of the integration of solar PV in a wind-dominated system has been explored in a real-case setting.
+In the presented example of Ireland, the analysis has demonstrated that transitioning to a more balanced system with similar amounts of wind and solar PV markedly reduces the frequency, duration, and seasonal variability of drought events.
 This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV typically peaks in summer while wind generation is more consistent in winter.
-However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries like Ireland.
-Still, a more diversified renewable energy mix mitigates extreme RES drought conditions and enhances overall system resilience.
+Thus, a more diversified renewable energy mix not only mitigates extreme drought conditions but also enhances overall system resilience, providing valuable insights for policymakers tasked with ensuring energy security.
 
-The results presented in this study have four main limitations.
-First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though the main takeaways would be expected to be the same.
-Second, the definition of RES droughts based on generation does not consider the mismatch between renewable generation and demand, which may be of interest to system operators, as these events put large amounts of strain on the system.
-Third, the availability of solar PV data is limited to a relatively short time period, as recent expansions in installed capacity have significantly changed the generation landscape.
-Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution.
+The results presented in this study have several limitations.
+Although ERA5 is among the best reanalysis datasets for renewable energy analysis, it still presents some biases, and its resolution may not capture local-scale phenomena.
+This is especially limiting if individual farms are considered, but its still shows good skill at statistically representing nation-wide behaviours considering the distribution of farms.
+Moreover, the methodology employs a fixed threshold to define RES drought events, which is necessary for comparing the three datasets considering only weather-derived drivers, but does not account for demand variations.
+Consequently, while this approach enables a consistent inter-comparison, it will not signal events that are most critical for power system operations for their mismatch of RES generation and demand.
 

### 2025-04-01 Article/RES_droughts_Morin_ReEn_R3.tex
Added R3 which are modifications since Thursday

@@
 
-This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S NAT) and grid-level data (C3S GRD), and one derived from the Atlite model (ATL).
+This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S~NAT) and grid-level data (C3S~GRD), and one derived from the Atlite model (ATL).
 
-\subsection{C3S Energy National: C3S NAT}
+\subsection{C3S Energy National: C3S~NAT}
 \label{sec:c3se_n}
 
-The C3S NAT dataset is created by combining two inputs provided by C3SE at the corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
+The C3S~NAT dataset is created by combining two inputs provided by C3SE at the corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
 The two inputs are combined, using the actual installed capacity as weights.
@@
 
-\subsection{C3S Energy Gridded: C3S GRD}
+\subsection{C3S Energy Gridded: C3S~GRD}
 \label{sec:c3se_g}
 
-The C3S GRD dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
+The C3S~GRD dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
 For each farm, the CF from the nearest grid point on the C3SE dataset was selected.
@@
 The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into CF.
-The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S GRD.
+The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S~GRD.
 This study identifies the most appropriate wind turbine power curve to use from the 121 power curves, each at five different levels of smoothing, made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within Atlite.
@@
 
-New time series were generated for both the ATL and C3S GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
@@
 
-For each dataset (ATL, C3S GRD, and C3S NAT), four distinct scenarios are examined, as summarised below:
+For each dataset (ATL, C3S~GRD, and C3S~NAT), four distinct scenarios are examined, as summarised below:
 
@@
 	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
-Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S GRD (blue), and C3S NAT (purple).
+Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
 The mean value for each dataset is marked with a green horizontal line.
@@
 	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S GRD and c) C3S NAT datasets}
+	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S~GRD and c) C3S~NAT datasets}
 	\label{fig:wind_verification_contour}
@@
 		\cline{2-4}
-		& \textbf{ATL} & \textbf{C3S GRD} & \textbf{C3S NAT} \\ \hline
+		& \textbf{ATL} & \textbf{C3S~GRD} & \textbf{C3S~NAT} \\ \hline
 		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.981           & 0.972            & 0.970            \\ \hline
@@
 	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for ATL (red), C3S GRD (blue), C3S NAT (purple), and the observed data (black outline).
+	\caption{Average annual number of wind drought events for ATL (red), C3S~GRD (blue), C3S~NAT (purple), and the observed data (black outline).
 The wind droughts are identified from 2014 to 2023, considering the actual capacity of the system at any given time}
@@
 The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts over Ireland.
-On the other hand, the power curve used for both C3S GRD and C3S NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
+On the other hand, the power curve used for both C3S~GRD and C3S~NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
 This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
@@
 	\includegraphics[width=\textwidth]{verification_pv_contour.png}
-	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) ATL, b) C3S GRD and c) C3S NAT datasets}	
+	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) ATL, b) C3S~GRD and c) C3S~NAT datasets}	
 	\label{fig:solar_verification_contour}
@@
 Fig.~\ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
-The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S GRD and C3S NAT perform better than ATL for solar PV CF, with lower RMSE and MBE, and higher CC scores.
+The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S~GRD and C3S~NAT perform better than ATL for solar PV CF, with lower RMSE and MBE, and higher CC scores.
 This may be due to the statistical approach taken by C3SE for the orientation of the PV panels.
@@
 		\cline{2-4}
-		& \textbf{ATL} & \textbf{C3S GRD} & \textbf{C3S NAT} \\ \hline
+		& \textbf{ATL} & \textbf{C3S~GRD} & \textbf{C3S~NAT} \\ \hline
 		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.931            \\ \hline
@@
 This results in uncertainties in solar PV generation data and the actual generating capacity in the first few months after each farm is connected.
-Overall, C3S GRD performs slightly better than the other datasets in reproducing observed solar PV drought events.
+Overall, C3S~GRD performs slightly better than the other datasets in reproducing observed solar PV drought events.
 
@@
 	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
-	\caption{Number of solar PV drought events for ATL (red), C3S GRD (blue), and C3S NAT (purple) and the observed data (black outline).
+	\caption{Number of solar PV drought events for ATL (red), C3S~GRD (blue), and C3S~NAT (purple) and the observed data (black outline).
 The solar PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
@@
 When wind and solar PV are combined under the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
-However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S GRD, and 50\% for C3S NAT, demonstrating the beneficial effects of a more balanced energy mix.
+However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S~GRD, and 50\% for C3S~NAT, demonstrating the beneficial effects of a more balanced energy mix.
 
@@
 	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red), C3S GRD (blue), and C3S NAT (purple).
+	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
 The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
@@
 The ATL dataset show a generally log-linear increase.
-For C3S GRD and C3S NAT, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
+For C3S~GRD and C3S~NAT, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
 Beyond this duration, there is a sharp rise in solar PV drought duration for events up to a one-year return period.
@@
 ATL remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented RES drought events.
-In contrast, C3S GRD and C3S NAT tend to fall below the threshold in similar conditions, resulting in longer continuous RES drought periods, especially during winter.
+In contrast, C3S~GRD and C3S~NAT tend to fall below the threshold in similar conditions, resulting in longer continuous RES drought periods, especially during winter.
 
@@
 	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red triangle), C3S GRD (blue circle), and C3S NAT (purple square).
+	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red triangle), C3S~GRD (blue circle), and C3S~NAT (purple square).
 The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
@@
 In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S GRD, and from 9\% to 4\% for C3S NAT.
+Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S~GRD, and from 9\% to 4\% for C3S~NAT.
 The balanced mix of wind and solar PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
@@
 	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red dotted), C3S GRD (blue dashed), and C3S NAT (purple solid).
+	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red dotted), C3S~GRD (blue dashed), and C3S~NAT (purple solid).
 The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.

### 2025-04-02 Article/RES_droughts_Morin_ReEn_R3.tex
Update Reviewed Version of the Article

@@
 Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events~\citep{kies2021drought}.
-This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country (Ireland).
-In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV.
+This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country.
+In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV capacity.
 
@@
 	\item Do generic datasets have sufficient skill to reliably quantify RES drought events?
-	\item How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES drought events?
+	\item How does the integration of solar PV capacity into a predominantly wind-based system alter the characteristics of RES drought events?
 \end{itemize}
@@
 
-This study analyses RES droughts across the island of Ireland using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S~NAT) and grid-level data (C3S~GRD), and one derived from the Atlite model (ATL).
+This study analyses RES droughts using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S~NAT) and grid-level data (C3S~GRD), and one derived from the Atlite model (ATL).
 
@@
 
-It is important to note that the specific capacity values used in this study are illustrative and are not intended to reflect precise future realities.
+It is important to note that the specific capacity values used in this study are illustrative and are not intended to reflect accurate future realities.
 Instead, they serve to explore the impact of transitioning from a wind-dominated system (91W-9PV) to a more evenly distributed system (57W-43PV).
@@
 This means that a single hour above the threshold does not "break" a RES drought event if it is surrounded by prolonged low-generation hours.
-As a result, fewer but longer-lasting RES drought events are identified, which may better reflect real-world conditions where energy supply constraints persist over extended periods.
+As a result, fewer but longer-lasting RES drought events are identified, which may better reflect actual conditions where energy supply constraints persist over extended periods.
 
@@
 This step allowed us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
-The general statistics of the observed data and the three datasets are presented in the violin plots in Fig.~\ref{fig:violin_plots}, considering the 2014-2023 validation period for wind and 2023 for solar PV.
-These plots illustrate the density of CF values for each dataset, highlighting differences between the datasets and their alignment with observed data.
-The general CF values show ATL aligns better with OBS for wind, and all datasets are similar for PV.
+This validation step evaluates how well the datasets represent actual renewable energy production by comparing them against observed data.
+The overall statistical distribution of CF values for wind (2014–2023) and solar PV (2023) is presented in the violin plots in Fig.~\ref{fig:violin_plots}.
+These plots illustrate the density of CF values for each dataset, highlighting their differences and alignment with observations.
+The results indicate that ATL aligns more closely with OBS for wind, while all datasets exhibit similar distributions for solar PV.
 
@@
 	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
-	\caption{Violin plots of CF distributions for wind a) and solar PV b) datasets.
-Each violin represents the distribution of CF over time for different datasets: Observations (black), ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
-The mean value for each dataset is marked with a green horizontal line.
+	\caption{Violin plots of CF distributions for wind a) and solar PV b) for the Observations (grey) and the three datasets: ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
+The black dot shows the median values, while the black vertical lines represent the first and third quartiles.
 The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
@@
 	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by ATL (orange) and the Vestas V136/3450 used by C3SE (blue) b) Histograms of wind CF for Ireland from ATL (orange), C3SE (blue) and Observed (shaded)}
+	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by the ATL dataset (orange) and the Vestas V136/3450 used by the two C3S datasets (blue) b) Histograms of wind CF for Ireland for the ATL dataset (orange), the C3S datasets (blue) and Observed (shaded)}
 	\label{fig:power_curve}
@@
 
-The two questions presented and answered in this study are: XXX?; YYY?
-Three different datasets were compared over a 45-year period: one created using a regionally validated model and two derived from a generic dataset for Europe, C3S-Energy.
-The two datasets derived from C3S-Energy present different approaches, with one using large-scale aggregated information only, and the other one including the locations of farms as well.
-The regionally validated model considered the locations of farms as well as tailored wind and solar PV models selected to best represent the actual generation in Ireland.
+This study aimed to answer two key questions: Do generic datasets have sufficient skill to reliably quantify RES drought events?
+How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts?
+To address these questions, three datasets were compared: one developed using a regionally validated model and two derived from the European C3S-Energy dataset.
+The C3S-Energy datasets differ in their assumptions, one assumes a homogeneous distribution of wind and solar PV across the region, while the other includes the actual locations of RES farms.
+In contrast, the regionally validated model accounts for farm locations and uses tailored wind and solar PV models selected to represent the actual generation.
 
-% Answer XXX - generic models are sufficient to quantify RES droughts
-Our results show the limitations in the quantification of RES droughts present in datasets that have not undergone regional validation.
-The three datasets used in this study are able to capture overall trends in RES drought occurrence such as the seasonal cycle or the effect of increasing the share of solar PV.
-However, significant differences in the quantitative values, particularly the extremes, emerge when using non-validated datasets for the study of RES droughts.
-This finding highlights that using a non-validated dataset can lead to undersized reserve capacity, with the associated negative consequences for grid stability and security of supply.
+Our results demonstrate that datasets without regional validation have notable limitations in quantifying RES drought events.
+While all three datasets capture broad trends, significant differences arise in extreme event assessment.
+In particular, datasets without regional validation tend to underestimate the return periods of RES droughts, which can lead to insufficient reserve capacity planning and potential risks to grid stability and security of supply.
+This highlights the critical need for regionally validated datasets when studying RES droughts in renewable energy planning.
+The inclusion of wind and solar PV farm locations has a limited impact on RES drought analysis compared to the choice of wind turbine power curves and solar PV models.
 
-This study has also revealed that differences in the wind turbine power curves and solar PV panel models have a stronger influence on the estimation of RES droughts than the consideration of RES farm locations.
-The two datasets derived from C3S-Energy consistently underestimated the number of wind drought events and the frequency of extremes when compared to the regionally validated dataset with a specifically selected wind turbine power curve.
-This suggests that a meticulous selection of the wind turbine power curve to match observed data is crucial for accurately quantifying RES drought risks, thereby supporting more effective energy system planning.
-
-% Answer YYY - role of intergrating PV
-Finally, the effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored in a real-case setting based on Ireland.
-Our analysis has demonstrated that transitioning to a system with similar amounts of wind and solar PV reduces the frequency, duration, and seasonal variability of RES drought events.
-This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV typically peaks in summer while wind generation is more consistent in winter.
-However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries like Ireland.
+The effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored.
+Our analysis has demonstrated that transitioning to a system with more equal amounts of wind and solar PV capacity reduces the frequency, duration, and seasonal variability of RES drought events.
+This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV generation typically peaks in summer while wind generation is more consistent in winter.
+However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries.
 Still, a more diversified renewable energy mix mitigates extreme RES drought conditions and enhances overall system resilience.
 
-% Limitations
 The results presented in this study have four main limitations.
-First, the presented study uses a fixed threshold to define RES drought events, but other methods could yield different results, even though the main takeaways would be expected to be the same.
+First, the presented study uses a fixed threshold to define RES drought events, but other methods such as ones using the percentile of the generation could yield different results.
 Second, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
-Third, the availability of solar PV data is limited to a relatively short time period, as recent expansions in installed capacity have significantly changed the generation landscape.
+Third, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, whereas a longer validation period would be preferable.
 Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
 
-% Future work
 Future work is planned to extend the current analysis.

### 2025-04-02 Article/RES_droughts_Morin_ReEn_R3.tex
Update Reviewed Version of the Article

@@
 	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
-	\caption{Violin plots of CF distributions for wind a) and solar PV b) for the Observations (grey) and the three datasets: ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
+	\caption{Violin plots of CF distributions for a)  wind and b) solar PV for the Observations (grey) and the three datasets: ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
 The black dot shows the median values, while the black vertical lines represent the first and third quartiles.
-The red line indicates the threshold of 0.1 used in the study to identify RES droughts}
+The black dashed line indicates the threshold of 0.1 used in the study to identify RES droughts}
 	\label{fig:violin_plots}
@@
 How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts?
-To address these questions, three datasets were compared: one developed using a regionally validated model and two derived from the European C3S-Energy dataset.
-The C3S-Energy datasets differ in their assumptions, one assumes a homogeneous distribution of wind and solar PV across the region, while the other includes the actual locations of RES farms.
-In contrast, the regionally validated model accounts for farm locations and uses tailored wind and solar PV models selected to represent the actual generation.
+To address these questions, three datasets were compared: two derived from the European C3S-Energy dataset, and one developed by the authors.
+The datasets derived from C3S-Energy differ in their assumptions, one assumes a homogeneous distribution of wind and solar PV capacity across the region, while the other includes the actual locations of RES farms.
+The dataset developed by the authors uses a regionally validated model which accounts for farm locations and uses tailored wind and solar PV models selected to represent the actual generation.
 
-Our results demonstrate that datasets without regional validation have notable limitations in quantifying RES drought events.
-While all three datasets capture broad trends, significant differences arise in extreme event assessment.
-In particular, datasets without regional validation tend to underestimate the return periods of RES droughts, which can lead to insufficient reserve capacity planning and potential risks to grid stability and security of supply.
-This highlights the critical need for regionally validated datasets when studying RES droughts in renewable energy planning.
-The inclusion of wind and solar PV farm locations has a limited impact on RES drought analysis compared to the choice of wind turbine power curves and solar PV models.
+Our results demonstrate that datasets without regional validation misrepresent the frequency and duration of RES drought events due to their limited ability to reproduce the observations.
+The inclusion of wind and solar PV farm locations has limited impact on RES drought analysis compared to the choice of wind turbine power curves and solar PV models.
+Whereas all three datasets capture broad trends in the duration and seasonality of RES drought events, the actual number of events is consistently underestimated by the non-validated datasets.
+This effect becomes clearer for extreme events, as not using regionally validated datasets can yield an overestimation of the return periods of RES droughts.
+This can lead to insufficient reserve capacity planning and potential risks to grid stability and security of supply.
 
-The effect of the integration of solar PV in a wind-dominated system on RES droughts has been explored.
-Our analysis has demonstrated that transitioning to a system with more equal amounts of wind and solar PV capacity reduces the frequency, duration, and seasonal variability of RES drought events.
-This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV generation typically peaks in summer while wind generation is more consistent in winter.
+The effect of the integration of solar PV capacity in a wind-dominated system on RES droughts has been explored.
+Our analysis has demonstrated that transitioning to a system with more equal amounts of wind and solar PV capacity reduces the occurrence of RES drought events, mitigates extreme RES drought conditions and enhances overall system resilience.
+This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV generation typically peaks in summer while wind generation predominates during winter.
 However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries.
-Still, a more diversified renewable energy mix mitigates extreme RES drought conditions and enhances overall system resilience.
 
 The results presented in this study have four main limitations.
-First, the presented study uses a fixed threshold to define RES drought events, but other methods such as ones using the percentile of the generation could yield different results.
+First, the presented study uses a fixed threshold to define RES drought events, but other methods such as ones based on percentiles could be explored.
 Second, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
-Third, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, whereas a longer validation period would be preferable.
-Lastly, the source for weather data is ERA5, which is among the best reanalysis datasets for renewable energy applications, but still comes in a limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
+Third, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, although a longer validation period would be preferable.
+Last, the source for weather data is ERA5 has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
 

### 2025-04-03 Article/RES_droughts_Morin_ReEn_R3.tex
Update Reviewed Version of the Article

@@
 
-For each dataset (ATL, C3S~GRD, and C3S~NAT), four distinct scenarios are examined, as summarised below:
+In summary, for each of the three datasets (ATL, C3S GRD and C3S NAT) four energy scenarios are examined:
 
@@
 When wind and solar PV are combined under the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
-However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S~GRD, and 50\% for C3S~NAT, demonstrating the beneficial effects of a more balanced energy mix.
+However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S~GRD, and 50\% for C3S~NAT, demonstrating the beneficial effects of a more equal share of wind and solar PV capacity.
 
@@
 Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
-In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more diversified energy mix can substantially mitigate the frequency of prolonged RES drought events.
+In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more balanced share of wind and solar PV capacity can substantially mitigate the frequency of prolonged RES drought events.
 For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
-Despite the lower wind share in the 57W-43PV scenario, typically known for its relative stability, the balanced share with solar PV leads to extended return periods for RES droughts.
+Although wind is the more stable of the two sources, the lower wind share in the 57W-43PV scenario actually leads to extended return periods for RES droughts thanks to the increased share of PV and the complementarity between the two.
 This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of RES drought events in a diversified energy portfolio.

### 2025-04-07 Article/RES_droughts_Morin_ReEn_R3.tex
CS minor edits

@@
 Three different RES datasets are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
-The skill of the RES models is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
-The RES models are used to generate a 45-year hourly time series of RES capacity factor, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
-Results show the importance of using an accurate, validated RES model for RES drought risk assessment.
+The skill of the RES datasets is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
+The RES datasets are used to generate a 45-year hourly time series of RES capacity factor, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
+Results show the importance of using an accurate, validated RES dataset for RES drought risk assessment.
 The addition of solar PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal RES drought patterns.
@@
 Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
-Despite undergoing a validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events~\citep{kies2021drought}.
-This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country.
+Despite undergoing a general validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events subsequently identified~\citep{kies2021drought}.
+This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country; Ireland.
 In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV capacity.
@@
 \begin{itemize}
-	\item Do generic datasets have sufficient skill to reliably quantify RES drought events?
+	\item Do generic datasets have sufficient skill to reliably quantify RES drought events for specific countries?
 	\item How does the integration of solar PV capacity into a predominantly wind-based system alter the characteristics of RES drought events?
@@
 The spreadsheet available from the EirGrid website contains two key variables: generation and availability.
+Generation and availability values are available from 2014 onward for wind power and from 2018 onward for solar PV power, although solar PV availability data only became present in the Republic of Ireland in 2023.
 Generation is the energy that a RES farm actually contributed to the grid, which may include limitations introduced by the TSO to maintain grid stability, such as constraints and curtailment.
 Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-Generation and availability values are available from 2014 onward for wind power and from 2018 onward for solar PV power, although solar PV availability data only became present in the Republic of Ireland in 2023.
 This study focuses on availability for all analyses.
@@
 
-This study analyses RES droughts using onshore wind and solar PV CF time series from three datasets: two from C3SE, based on national-level data (C3S~NAT) and grid-level data (C3S~GRD), and one derived from the Atlite model (ATL).
+This study analyses RES droughts using onshore wind and solar PV CF time series from three datasets: two from C3SE; one based on national-level data (C3S~NAT) and the other on grid-level data (C3S~GRD), and a third dataset derived using the Atlite model (ATL).
 
@@
 
-New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap.
+New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap \textbf{[ADD REFERENCE]}.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
@@
 There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of RES droughts.
-This highlights the need to use regionally verified models to assess RES droughts.
+This highlights the need to use regionally verified datasets to assess RES droughts.
 
@@
 The figure reveals partial agreement between the three datasets and the observed data, with consistent results noticed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
-However, discrepancies appear in the other ranges, where the models diverge from the observed data.
+However, discrepancies appear in the other ranges, where the datasets diverge from the observed data.
 The main challenge in validating solar PV data stems from the recent installation of a large share of Ireland’s solar PV capacity, with over 65\% of the total solar PV capacity installed in 2023.
@@
 Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
-In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic increase in return periods across all durations, suggesting that a more balanced share of wind and solar PV capacity can substantially mitigate the frequency of prolonged RES drought events.
-For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) extends from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
-Although wind is the more stable of the two sources, the lower wind share in the 57W-43PV scenario actually leads to extended return periods for RES droughts thanks to the increased share of PV and the complementarity between the two.
+In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic reduction in RES drought durations, suggesting that a more balanced share of wind and solar PV capacity can substantially mitigate the frequency of prolonged RES drought events.
+For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) increases from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
 This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of RES drought events in a diversified energy portfolio.
@@
 
-Across Fig.~\ref{fig:return_periods}a, c, and, d, the return periods in the ATL dataset are consistently higher than those in the two C3S datasets.
+Across Fig.~\ref{fig:return_periods}a, c, and d, the return periods in the ATL dataset are consistently higher than those in the two C3S datasets.
 For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the ATL dataset, compared to only five days in the C3S datasets.
-This difference underscores the importance of model selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES droughts duration estimates.
+This difference underscores the importance of dataset selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES droughts duration estimates.
 Additionally, in all four graphs, the similarity between results from the two C3S datasets suggests that assumptions in the ATL dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
@@
 Wind droughts tend to be more frequent during summer, whereas solar PV droughts are more common in winter due to reduced sunlight.
-By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how model-specific assumptions and variations in capacity mix affect the overall characterisation of RES drought events.
+By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how dataset-specific assumptions and variations in capacity mix affect the overall characterisation of RES drought events.
 
@@
 ATL tends to record a slightly higher percentage of RES drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
-These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, significantly influences the apparent seasonal dynamics of RES droughts.
+These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, influences the seasonal dynamics of RES droughts.
 
-The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
+The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns similar to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% of total capacity.
+This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% \textbf{[SHOULD THIS BE 9\%?]}of total capacity.
 Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall RES drought hours.
@@
 This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV generation typically peaks in summer while wind generation predominates during winter.
-However, this integration is unable to counter the critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries.
+However, this integration is unable to counter critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries.
 
-The results presented in this study have four main limitations.
-First, the presented study uses a fixed threshold to define RES drought events, but other methods such as ones based on percentiles could be explored.
-Second, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
-Third, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, although a longer validation period would be preferable.
-Last, the source for weather data is ERA5 has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
+The results presented in this study have three main limitations.
+First, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
+Second, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, although a longer validation period would be preferable.
+Third, the source for weather data is ERA5 has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
 
 Future work is planned to extend the current analysis.
-First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change might affect RES droughts.
-Second, expanding the geographic domain of the study to include the rest of Europe would provide a more comprehensive understanding of RES droughts in an interconnected electricity grid.
+First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change and offshore wind may affect RES droughts.
+Second, expanding the geographic domain of the study to include the rest of Europe, while also including the role of electricity interconnects between countries, would provide a more comprehensive understanding of RES droughts.
 This would require extensive verification across other European countries, making it a more complex but highly relevant challenge.

### 2025-04-11 Article/RES_droughts_Morin_ReEn_R3.tex
Revert "Update RES_droughts_Morin_ReEn_R3.tex"
This reverts commit e5261afa93834a1c33017bc2040b79d91e3056d9.

@@
 While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and solar photovoltaic (PV) power.
-This challenge is amplified by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
+This challenge is amplified by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{staffell2018demand}.
 Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasising the need for a resilient energy system to meet both growing electricity demand and decarbonisation targets.
@@
 
-The wind CF in C3SE was calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
+The wind CF in C3SE was calculated using wind speeds at 100 m ($u_{100}$, $v_{100}$) and a standard wind turbine model, the Vestas V136/3450, with a fixed hub height of 100 m.
 As data on wind turbine fleet locations and specifications are difficult to obtain across Europe, C3SE assumes a homogeneous distribution of wind turbines across the ERA5 grid.
@@
 For each farm, the CF from the nearest grid point on the C3SE dataset was selected.
-A weighted average of the CF associated with each farm, using the farm's installed capacities, was used to produce the total CF time series.
+A weighted average of the CF associated with each farm, using the farm's installed capacities, was used to produce the combined CF time series.
 
@@
 Atlite allows the user to define the wind turbine power curve and PV panel model to use when converting weather variables to wind and solar PV generation.
-The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
-The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into CF.
+The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 m ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
+The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into a CF.
 The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S~GRD.
@@
 
-New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap \textbf{[ADD REFERENCE]}.
+New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap~\citep{eirgrid2023future}.
 For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
@@
 \begin{itemize}
-	\item Wind Power - based on the actual capacity at the end of 2023
-	\item Solar PV Power - based on the actual capacity at the end of 2023
-	\item Combined RES / 91W-9PV - based on the actual capacity at the end of 2023
-	\item Combined RES / 57W-43PV - based on the projected capacity for 2030
+	\item Wind Power - based on actual capacity at the end of 2023
+	\item Solar PV Power - based on actual capacity at the end of 2023
+	\item Combined RES / 91W-9PV - based on actual capacity at the end of 2023
+	\item Combined RES / 57W-43PV - based on projected capacity for 2030
 \end{itemize}
@@
 For the verification process, time-varying values of installed capacity were used to account for changes in RES development over the verification period.
-This step allowed us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
+This step allowed us to assess how well the datasets represent renewable generation profiles by comparing them against observed data.
 This validation step evaluates how well the datasets represent actual renewable energy production by comparing them against observed data.
@@
 These plots illustrate the density of CF values for each dataset, highlighting their differences and alignment with observations.
-The results indicate that ATL aligns more closely with OBS for wind, while all datasets exhibit similar distributions for solar PV.
+The results indicate that ATL aligns more closely with observations for wind, while all datasets exhibit similar distributions for solar PV.
 
@@
 
-This verification for wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
-The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts over Ireland.
+Verification of wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
+The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts across Ireland.
 On the other hand, the power curve used for both C3S~GRD and C3S~NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
 This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
-The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in their ability to reproduce the changes in CF, as seen by their similar CC scores.
-However, their ability to reproduce the actual CF values is much lower than that of ATL, with RMSE scores almost four times bigger for the two C3S datasets.
-There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of RES droughts.
+The skill scores for the three datasets (Table~\ref{tab:wind_skill_scores}) show only a small difference in their ability to reproduce the changes in CF, as seen by their similar CC scores.
+However, their ability to reproduce the actual CF values is much lower than that for ATL, with RMSE scores almost four times higher than for the two C3S datasets.
+There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to an underestimation of RES droughts.
 This highlights the need to use regionally verified datasets to assess RES droughts.
@@
 Fig.~\ref{fig:return_periods} shows the return periods for different RES drought durations, which can be used to capture the most extreme events affecting the system.
-Understanding their return periods is crucial, as extreme yet rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
+Understanding their return periods is crucial, as extreme but rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
 
@@
 In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic reduction in RES drought durations, suggesting that a more balanced share of wind and solar PV capacity can substantially mitigate the frequency of prolonged RES drought events.
-For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) increases from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
+For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) increases from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario for the ATL dataset, and from about fifteen months to around five years for the two C3S datasets.
 This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of RES drought events in a diversified energy portfolio.
@@
 However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 11\% \textbf{[SHOULD THIS BE 9\%?]}of total capacity.
+This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 9\% of total capacity.
 Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall RES drought hours.

### 2025-04-11 Article/RES_droughts_Morin_ReEn_R3.tex
Reapply "Update RES_droughts_Morin_ReEn_R3.tex"
This reverts commit c23a2f25f530743537f59824f32df1171d5de3ca.

@@
 Despite undergoing a general validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events subsequently identified~\citep{kies2021drought}.
-This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country; Ireland.
+This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country; Ireland, representative of Northwestern Europe.
 In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV capacity.
@@
 As data on wind turbine fleet locations and specifications are difficult to obtain across Europe, C3SE assumes a homogeneous distribution of wind turbines across the ERA5 grid.
-While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of climate variability on wind power generation.
+While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of weather variability on wind power generation.
 The C3SE solar PV CF was also calculated for the ERA5 grid.
 It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference solar PV plant model.
-This model incorporates empirical calculations for key system components such as optical losses, module efficiency, and inverters.
+This model incorporates empirical calculations for key system components, such as optical losses, the electrical characteristics of the power module and the power curve of the PV inverter.
 The final CF accounts for a mix of module orientations typical for each location~\citep{saintdrenan2018solar}.
@@
 
-The three datasets provide CF time series for both wind and solar PV.
+The three datasets provide time series for both wind and solar PV CF.
 In addition to analysing the CF of wind and solar PV separately, a combined CF was computed for each dataset by averaging wind and solar PV CF, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
@@
 
-The accuracy of the datasets used in this study was verified, before continuing to the analysis of RES droughts.
-For the verification process, time-varying values of installed capacity were used to account for changes in RES development over the verification period.
-This step allowed us to assess how well the datasets represent renewable generation profiles by comparing them against observed data.
-This validation step evaluates how well the datasets represent actual renewable energy production by comparing them against observed data.
+The accuracy of the datasets used in this study was verified against national generation data, before continuing to the analysis of RES droughts.
+For the verification process, time varying values of installed capacity were used to account for changes in RES development over the verification period.
+This validation step evaluates how well the datasets represent actual renewable generation profiles by comparing them against observed data.
 The overall statistical distribution of CF values for wind (2014–2023) and solar PV (2023) is presented in the violin plots in Fig.~\ref{fig:violin_plots}.
@@
 However, their ability to reproduce the actual CF values is much lower than that for ATL, with RMSE scores almost four times higher than for the two C3S datasets.
-There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to an underestimation of RES droughts.
-This highlights the need to use regionally verified datasets to assess RES droughts.
+There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to an underestimation of RES droughts, which highlights the strong motivation to use regionally verified models to assess RES droughts.
 
@@
 
-In this section, RES droughts are analysed by calculating the frequency and duration of RES drought events, the return periods for different RES drought durations, and the seasonality of RES drought events.
-Understanding the characteristics and timing of RES drought events enables system operators to optimally plan for reserve capacity requirements, ensuring grid stability and security of supply.
-Results are presented for the three datasets, allowing their differences on the characterisation of RES droughts to be clearly identified.
+In this section, RES droughts are analysed by calculating the frequency and duration of events, their return periods, and their seasonality.
+By examining both the frequency of occurrence and duration of these events, TSOs can better assess the need for alternative fossil fuel-based generation, long-duration energy storage, demand-side measures, and enhanced interconnections with neighbouring power systems to ensure adequate supply in a cost-effective manner.
+Results are presented for the three datasets, which clearly illustrate how different modelling assumptions influence the characterisation of RES droughts.
 
@@
 Fig.~\ref{fig:return_periods} shows the return periods for different RES drought durations, which can be used to capture the most extreme events affecting the system.
-Understanding their return periods is crucial, as extreme but rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
+Understanding their return periods is crucial to inform decision makers about the trade-off between the risk of extreme events and the associated costs of mitigation measures.
+Extreme but rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
 
@@
 For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the ATL dataset, compared to only five days in the C3S datasets.
-This difference underscores the importance of dataset selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES droughts duration estimates.
+This difference underscores the importance of dataset selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES drought duration estimates.
 Additionally, in all four graphs, the similarity between results from the two C3S datasets suggests that assumptions in the ATL dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
@@
 Wind droughts tend to be more frequent during summer, whereas solar PV droughts are more common in winter due to reduced sunlight.
-By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how dataset-specific assumptions and variations in capacity mix affect the overall characterisation of RES drought events.
+By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how differences in model assumptions and capacity mix influence the characterisation of RES drought events.
+In Northwestern Europe, winter droughts are primarily driven by variations in wind generation, whereas summer droughts are mainly related to solar generation.
+This seasonal analysis is critical to ensure a cost-effective balance between generation and demand throughout the year.
 
@@
 
-The effect of the integration of solar PV capacity in a wind-dominated system on RES droughts has been explored.
+The effect of adding solar PV capacity to a wind dominated power system, as is currently happening in Northwestern Europe, has been explored in the context of RES droughts.
 Our analysis has demonstrated that transitioning to a system with more equal amounts of wind and solar PV capacity reduces the occurrence of RES drought events, mitigates extreme RES drought conditions and enhances overall system resilience.
@@
 Second, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, although a longer validation period would be preferable.
-Third, the source for weather data is ERA5 has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
+Third, the source for weather data is ERA5 which has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
 

### 2025-04-14 Article/RES_droughts_Morin_ReEn_Review.tex
Updated file names for submission

--- 2025-04-11 Article/RES_droughts_Morin_ReEn_R3.tex
@@
 \begin{itemize}
-	\item Do generic datasets have sufficient skill to reliably quantify RES drought events for specific countries?
-	\item How does the integration of solar PV capacity into a predominantly wind-based system alter the characteristics of RES drought events?
+	\item Do generic datasets have sufficient skill to reliably quantify RES drought events?
+	\item How does the integration of solar PV generation into a predominantly wind-based system alter the characteristics of RES drought events?
 \end{itemize}

### 2025-04-14 Article/RES_droughts_Morin_ReEn_Title.tex
Updated file names for submission

--- 2025-04-11 Article/RES_droughts_Morin_ReEn_R3.tex
@@
-%% 
-%% Copyright 2007-2025 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
-
 \documentclass[preprint, 12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsmath package provides various useful equation environments.
-\usepackage{amsmath}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers.
-\usepackage{lineno}
-
-%% Use this package to add line brakes in URL
-\usepackage{xurl}
-
-\graphicspath{ {../Scripts/} }
 
@@
 
-%% Abstract
-\begin{abstract}
-Increasing the share of electricity produced from renewable energy sources (RES), combined with RES dependence on weather, poses a critical challenge for energy systems.
-This study investigates the importance of the balance between wind and solar photovoltaic (PV) capacity on periods of low renewable generation, known as RES droughts.
-Three different RES datasets are used to estimate the capacity factors for different scenarios of installed capacities for wind and solar PV power.
-The skill of the RES datasets is quantified by comparing capacity factor time series to observed hourly data and by assessing their representation of observed RES droughts.
-The RES datasets are used to generate a 45-year hourly time series of RES capacity factor, enabling analysis of the frequency, duration and return periods of RES droughts at a climatological scale.
-Results show the importance of using an accurate, validated RES dataset for RES drought risk assessment.
-The addition of solar PV capacity to a wind-dominated system results in a significant reduction in the frequency and duration of RES droughts, while also reducing extremes and seasonal RES drought patterns.
-These findings underscore the importance of diversification in RES capacity to enhance energy security and resilience.
-\end{abstract}
-
-%%Research highlights
-\begin{highlights}
-\item RES droughts are analysed using 45 years of hourly wind and solar PV generation data
-
-\item RES droughts from C3S-Energy and ERA5-Atlite datasets are compared
-
-\item Adding solar PV to a wind-dominated system reduces RES drought frequency and duration
-
-\item Validated RES datasets are crucial to accurately identify RES drought extremes
-
-\end{highlights}
-
-%% Keywords
-\begin{keyword}
-RES Drought \sep Wind Power \sep Solar PV Power \sep Renewable Energy Sources \sep Return Periods
-
-\end{keyword}
-
 \end{frontmatter}
 
-%% Add \usepackage{lineno} before \begin{document} and uncomment 
-%% following line to enable line numbers
-\linenumbers
-
-\section{Introduction}
-\label{sec:intro}
-
-The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022~\citep{eurostat2023share}.
-While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and solar photovoltaic (PV) power.
-This challenge is amplified by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions, both in historical~\citep{bloomfield2016} and future climates~\citep{bloomfield2021}.
-Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasising the need for a resilient energy system to meet both growing electricity demand and decarbonisation targets.
-
-RES drought events do not have a fixed definition, with various approaches present in the literature.
-One common method defines a RES drought as a period during which the average capacity factor (CF) remains below a fixed threshold for a specified duration.
-For example, Kaspar et al.~\citep{kaspar2019drought} used this method to investigate the shortfall risks of low wind and solar PV generation in Europe, with a focus on Germany, testing multiple CF thresholds and durations.
-Similarly, Mockert et al.~\citep{mockert2023drought} examined the link between weather regimes and RES droughts in Germany using a 48-hour rolling window under a threshold to define RES droughts.
-Similar fixed-threshold approaches have also been applied using CF series reconstructed through machine learning in regions such as Japan~\citep{ohba2022drought} and Hungary~\citep{mayer2023drought}.
-
-Alternative methods adjust the CF threshold dynamically over the year to account for seasonal variations in renewable production.
-Raynaud et al.~\citep{raynaud2018drought} defined RES droughts as sequences of days with renewable electricity generation below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}.
-Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models.
-Other studies have defined RES droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S.
-Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe RES drought events.
-Another method defines RES drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}.
-This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles.
-Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.
-
-In addition to examining periods of low renewable electricity generation, several studies also explore the periods when the imbalance between renewable generation and electricity demand (residual demand) is high.
-Raynaud et al.~\citep{raynaud2018drought} showed the difference between RES droughts and high residual demand events in a hypothetical fully renewable system composed of wind, solar PV and run-of-the-river hydropower.
-Similarly, Allen and Otero~\citep{allen2023drought} also defined a standardised index based on meteorological droughts to address residual demand, whose correlation to the electricity generation index is mostly negative (as expected, although quite low anticorrelations and even small positive correlations appear for some European countries).
-This index was also applied to the U.S.
-by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when demand is considered, despite showing differing results across regions.
-
-In this paper, the focus is exclusively on renewable electricity generation, to keep the focus on RES droughts driven by the weather.
-A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.
-The case study used in this paper is Ireland, a region where most RES generation comes from wind power and with ambitious targets for solar PV power expansion.
-This provides valuable insights into the potential benefits of adding solar PV installations in wind-dominated countries.
-
-RES droughts are identified using onshore wind and solar PV CF time series.
-In this study, three different datasets are used and compared, all of which are driven by the ERA5 reanalysis~\citep{hersbach2020era5}.
-Two of the datasets are part of C3S Energy (C3SE), an energy-based operational dataset produced by the EU Copernicus Climate Change Service~\citep{dubus2023energy}.
-One of the C3SE datasets provides CF time series aggregated at the national scale, while the other provides the CF time series at each grid point, at the ERA5 resolution of 0.25°.
-The third dataset produced by the authors was generated using the Atlite model~\citep{hofman2021atlite}, which converts the ERA5 atmospheric data to a generation time series using specified wind turbine and PV panel models.
-Atlite is an open-source tool developed by PyPSA~\citep{hofman2021atlite} and has been used for estimating wind and solar PV generation in order to study RES droughts in Germany~\citep{mockert2023drought}.
-
-Generic datasets for wind and solar PV CF are often used for the quantification of RES droughts.
-Despite undergoing a general validation process, they are often not fully representative of each geographical location, and can show differences in the number of RES drought events subsequently identified~\citep{kies2021drought}.
-This study evaluates the skill of a dataset developed for the European region (C3SE) when applied to a specific country; Ireland.
-In particular, the analysis explores the impact of using a generic versus a tailored dataset on RES drought assessments, in the context of a transition from a wind-dominated system to one with a greater share of solar PV capacity.
-
-The aim of this study is to answer two questions which are relevant for systems with a large share of RES generation:
-\begin{itemize}
-	\item Do generic datasets have sufficient skill to reliably quantify RES drought events for specific countries?
-	\item How does the integration of solar PV capacity into a predominantly wind-based system alter the characteristics of RES drought events?
-\end{itemize}
-
-The datasets used in this study are detailed in section~\ref{sec:data}, which describes their characteristics and relevance for evaluating RES droughts.
-Section~\ref{sec:methods} outlines the RES datasets used to simulate wind and solar PV generation and provides the methodology for defining and identifying RES drought events, including the thresholds and metrics applied.
-In section~\ref{sec:results}, the datasets are first verified against observed energy data to assess their accuracy, followed by an analysis of RES drought occurrences for two scenarios with different ratios of installed wind to solar PV capacities.
-Finally, section~\ref{sec:conclusions} offers a discussion of the results in the context of energy reliability and future planning, followed by the main conclusions and recommendations for further research.
-
-\section{Data}
-\label{sec:data}
-
-This study uses publicly available datasets to construct and validate the datasets for estimating the CF of wind and solar PV power.
-The primary data sources include: EirGrid and SONI, the transmission system operators (TSO) for the Republic of Ireland and Northern Ireland, respectively; the ERA5 reanalysis dataset; and the C3SE dataset.
-
-\subsection{Wind and solar PV Capacity and Availability}
-\label{sec:eirgrid}
-
-EirGrid, the TSO for the Republic of Ireland, and SONI, the Northern Ireland TSO, provide detailed datasets on all wind and solar PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\citep{eirgrid2023spreadsheet}.
-These datasets include information such as each farm’s installed capacity, name, and connection date.
-To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches.
-For simplicity, this data will be referred to as originating from EirGrid, as all-island data was directly obtained from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.
-
-The spreadsheet available from the EirGrid website contains two key variables: generation and availability.
-Generation and availability values are available from 2014 onward for wind power and from 2018 onward for solar PV power, although solar PV availability data only became present in the Republic of Ireland in 2023.
-Generation is the energy that a RES farm actually contributed to the grid, which may include limitations introduced by the TSO to maintain grid stability, such as constraints and curtailment.
-Availability represents the energy that would have been generated from a RES farm if no grid constraints had been applied, making it representative of the weather-related response.
-This study focuses on availability for all analyses.
-
-\subsection{Atmospheric Variables}
-\label{sec:era5}
-
-All of the datasets used in this study are driven by data from the ERA5 reanalysis~\citep{hersbach2020era5}, produced by the European Centre for Medium-Range Weather Forecasts (ECMWF).
-This global gridded dataset provides hourly atmospheric variables from 1940 to the present at a horizontal resolution of 0.25\textdegree.
-Table~\ref{tab:var_name} lists the relevant ERA5 variables.
-
-\begin{table}[h!]
-	\centering
-	\caption{ERA5 variables used to calculate wind and solar PV generation}
-	\begin{tabular}{|l|c|}
-		\hline
-		{\textbf{ERA5 name}}      & \textbf{variable} \\ \hline
-		100 metre zonal and meridional wind speed   & $u_{100}$, $v_{100}$ \\
-		2 metre temperature                         & $t2m$ \\
-		Surface net solar radiation                 & $ssr$ \\
-		Surface solar radiation downwards           & $ssrd$  \\
-		Top of atmosphere incident radiation        & $tisr$  \\
-		Total sky direct solar radiation at surface & $fdir$  \\ \hline
-	\end{tabular}
-	\label{tab:var_name}
-\end{table}
-
-\subsection{C3S Energy}
-\label{sec:c3se}
-
-The EU Copernicus Climate Change Service developed the C3S-Energy (C3SE) renewable energy dataset for Europe~\citep{dubus2023energy}, using ERA5 atmospheric variables and weather-to-energy models.
-This dataset provides hourly CF for wind and solar PV power from 1979 to the present.
-The data are available on the same grid as the ERA5 data, which has a horizontal resolution of 0.25°.
-The time series are also available for download at two aggregated scales: regional (NUTS 2) and national.
-
-The wind CF in C3SE was calculated using wind speeds at 100 metres ($u_{100}$, $v_{100}$) and a standard turbine model, the Vestas V136/3450, with a fixed hub height of 100 meters.
-As data on wind turbine fleet locations and specifications are difficult to obtain across Europe, C3SE assumes a homogeneous distribution of wind turbines across the ERA5 grid.
-While this approach does not capture the precise capacity factors reported by grid operators, it provides a well-correlated time series that effectively represents the impact of climate variability on wind power generation.
-The C3SE solar PV CF was also calculated for the ERA5 grid.
-It is derived from meteorological data, including surface solar radiation downwards ($ssrd$) and air temperature ($t2m$), using a reference solar PV plant model.
-This model incorporates empirical calculations for key system components such as optical losses, module efficiency, and inverters.
-The final CF accounts for a mix of module orientations typical for each location~\citep{saintdrenan2018solar}.
-
-\section{Methods}
-\label{sec:methods}
-
-This study analyses RES droughts using onshore wind and solar PV CF time series from three datasets: two from C3SE; one based on national-level data (C3S~NAT) and the other on grid-level data (C3S~GRD), and a third dataset derived using the Atlite model (ATL).
-
-\subsection{C3S Energy National: C3S~NAT}
-\label{sec:c3se_n}
-
-The C3S~NAT dataset is created by combining two inputs provided by C3SE at the corresponding NUTS levels: Republic of Ireland (NUTS0: IE) and Northern Ireland (NUTS2: UKN0).
-The two inputs are combined, using the actual installed capacity as weights.
-This dataset assumes that RES generation occurs at every ERA5 grid point in Ireland.
-
-\subsection{C3S Energy Gridded: C3S~GRD}
-\label{sec:c3se_g}
-
-The C3S~GRD dataset uses, as inputs, the actual locations of the RES farms in Ireland, and the CF from C3SE over the ERA5 grid.
-For each farm, the CF from the nearest grid point on the C3SE dataset was selected.
-A weighted average of the CF associated with each farm, using the farm's installed capacities, was used to produce the total CF time series.
-
-\subsection{Atlite: ATL} 
-\label{sec:atlite}
-
-The ATL dataset is produced using the Atlite model.
-Atlite allows the user to define the wind turbine power curve and PV panel model to use when converting weather variables to wind and solar PV generation.
-The Atlite model takes as inputs the locations of RES farms and ERA5 weather variables: wind speed at 100 metres ($u_{100}$, $v_{100}$) for wind generation, and radiation variables ($ssr$, $ssrd$, $tisr$, and $fdir$) along with air temperature ($t2m$) for solar PV generation.
-The output of the Atlite model is a generation time series, which is divided by the total capacity to transform it back into CF.
-The selection of the wind turbine power curve and PV panel model represents the key difference between this dataset and C3S~GRD.
-This study identifies the most appropriate wind turbine power curve to use from the 121 power curves, each at five different levels of smoothing, made available by Renewables.ninja~\citep{staffell2016wake}, and selects the PV panel model out of the options available within Atlite.
-
-\subsection{Energy Scenarios}
-\label{sec:scenarios}
-
-The three datasets provide CF time series for both wind and solar PV.
-In addition to analysing the CF of wind and solar PV separately, a combined CF was computed for each dataset by averaging wind and solar PV CF, weighted by their installed capacities at the end of 2023 (5.9 GW for wind power and 0.6 GW for solar PV power).
-This configuration is referred to as the 91W-9PV scenario, reflecting the distribution of 91\% wind and 9\% solar PV capacity.
-Given that solar PV capacity in Ireland is low in 2023, and to explore how a more balanced distribution of wind and solar PV capacities might impact RES droughts, this study also considered a second scenario, referred to as 57W-43PV, where the installed solar PV capacity is assumed to increase to 8.6 GW, while wind capacity rises to 11.45 GW.
-These values are based on targets outlined in the roadmap published by the 2024 Climate Action Plan~\citep{cap2024future}.
-This study does not include offshore wind in the analysis.
-Recent reports suggest that even by 2030, Ireland is unlikely to have any significant new offshore wind farms, with projected offshore capacity expected to remain near zero using realistic scenarios~\citep{seai2024future}.
-
-New time series were generated for both the ATL and C3S~GRD solar PV datasets, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap~\citep{eirgrid2023future}.
-For wind power, the CF time series remains unchanged, as significant shifts in the location of wind farms are not expected.
-In total, twelve CF time series were analysed in this study, six for individual wind and solar PV CF (three datasets for each source) in the 91W-9PV scenario, and an additional six time series that include the combined CF for 91W-9PV and 57W-43PV scenarios across the different datasets.
-
-It is important to note that the specific capacity values used in this study are illustrative and are not intended to reflect accurate future realities.
-Instead, they serve to explore the impact of transitioning from a wind-dominated system (91W-9PV) to a more evenly distributed system (57W-43PV).
-This approach allows for a comparative analysis between the two scenarios, assessing how the balance of RES capacity affects the occurrence of RES droughts.
-
-In summary, for each of the three datasets (ATL, C3S GRD and C3S NAT) four energy scenarios are examined:
-
-\begin{itemize}
-	\item Wind Power - based on the actual capacity at the end of 2023
-	\item Solar PV Power - based on the actual capacity at the end of 2023
-	\item Combined RES / 91W-9PV - based on the actual capacity at the end of 2023
-	\item Combined RES / 57W-43PV - based on the projected capacity for 2030
-\end{itemize}
-
-\subsection{RES Drought Definition}
-\label{sec:res_drought}
-
-In this study, a RES drought event was defined as occurring when the 24-hour moving average of CF remains below a fixed threshold of 0.1 for a period of longer than 24 hours.
-By using a 24-hour moving average, fewer but longer-lasting events were captured compared to using the raw CF time series, which can be more sensitive to short-term fluctuations.
-The 24-hour rolling average also avoids potential masking of day-long events due to their start time.
-A fixed threshold approach was chosen in this study to enable consistent inter-comparison between datasets.
-
-\begin{figure}[ht!]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_methodology.pdf}
-	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021.
-The black dashed line indicates the CF threshold.
-The grey bar shows the period identified as a wind drought under our definition}
-	\label{fig:find_res_droughts}
-\end{figure}
-
-The moving average approach smooths out short-term fluctuations, so that brief periods above the threshold do not interrupt an otherwise continuous low-CF period (Fig.~\ref{fig:find_res_droughts}).
-This means that a single hour above the threshold does not "break" a RES drought event if it is surrounded by prolonged low-generation hours.
-As a result, fewer but longer-lasting RES drought events are identified, which may better reflect actual conditions where energy supply constraints persist over extended periods.
-
-\section{Results}
-\label{sec:results}
-
-\subsection{Verification}
-\label{sec:verification}
-
-The accuracy of the datasets used in this study was verified, before continuing to the analysis of RES droughts.
-For the verification process, time-varying values of installed capacity were used to account for changes in RES development over the verification period.
-This step allowed us to assess how well the datasets represent the production of renewable energy by comparing them against observed data.
-This validation step evaluates how well the datasets represent actual renewable energy production by comparing them against observed data.
-The overall statistical distribution of CF values for wind (2014–2023) and solar PV (2023) is presented in the violin plots in Fig.~\ref{fig:violin_plots}.
-These plots illustrate the density of CF values for each dataset, highlighting their differences and alignment with observations.
-The results indicate that ATL aligns more closely with OBS for wind, while all datasets exhibit similar distributions for solar PV.
-
-\begin{figure}[ht!]
-	\centering
-	\includegraphics[width=\textwidth]{violin_plots_verification.pdf}
-	\caption{Violin plots of CF distributions for a)  wind and b) solar PV for the Observations (grey) and the three datasets: ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
-The black dot shows the median values, while the black vertical lines represent the first and third quartiles.
-The black dashed line indicates the threshold of 0.1 used in the study to identify RES droughts}
-	\label{fig:violin_plots}
-\end{figure}
-
-\subsubsection{Wind Energy}
-\label{sec:wind_verification}
-
-The C3S datasets use the Vestas V136/3450 wind turbine power curve (Fig.~\ref{fig:power_curve}a).
-The Atlite model allows the user to specify the power curve.
-We considered the 121 power curves available for download from Renewables.ninja~\citep{staffell2016wake}.
-For each power curve, Renewables.ninja also provides four associated smoothed power curves.
-The smoothing is done using a Gaussian filter with different standard deviations that depend on the wind speed.
-A separate wind CF time series for Ireland was generated for each of the wind turbine power curves and smoothing levels.
-
-The performance of each CF time series is then assessed based on four skill scores: correlation coefficient (CC), root mean square error (RMSE), mean bias error (MBE), and the percentage of overlap.
-The percentage of overlap quantifies the similarity between the observed and modelled distributions.
-It is a positively oriented skill score, where 100\% shows full agreement between the two distributions, and 0\% indicates no overlap.
-The histograms of hourly CF values for the most recent decade (2014-2023) are used to calculate this skill score.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
-	\caption{a) Power curves of the Enercon E112.4500 with a 0.3w smoothing filter used by the ATL dataset (orange) and the Vestas V136/3450 used by the two C3S datasets (blue) b) Histograms of wind CF for Ireland for the ATL dataset (orange), the C3S datasets (blue) and Observed (shaded)}
-	\label{fig:power_curve}
-\end{figure}
-
-Based on these metrics, the most representative power curve for Ireland is the Enercon E112.4500 power curve with the $0.3w$ smoothing filter.
-The smoothing of the wind turbine power curve represents losses associated with each turbine, as well as losses such as wake effects between turbines, which are important when modelling wind energy on larger spatial scales.
-The histogram in Fig.~\ref{fig:power_curve}b shows that the C3SE power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed ATL power curve more closely follows the observed wind availability data.
-This is further supported by the percentage of overlap which is higher for ATL (97.2\%) than for C3SE (83.2\%), indicating better agreement with observed data.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{verification_wind_contour.png}
-	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) ATL, b) C3S~GRD and c) C3S~NAT datasets}
-	\label{fig:wind_verification_contour}
-\end{figure}
-
-The effect of the difference between the power curves is also visible in Fig.~\ref{fig:wind_verification_contour}, which shows a density plot of wind CF values.
-The two C3S datasets are shown to overestimate the observed CF, whereas the ATL dataset is in good agreement with the observed data.
-The skill scores presented in Table~\ref{tab:wind_skill_scores} show that ATL performs better than the two C3S datasets for all of the skill scores.
-
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{l|lll|}
-		\cline{2-4}
-		& \textbf{ATL} & \textbf{C3S~GRD} & \textbf{C3S~NAT} \\ \hline
-		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.981           & 0.972            & 0.970            \\ \hline
-		\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.045           & 0.177            & 0.162            \\ \hline
-		\multicolumn{1}{|l|}{\textbf{MBE}}   & -0.003          & 0.137            & 0.121            \\ \hline
-	\end{tabular}
-	\caption{Skill scores for wind power for the three datasets compared to observed data}
-	\label{tab:wind_skill_scores}
-\end{table}
-
-Fig.~\ref{fig:bar_number_events_verification_wind} shows the average annual number of wind drought events during the 2014 to 2023 validation period.
-The figure reveals that ATL presents the best overall agreement with the observed frequency and duration of wind drought events.
-This pattern is particularly evident for shorter-duration events, which are the most frequent.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
-	\caption{Average annual number of wind drought events for ATL (red), C3S~GRD (blue), C3S~NAT (purple), and the observed data (black outline).
-The wind droughts are identified from 2014 to 2023, considering the actual capacity of the system at any given time}
-	\label{fig:bar_number_events_verification_wind}
-\end{figure}
-
-This verification for wind generation data highlights the importance of selecting a representative wind turbine power curve for the region being analysed.
-The ATL dataset, which uses a representative wind turbine power curve, is skilled at reproducing wind CF and RES droughts over Ireland.
-On the other hand, the power curve used for both C3S~GRD and C3S~NAT is not representative for Ireland, as it severely overestimates generation, underestimating the occurrence of RES droughts.
-This highlights a problem with using generalised datasets for analysing RES droughts: biases severely affect their ability to accurately reproduce RES drought events.
-The skill scores for the three datasets (Tab.~\ref{tab:wind_skill_scores}) show only a small difference in their ability to reproduce the changes in CF, as seen by their similar CC scores.
-However, their ability to reproduce the actual CF values is much lower than that of ATL, with RMSE scores almost four times bigger for the two C3S datasets.
-There is a clear bias towards an overestimation of CF, seen in the MBE values, which leads to the underestimation of RES droughts.
-This highlights the need to use regionally verified datasets to assess RES droughts.
-
-\subsubsection{Solar PV Energy}
-\label{sec:pv_verification}
-
-The Atlite model allows the user to select certain PV panel characteristics.
-In this study, the three PV panel types available in the Atlite model were considered (CSi, CdTe, Kaneka).
-Following the same methodology as in the previous section, the three available models were compared using four skill scores (CC, RMSE, MBE, and the percentage of overlap).
-Based on the best-performing metrics, the Beyer PV panel model was selected~\citep{beyer2004pv}, using the Kaneka Hybrid panel option.
-For all solar PV farm locations, the azimuth angle is fixed at 180\textdegree (due south), and the optimal tilt angle option is applied.
-
-The solar PV installed capacity available on the spreadsheets from EirGrid represents the Maximum Export Capacity (MEC) and does not accurately reflect the installed solar PV capacity.
-To enable actual solar PV generation potential to be modelled correctly, installed capacities were set at 1.4 times the MEC values.
-This scaling factor was estimated by analysing proprietary data from individual solar PV farms provided by EirGrid, which showed that, on average, assuming that the installed capacities of farms exceed their MEC values by 40\% yields the best agreement with the observed availability.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\textwidth]{verification_pv_contour.png}
-	\caption{Solar PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) ATL, b) C3S~GRD and c) C3S~NAT datasets}	
-	\label{fig:solar_verification_contour}
-\end{figure}
-
-Fig.~\ref{fig:solar_verification_contour} shows that the three datasets have a similar tendency to overestimate the CF compared to the observed values, especially for high CF values.
-The skill scores presented in Table~\ref{tab:pv_skill_scores} indicate that C3S~GRD and C3S~NAT perform better than ATL for solar PV CF, with lower RMSE and MBE, and higher CC scores.
-This may be due to the statistical approach taken by C3SE for the orientation of the PV panels.
-
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{l|lll|}
-		\cline{2-4}
-		& \textbf{ATL} & \textbf{C3S~GRD} & \textbf{C3S~NAT} \\ \hline
-		\multicolumn{1}{|l|}{\textbf{CC}}   & 0.921           & 0.931            & 0.931            \\ \hline
-		\multicolumn{1}{|l|}{\textbf{RMSE}} & 0.119           & 0.090            & 0.113            \\ \hline
-		\multicolumn{1}{|l|}{\textbf{MBE}}   & 0.046           & 0.027           & 0.021           \\ \hline
-	\end{tabular}
-	\caption{Skill scores for solar PV CF for the three datasets compared to observed data}
-	\label{tab:pv_skill_scores}
-\end{table}
-
-Fig.~\ref{fig:bar_number_events_verification_pv} shows the number of solar PV drought events during the 2023 validation period across different duration ranges.
-The figure reveals partial agreement between the three datasets and the observed data, with consistent results noticed for duration ranges of 1-2, 3-4, 7-8, and 8+ days.
-However, discrepancies appear in the other ranges, where the datasets diverge from the observed data.
-The main challenge in validating solar PV data stems from the recent installation of a large share of Ireland’s solar PV capacity, with over 65\% of the total solar PV capacity installed in 2023.
-This results in uncertainties in solar PV generation data and the actual generating capacity in the first few months after each farm is connected.
-Overall, C3S~GRD performs slightly better than the other datasets in reproducing observed solar PV drought events.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
-	\caption{Number of solar PV drought events for ATL (red), C3S~GRD (blue), and C3S~NAT (purple) and the observed data (black outline).
-The solar PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
-	\label{fig:bar_number_events_verification_pv}
-\end{figure}
-
-\subsection{Analysis}
-\label{sec:analysis}
-
-In this section, RES droughts are analysed by calculating the frequency and duration of RES drought events, the return periods for different RES drought durations, and the seasonality of RES drought events.
-Understanding the characteristics and timing of RES drought events enables system operators to optimally plan for reserve capacity requirements, ensuring grid stability and security of supply.
-Results are presented for the three datasets, allowing their differences on the characterisation of RES droughts to be clearly identified.
-
-RES drought events are evaluated under two different scenarios with fixed installed capacities: the 91W-9PV scenario, with 5.9 GW of wind capacity and 0.6 GW of solar PV capacity; and the 57W-43PV scenario, where wind capacity comprises 11.45 GW and solar PV capacity increases to 8.6 GW.
-Both scenarios were driven by 45 years of ERA5 data.
-Using the RES drought identification process described in Section~\ref{sec:res_drought}, wind and solar PV droughts are first analysed separately before presenting the results for combined (wind + solar PV) RES droughts under both scenarios.
-
-\subsubsection{Annual Number of RES Droughts}
-
-The first part of the analysis examines the annual number of RES drought events.
-When only wind energy is considered (Fig.~\ref{fig:boxplot_number_events}a), the number of RES drought events decreases as the duration range increases, with very few events lasting more than seven days.
-In contrast, for solar PV energy (Fig.~\ref{fig:boxplot_number_events}b), RES drought frequency declines from one to eight days and then slightly increases for longer durations.
-This behaviour is attributable to Ireland's high-latitude location, where reduced sunlight in winter (from November to March) leads to consistently low solar PV output.
-
-Moreover, the comparison between wind and solar PV results indicates that the median, first, and third quartiles for solar PV are consistently higher than or equal to those for wind.
-This is expected, given that solar PV generation is inherently lower, zero at night, and limited by the solar cycle.
-When wind and solar PV are combined under the 91W-9PV scenario (Fig.~\ref{fig:boxplot_number_events}c), the results closely mirror those of wind alone, due to the dominance of wind power in the current energy mix.
-However, in the 57W-43PV scenario (Fig.~\ref{fig:boxplot_number_events}d), a marked reduction in RES drought events is observed across all datasets, with a decrease of the total number of events of 56\% for ATL, 52\% for C3S~GRD, and 50\% for C3S~NAT, demonstrating the beneficial effects of a more equal share of wind and solar PV capacity.
-
-The consistently higher RES drought counts reported by the ATL dataset, compared to the C3S datasets, underscore the importance of wind turbine power curve representation when quantifying RES droughts.
-Whereas the three datasets agree on the overall effect of balancing the share of wind and solar PV generation, they differ at a quantitative level, which has crucial implications for energy planning.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
-	\caption{Average annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red), C3S~GRD (blue), and C3S~NAT (purple).
-The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events.
-The boxes display the first and third quartiles and the median is marked by a black line.
-The whiskers indicate the 5th and 95th percentiles}
-	\label{fig:boxplot_number_events}	
-\end{figure}
-
-\subsubsection{Return Periods of RES Drought Duration}
-
-RES drought events identified over the 45-year period were used to calculate the return periods for different RES drought durations.
-A return period is the estimated average time interval between events of a specified duration (not to be confused with the frequency of their occurrence within a fixed time frame).
-Fig.~\ref{fig:return_periods} shows the return periods for different RES drought durations, which can be used to capture the most extreme events affecting the system.
-Understanding their return periods is crucial, as extreme yet rare RES droughts pose the toughest challenge to energy security by placing significant strain on the conventional backup sources necessary to maintain security of supply during these events.
-
-The duration of wind droughts (Fig.~\ref{fig:return_periods}a) increases in a log-linear fashion across the three datasets.
-The log-linear trend indicates a predictable relationship between wind drought duration and occurrence, with longer wind droughts becoming exponentially less likely as duration increases.
-In the case of solar PV droughts (Fig.~\ref{fig:return_periods}b), Atlite behaves differently than the two C3S datasets.
-The ATL dataset show a generally log-linear increase.
-For C3S~GRD and C3S~NAT, the duration of PV droughts increases in a log-linear pattern for events lasting less than 16 days.
-Beyond this duration, there is a sharp rise in solar PV drought duration for events up to a one-year return period.
-This sudden increase again reflects the impact of extended periods of low PV generation during winter in Ireland.
-The difference between the ATL and the C3S results arises from differences in the datasets near the threshold of 0.1 CF.
-ATL remains slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented RES drought events.
-In contrast, C3S~GRD and C3S~NAT tend to fall below the threshold in similar conditions, resulting in longer continuous RES drought periods, especially during winter.
-
-Under the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), the combined RES drought return periods mirror those for wind alone, reflecting the dominance of wind in the current energy mix.
-In contrast, the 57W-43PV scenario (Fig.~\ref{fig:return_periods}d) shows a dramatic reduction in RES drought durations, suggesting that a more balanced share of wind and solar PV capacity can substantially mitigate the frequency of prolonged RES drought events.
-For example, the return period for a five-day RES drought event (shown by the vertical dashed lines in Fig.~\ref{fig:return_periods}) increases from roughly six months for the 91W-9PV scenario, to four years for the 57W-43PV scenario in the ATL dataset, and from about fifteen months to around five years in the two C3S datasets.
-This result indicates that the complementarity between wind and solar PV plays a crucial role in reducing the occurrence of RES drought events in a diversified energy portfolio.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
-	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red triangle), C3S~GRD (blue circle), and C3S~NAT (purple square).
-The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it.
-The horizontal dashed line marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
-	\label{fig:return_periods}
-\end{figure}
-
-Across Fig.~\ref{fig:return_periods}a, c, and d, the return periods in the ATL dataset are consistently higher than those in the two C3S datasets.
-For instance, in the 91W-9PV scenario (Fig.~\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the ATL dataset, compared to only five days in the C3S datasets.
-This difference underscores the importance of dataset selection when quantifying RES droughts, as each dataset’s assumptions and parametrisations significantly influence RES droughts duration estimates.
-Additionally, in all four graphs, the similarity between results from the two C3S datasets suggests that assumptions in the ATL dataset, such as wind turbine power curve selection and PV panel specifications, have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.
-
-The return periods calculated from the three datasets show large differences, in particular for the more extreme events with longer return periods.
-The C3S datasets produce shorter RES drought durations for these events, which would have the largest impact on the power system.
-This shows that system planning based on the wrong datasets could yield an underestimation of the duration of extreme RES droughts, potentially leading to shortages linked to undersized reserve capacity.
-
-\subsubsection{Seasonal Distribution of RES Droughts}
-
-The seasonal analysis of RES droughts is based on the percentage of hours in each month classified as part of a RES drought event.
-Wind droughts tend to be more frequent during summer, whereas solar PV droughts are more common in winter due to reduced sunlight.
-By comparing these seasonal patterns across different datasets and energy scenarios, this study examines how dataset-specific assumptions and variations in capacity mix affect the overall characterisation of RES drought events.
-
-For the wind-only scenario (Fig.~\ref{fig:res_droughts_seasonality}a), the ATL dataset exhibits a pronounced seasonal pattern, with about 24\% of summer hours (June, July, August) identified as RES droughts compared to only 4\% in winter (December, January, February).
-This strong seasonal signal is less evident in the C3S datasets, which suggests that the differences in the underlying wind power curves play a significant role.
-In ATL, CF near or below the 0.1 threshold occurs at relatively higher wind speeds, resulting in a higher count of RES drought hours during the summer months.
-In contrast, solar PV droughts (Fig.~\ref{fig:res_droughts_seasonality}b) display an opposite seasonal trend.
-Across all datasets, over 60\% of winter hours are classified as solar PV droughts, reflecting the naturally low solar irradiance in Ireland during winter.
-
-ATL tends to record a slightly higher percentage of RES drought hours for wind and a marginally lower percentage for solar PV relative to the C3S datasets.
-These differences highlight how dataset-specific assumptions, such as the treatment of wind turbine power curves and PV panel characteristics, influences the seasonal dynamics of RES droughts.
-
-The 91W-9PV scenario (Fig.~\ref{fig:res_droughts_seasonality}c) shows patterns similar to the ones for wind droughts (Fig.~\ref{fig:res_droughts_seasonality}a).
-However, in the 91W/9PV scenario, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario.
-This reduction can be explained by the contribution of solar PV generation during the summer months in the 91W-9PV scenario, even though it constitutes only 9\% of total capacity.
-Since the number of RES drought hours for solar PV in summer is near zero, this small contribution has a noticeable impact on reducing overall RES drought hours.
-In the 57W-43PV scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency.
-Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for ATL, from 8\% to 3\% for C3S~GRD, and from 9\% to 4\% for C3S~NAT.
-The balanced mix of wind and solar PV power in this scenario reduces the seasonal signal overall and significantly decreases the percentage of RES drought hours in the summer.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
-	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~Solar PV, c)~91W-9PV and d)~57W-43PV for ATL (red dotted), C3S~GRD (blue dashed), and C3S~NAT (purple solid).
-The x-axis represents the month of the year, and the y-axis indicates the percentage of hours.
-Lines correspond to the median values and the area between the first and third quartiles is shaded.
-Note the different y-axis scale for b).}
-	\label{fig:res_droughts_seasonality}
-\end{figure}
-
-The seasonal variations of RES droughts observed in this study have important implications for energy planning.
-Energy demand peaks in winter for Northern European countries, making the seasonality of RES droughts critical for the sizing of reserve capacity.
-Our results show that selecting the wrong dataset could severely underestimate RES droughts during winter months, thereby affecting the reliability of the energy system during critical periods.
-Additionally, the integration of large shares of solar PV in the system leads to a generalised reduction of RES droughts, yet winter months present a slight increase.
-The natural limitations of solar PV lead to inevitably higher reserve capacity needs during winter months as reliance on RES increases.
-These types of insights are essential to develop targeted strategies that enhance grid resilience and ensure a stable energy supply throughout the year.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-This study aimed to answer two key questions: Do generic datasets have sufficient skill to reliably quantify RES drought events?
-How does the integration of solar PV into a predominantly wind-based system alter the characteristics of RES droughts?
-To address these questions, three datasets were compared: two derived from the European C3S-Energy dataset, and one developed by the authors.
-The datasets derived from C3S-Energy differ in their assumptions, one assumes a homogeneous distribution of wind and solar PV capacity across the region, while the other includes the actual locations of RES farms.
-The dataset developed by the authors uses a regionally validated model which accounts for farm locations and uses tailored wind and solar PV models selected to represent the actual generation.
-
-Our results demonstrate that datasets without regional validation misrepresent the frequency and duration of RES drought events due to their limited ability to reproduce the observations.
-The inclusion of wind and solar PV farm locations has limited impact on RES drought analysis compared to the choice of wind turbine power curves and solar PV models.
-Whereas all three datasets capture broad trends in the duration and seasonality of RES drought events, the actual number of events is consistently underestimated by the non-validated datasets.
-This effect becomes clearer for extreme events, as not using regionally validated datasets can yield an overestimation of the return periods of RES droughts.
-This can lead to insufficient reserve capacity planning and potential risks to grid stability and security of supply.
-
-The effect of the integration of solar PV capacity in a wind-dominated system on RES droughts has been explored.
-Our analysis has demonstrated that transitioning to a system with more equal amounts of wind and solar PV capacity reduces the occurrence of RES drought events, mitigates extreme RES drought conditions and enhances overall system resilience.
-This improvement is attributed to the complementary nature of wind and solar PV generation, as solar PV generation typically peaks in summer while wind generation predominates during winter.
-However, this integration is unable to counter critical winter RES droughts, which coincide with the strongest electricity demand in Northern European countries.
-
-The results presented in this study have three main limitations.
-First, the definition of RES droughts based on generation does not consider the important role of demand, which could be of interest to system operators.
-Second, recent solar PV capacity expansions have changed the generation profile, limiting solar PV data for model training to a single year, although a longer validation period would be preferable.
-Third, the source for weather data is ERA5 has limited spatial resolution, an issue that can be addressed once higher resolution datasets become available.
-
-Future work is planned to extend the current analysis.
-First, climate projection data will be integrated with different energy scenarios, incorporating the addition of offshore wind, to better understand how climate change and offshore wind may affect RES droughts.
-Second, expanding the geographic domain of the study to include the rest of Europe, while also including the role of electricity interconnects between countries, would provide a more comprehensive understanding of RES droughts.
-This would require extensive verification across other European countries, making it a more complex but highly relevant challenge.
-
-\section*{Data Availability}
-
-The ERA5 data can be obtained from the Climate Data Store (\url{https://doi.org/10.24381/cds.adbb2d47}).
-The C3SE dataset is also available from the Climate Data Store (\url{https://doi.org/10.24381/cds.4bd77450}).
-Information on wind and solar PV farms in Ireland can be obtained from the EirGrid website (\url{https://www.eirgrid.ie/grid/system-and-renewable-data-reports}).
-The Atlite model used in this study is open-source and can be found on GitHub (\url{https://github.com/pypsa/atlite}).
-The data and code required to reproduce the analysis in this article will be made available upon acceptance of the manuscript in a public GitHub repository.
-
-\section*{Acknowledgments}
-
-The research conducted in this publication was funded by Science Foundation Ireland and co-funding partners under grant number 21/SPP/3756 through the NexSys Strategic Partnership Programme.
-
-\bibliographystyle{elsarticle-num-names}
-\bibliography{RES_droughts_Morin_ref}
-
 \end{document}
-
-\endinput
